### 2024-04-22 tex/aq_sankey.tex
Org -> Make

\documentclass[11pt]{article}
\usepackage{sankey}
\usepackage{tikz}
\usepackage{pgf}
\pagestyle{empty}

\begin{document}
\begin{tikzpicture}[remember picture, overlay, shift={(5.5,-5.5)}]

  \begin{sankeydiagram}
    %% \sankeyset{debug=true}

    \colorlet{water}[rgb]{cyan!75!blue!50!white}
    \colorlet{ice}[rgb]{gray!20!gray!20!gray!20}
    \colorlet{air}[rgb]{gray!1!gray!1!gray!1}
    \colorlet{red}[rgb]{red!50!orange!75!white}

    \sankeyset{
      every node/.style={angle=-90},
      start style=arrow,end style=arrow,
      %% fill/.style={fill=yellow!10,line width=0pt,draw=yellow!10},
      ratio=1mm/50,
      minimum radius=1mm,
      water/.style={fill/.style={draw=water,line width=0pt,fill=water,},},
      ice/.style={fill/.style={draw=ice,line width=0pt,fill=ice,},},
      air/.style={fill/.style={draw=air,line width=0pt,fill=air,},},
      ice-to-water/.style={fill/.style={fill=none, top color=ice, bottom color=water},},
      ice-to-air/.style={fill/.style={fill=none, top color=ice, bottom color=air},},
      water-to-air/.style={fill/.style={fill=none, top color=water, bottom color=air},},
      air-to-ice/.style={fill/.style={fill=none, top color=air, bottom color=ice},},
      red/.style={fill/.style={fill=none, top color=red, bottom color=red},},
      red-to-ice/.style={fill/.style={fill=none, top color=red, bottom color=ice},},
      water-to-ice/.style={fill/.style={fill=none, top color=water, bottom color=ice},},
      debug label/.style={
        overlay,
        draw,
        font=\ttfamily,
        text=debug color,text opacity=1,
        inner sep=.1em,
        fill=white,fill opacity=1,
        rounded corners=.1em,
        node contents={\name (\qty)},
      },
    }

    % storage of labels
    \newcommand\LabSet[2]{% node name, label
      \expandafter\edef\csname#1@Lab\endcsname{#2}}
    \newcommand\Lab[1]{% node name
      \csname#1@Lab\endcsname}
    % storage of quantities
    \newcommand\QSet[2]{% node name, quantity
      \expandafter\edef\csname#1@Q\endcsname{\fpeval{#2}}}
    \newcommand\Q[1]{% node name
      \csname#1@Q\endcsname}
    % all nodes with their name, label and quantity
    %% \newcommand\Q[1]{\sankeygetnodeqty{#1}}

    \sankeyset{
      def data/.code args={#1/#2/#3}{% node name/label/values
        \LabSet{#1}{#2}
        \QSet{#1}{#3}
        \typeout{#1: \Q{#1} Gt (\Lab{#1})}
      },
      def data/.list={
        %% INSERT_CSV_HERE
        {MELT/Surface melt/\Q{RU}+\Q{RFZ}},
        {ICE-FRET2/ICE and some FRET/\Q{ICE}+0.5*\Q{FRET}},
        {SUB-FRET2/SUB and some FRET/\Q{SUB}+0.5*\Q{FRET}},
        {GROSS/Gross\\accumulation/\Q{DP}+\Q{RFZ}+\Q{SF}+\Q{RF}+\Q{CD}},
        {IL/Ice Loss/
          (\Q{SU}+\Q{RU}+\Q{BM}+\Q{DYN}-(\Q{SF}+\Q{RF}+\Q{CD}+\Q{DP}))},
        {ML/Net mass Loss/\Q{FRET}+\Q{IL}}
      },
    }

    \sankeyset{ice} % color

    %% GROSS inputs
    \sankeynode{
      name=GROSS,
      quantity=\Q{GROSS}+\Q{IL},
      at={0,0},
      % inputs
      forked={
        \Q{IL}/GROSS-from-IL,  % Ice loss
        \Q{SF}/GROSS-from-SF,   % snowfall
        \Q{DP}/GROSS-from-DP,   % deposition
        \Q{CD}/GROSS-from-CD,   % condensation
        \Q{RF}/GROSS-from-RF,   % rainfall
        \Q{RFZ}/GROSS-from-RFZ  % refrozen melt
      },
    }
    \sankeyadvance{GROSS}{2cm}
    %% GROSS outputs
    \sankeyfork{GROSS}{
      \Q{SU}/GROSS-to-SU,
      \Q{DYN}/GROSS-to-DYN,
      \Q{BM}/GROSS-to-BM,
      \Q{MELT}/GROSS-to-MELT}

    \sankeynode{name=BM, quantity=\Q{BM}, at={[xshift=-3mm, yshift=-3cm]GROSS-to-BM.center}, anchor=center}
    \sankeynode{name=MELT, quantity=\Q{MELT}, at={[xshift=-4mm, yshift=0mm]BM.center}, anchor=left}
    \sankeynode{name=DYN, quantity=\Q{DYN}, at={[xshift=3mm, yshift=0mm]BM.center}, anchor=right}
    \sankeynode{name=DYN-FRET, quantity=\Q{DYN}+\Q{FRET}, at={[xshift=3mm, yshift=-3cm]DYN.center}, anchor=center}

    %% BM
    \sankeydubins[ice-to-water]{GROSS-to-BM}{BM}
    \sankeyadvance[water]{BM}{3cm}
    \sankeyadvance[water]{BM}{3cm}
    \sankeyend[water]{BM}

    %% %% MELT
    \sankeydubins[ice-to-water]{GROSS-to-MELT}{MELT}
    \sankeyfork{MELT}{\Q{RU}/RU-from-MELT, \Q{RFZ}/RFZ-from-MELT}
    \sankeyadvance[water]{RU-from-MELT}{3cm}
    \sankeynode{name=RU-GROSS, quantity=\Q{RU}+\Q{RF-LIQ}, at={RU-from-MELT.left}, anchor=left,
      forked={\Q{RU}/other, \Q{RF-LIQ}/RF-LIQ-to-RF-NET}}
    \sankeyadvance[water]{RU-GROSS}{1.5cm}
    \sankeyfork{RU-GROSS}{\Q{RU}+\Q{RF-LIQ}-\Q{EV}/RU-NET, \Q{EV}/EVAP}
    \sankeyadvance[water]{RU-NET}{1.5cm}
    \sankeyend[water]{RU-NET}

    %% RFZ
    \sankeynode{ name=RFZ-waypoint0, quantity=\Q{RFZ}, angle=90,
      at={[xshift=-20mm, yshift=0cm]RFZ-from-MELT.right}, anchor=right}
    \sankeynode{ name=RFZ-waypoint1, quantity=\Q{RFZ}, angle=90,
      at={[xshift=-20mm, yshift=0cm]GROSS-from-RFZ.right}, anchor=right}
    \sankeydubins[water]{RFZ-from-MELT}{RFZ-waypoint0}
    \sankeydubins[ice-to-water]{RFZ-waypoint0}{RFZ-waypoint1}
    \sankeydubins[ice]{RFZ-waypoint1}{GROSS-from-RFZ}

    %% DYN
    \sankeydubins{GROSS-to-DYN}{DYN}
    \sankeyoutin{DYN}{DYN-FRET}
    \sankeynode{name=FRET-DYN, quantity=\Q{FRET}, at={DYN-FRET.center}, anchor=center}
    \sankeyadvance{DYN-FRET}{5mm}
    \sankeyfork{DYN-FRET}{\Q{ICE-FRET2}/ICE-from-DYN, \Q{SUB-FRET2}/SUB}
    \sankeyadvance[ice-to-water]{SUB}{2.5cm}
    \sankeyend[water]{SUB}
    \sankeynode{name=ICE, quantity=\Q{ICE-FRET2}, at={[xshift=3mm, yshift=0mm]SUB.left}, anchor=right}
    \sankeydubins{ICE-from-DYN}{ICE}
    \sankeyend{ICE}

    %% EV
    \sankeynode{name=EV-out, quantity=\Q{EV}, at={[xshift=-10mm, yshift=-10mm]EVAP.center}, anchor=center, angle=180}
    \sankeydubins[water-to-air]{EVAP}{EV-out}
    \sankeyend[air]{EV-out}

    %% SU
    \sankeynode{name=SU-out, quantity=\Q{SU},
      at={[xshift=10mm, yshift=-10mm]GROSS-to-SU.center}, anchor=center, angle=0}
    \sankeydubins[ice-to-air]{GROSS-to-SU}{SU-out}
    \sankeyend[air]{SU-out}

    %%% INPUTS

    \sankeynode{name=SF, quantity=\Q{SF}, at={[xshift=0, yshift=3cm]GROSS-from-SF.center}, align=center}
    \sankeynode{name=DP, quantity=\Q{DP}, at={[xshift=-3mm, yshift=0]SF.right}, align=left}
    \sankeynode{name=CD, quantity=\Q{CD}, at={[xshift=-3mm, yshift=0]DP.right}, align=left}
    \sankeynode{name=RF-NET, quantity=\Q{RF}+\Q{RF-LIQ}, at={[xshift=-6mm, yshift=0]CD.right}, align=left}
    \sankeystart[water]{RF-NET}
    \sankeyadvance[water]{RF-NET}{0.5cm}

    \sankeydubins{SF}{GROSS-from-SF}
    \sankeydubins[air-to-ice]{DP}{GROSS-from-DP}
    \sankeydubins{CD}{GROSS-from-CD}
    \sankeyfork{RF-NET}{\Q{RF}/RFZ-from-RF-NET, \Q{RF-LIQ}/RF-LIQ-from-RF-NET}
    \sankeydubins[water-to-ice]{RFZ-from-RF-NET}{GROSS-from-RF}
    \sankeynode{name=RF-wpt1, quantity=\Q{RF-LIQ}, at={[xshift=-10mm]GROSS-to-MELT.right}, align=right}
    \sankeynode{name=RF-wpt2, quantity=\Q{RF-LIQ}, at={[xshift=-10mm]RFZ-from-MELT.right}, align=right}
    \sankeydubins[water]{RF-LIQ-from-RF-NET}{RF-wpt1}
    \sankeydubins[water]{RF-wpt1}{RF-wpt2}
    \sankeydubins[water]{RF-wpt2}{RF-LIQ-to-RF-NET}

    \sankeystart[air]{DP}
    \sankeystart{SF}
    \sankeystart{CD}

    \sankeyset{ice}
    \sankeynode{ name=ML, quantity=\Q{FRET}+\Q{IL}, at={[xshift=8mm]SF.left}, anchor=right}
    \sankeystart[red]{ML}
    \sankeyadvance[red-to-ice]{ML}{1cm}
    \sankeyfork{ML}{\Q{FRET}/FRET, \Q{IL}/IL}
    \sankeydubins[minimum radius=2mm]{IL}{GROSS-from-IL}
    \sankeyadvance{FRET}{2cm}
    \sankeyturnleft{FRET}{30}
    \sankeyadvance{FRET}{2cm}
    \sankeyturnleft{FRET}{-30}
    \sankeyadvance{FRET}{3cm}
    \sankeydubins{FRET}{FRET-DYN}

    %% DRIFT
    \sankeynode{name=DRIFT-OUT, quantity=33, at={(3mm, -8mm)}, angle=150}
    \sankeynode{name=DRIFT-IN, quantity=33, at={(-3mm, -8mm)}, angle=210}
    \sankeydubins[minimum radius=5mm]{DRIFT-OUT}{DRIFT-IN}
    \sankeystart{DRIFT-OUT}
    \sankeyend{DRIFT-IN}
    \node[anchor=center, align=center, rotate=0, shift=({0,0})] at (0,-15mm) {Drifting (?)};

    \node[anchor=south east, align=center, shift=({2mm,-3mm})] at (EV-out.right) {Evaporation\\(\Q{EV})};

    \node[anchor=center, align=center, shift=({0mm,-1cm})] at (RU-NET.center) {Net Runoff\\(\sankeygetnodeqty{RU-NET})};

    \node[anchor=center, align=center, shift=({0mm,-1cm})] at (SUB.center) {Submarine\\melt (\sankeygetnodeqty{SUB})};

    \node[anchor=center, align=center, shift=({0mm,-1cm})] at (ICE.center) {Calving\\(\sankeygetnodeqty{ICE})};

    \node[anchor=center, align=left, rotate=-90, shift=({1.5cm, -2mm})] at (BM.center) {Basal\\melting (\sankeygetnodeqty{BM})};

    \node[anchor=center, align=center, shift=({0mm,0cm})] at (DYN.center) {Dynamics\\(\sankeygetnodeqty{DYN})};

    \node[anchor=center, align=center, shift=({0mm,1cm})] at (RU-from-MELT.center) {Melt runoff\\(\sankeygetnodeqty{RU-from-MELT})};

    \node[anchor=center, align=center, shift=({0mm,-1cm})] at (GROSS-to-MELT.center) {Melting \\(\sankeygetnodeqty{GROSS-to-MELT})};

    \node[anchor=center, align=center, shift=({-2.5cm,-1.5cm})] at (RFZ-from-MELT.center) {Refreezing\\loop (\sankeygetnodeqty{RFZ-from-MELT})};

    \node[anchor=center, align=center, shift=({0cm,-1cm})] at (SU-out.center) {Sublimation\\(\sankeygetnodeqty{SU-out})};

    \node[anchor=center, align=center, shift=({0cm,2.25cm})] at (ML.center) {Mass loss (\sankeygetnodeqty{ML}) = \\SMB \& Dynamics (\sankeygetnodeqty{IL})\\+ Frontal retreat (\sankeygetnodeqty{FRET})};

    \node[anchor=center, align=center, shift=({0cm,0.75cm})] at (SF.center) {Snowfall  (\sankeygetnodeqty{SF})};

    \node[anchor=east, align=center, rotate=-90, shift=({-3mm,0cm})] at (DP.center) {Deposition (\sankeygetnodeqty{DP})};

    \node[anchor=east, align=center, rotate=-90, shift=({-3mm,0cm})] at (CD.center) {Condensation (\sankeygetnodeqty{CD})};

    \node[anchor=center, align=center, rotate=-90, shift=({0cm,-1cm})] at (RF-NET.center) {Net rainfall (\sankeygetnodeqty{RF-NET})};

  \end{sankeydiagram}
\end{tikzpicture}
\end{document}

### 2024-04-22 tex/gl_sankey.tex
Org -> Make

--- 2024-04-22 tex/aq_sankey.tex
@@
       %% fill/.style={fill=yellow!10,line width=0pt,draw=yellow!10},
-      ratio=1mm/50,
+      ratio=1mm/15,
       minimum radius=1mm,
@@
         inner sep=.1em,
-        fill=white,fill opacity=1,
+        fill=white,fill opacity=0.5,
         rounded corners=.1em,
@@
         \Q{CD}/GROSS-from-CD,   % condensation
-        \Q{RF}/GROSS-from-RF,   % rainfall
-        \Q{RFZ}/GROSS-from-RFZ  % refrozen melt
+        \Q{RFZ}+\Q{RF}/GROSS-from-RFZ  % refrozen melt
       },
@@
       \Q{SU}/GROSS-to-SU,
+      \Q{BM}/GROSS-to-BM,
       \Q{DYN}/GROSS-to-DYN,
-      \Q{BM}/GROSS-to-BM,
       \Q{MELT}/GROSS-to-MELT}
 
-    \sankeynode{name=BM, quantity=\Q{BM}, at={[xshift=-3mm, yshift=-3cm]GROSS-to-BM.center}, anchor=center}
-    \sankeynode{name=MELT, quantity=\Q{MELT}, at={[xshift=-4mm, yshift=0mm]BM.center}, anchor=left}
-    \sankeynode{name=DYN, quantity=\Q{DYN}, at={[xshift=3mm, yshift=0mm]BM.center}, anchor=right}
-    \sankeynode{name=DYN-FRET, quantity=\Q{DYN}+\Q{FRET}, at={[xshift=3mm, yshift=-3cm]DYN.center}, anchor=center}
+    \sankeynode{name=BM, quantity=\Q{BM}, at={[xshift=8mm, yshift=-3cm]GROSS-to-BM.center}, anchor=center}
+    \sankeynode{name=DYN, quantity=\Q{DYN}, at={[xshift=-8mm, yshift=0mm]BM.right}, anchor=left}
+    \sankeynode{name=DYN-FRET, quantity=\Q{DYN}+\Q{FRET}, at={[xshift=0mm, yshift=-3cm]DYN.center}, anchor=center}
+    \sankeynode{DYN-FRET-attach-x, quantity=\Q{FRET}, at={[yshift=0.5cm]DYN-FRET}, anchor=center, forked={\Q{FRET}/DYN-FRET-attach}}
+    \sankeynode{name=MELT, quantity=\Q{MELT}, at={[xshift=-4mm, yshift=0mm]DYN.right}, anchor=left}
 
@@
     %% RFZ
-    \sankeynode{ name=RFZ-waypoint0, quantity=\Q{RFZ}, angle=90,
-      at={[xshift=-20mm, yshift=0cm]RFZ-from-MELT.right}, anchor=right}
-    \sankeynode{ name=RFZ-waypoint1, quantity=\Q{RFZ}, angle=90,
+    \sankeynode{ name=RFZ-waypoint0, quantity=\Q{RFZ}+\Q{RF}, angle=90,
+      at={[xshift=-20mm, yshift=0cm]RFZ-from-MELT.right}, anchor=right,
+      forked={\Q{RF}/RF-to-RFZ2, \Q{RFZ}/RFZ-waypoint0-in}}
+    \sankeynode{ name=RFZ-waypoint1, quantity=\Q{RFZ}+\Q{RF}, angle=90,
       at={[xshift=-20mm, yshift=0cm]GROSS-from-RFZ.right}, anchor=right}
-    \sankeydubins[water]{RFZ-from-MELT}{RFZ-waypoint0}
+    \sankeydubins[water]{RFZ-from-MELT}{RFZ-waypoint0-in}
     \sankeydubins[ice-to-water]{RFZ-waypoint0}{RFZ-waypoint1}
@@
     \sankeynode{name=FRET-DYN, quantity=\Q{FRET}, at={DYN-FRET.center}, anchor=center}
-    \sankeyadvance{DYN-FRET}{5mm}
     \sankeyfork{DYN-FRET}{\Q{ICE-FRET2}/ICE-from-DYN, \Q{SUB-FRET2}/SUB}
-    \sankeyadvance[ice-to-water]{SUB}{2.5cm}
+    \sankeyadvance[ice-to-water]{SUB}{3cm}
     \sankeyend[water]{SUB}
@@
     \sankeystart[water]{RF-NET}
-    \sankeyadvance[water]{RF-NET}{0.5cm}
+    %% \sankeyadvance[water]{RF-NET}{0.5cm}
 
@@
     \sankeydubins{CD}{GROSS-from-CD}
-    \sankeyfork{RF-NET}{\Q{RF}/RFZ-from-RF-NET, \Q{RF-LIQ}/RF-LIQ-from-RF-NET}
-    \sankeydubins[water-to-ice]{RFZ-from-RF-NET}{GROSS-from-RF}
-    \sankeynode{name=RF-wpt1, quantity=\Q{RF-LIQ}, at={[xshift=-10mm]GROSS-to-MELT.right}, align=right}
-    \sankeynode{name=RF-wpt2, quantity=\Q{RF-LIQ}, at={[xshift=-10mm]RFZ-from-MELT.right}, align=right}
-    \sankeydubins[water]{RF-LIQ-from-RF-NET}{RF-wpt1}
-    \sankeydubins[water]{RF-wpt1}{RF-wpt2}
-    \sankeydubins[water]{RF-wpt2}{RF-LIQ-to-RF-NET}
+
+    %% FRZ loop with RF
+    \sankeynode{name=RF-NET1, quantity=\Q{RF-LIQ}+\Q{RF}, at={[xshift=-10mm]GROSS-to-MELT.right}, align=right}
+    \sankeynode{name=RF-NET2, quantity=\Q{RF-LIQ}+\Q{RF},
+      at={[xshift=-10mm, yshift=-1cm]RFZ-from-MELT.right}, align=right}
+    \sankeydubins[water]{RF-NET}{RF-NET1}
+    \sankeydubins[water]{RF-NET1}{RF-NET2}
+    \sankeyfork{RF-NET2}{\Q{RF-LIQ}/RF-to-RU-NET, \Q{RF}/RF-to-RFZ}
+    \sankeydubins[water]{RF-to-RU-NET}{RF-LIQ-to-RF-NET}
+    \sankeydubins[water, minimum radius=6mm]{RF-to-RFZ}{RF-to-RFZ2}
+    %% %% \sankeyfork{RF-wpt2}{\Q{RF}/RFZ-from-RF-NET, \Q{RF-LIQ}/RF-LIQ-from-RF-NET}
+    %%   \sankeydubins[water]{RF-to-RU}{RF-LIQ-to-RF-NET}
 
@@
     \sankeyadvance{FRET}{3cm}
-    \sankeydubins{FRET}{FRET-DYN}
+    %% \sankeydubins{FRET}{FRET-DYN}
+    \sankeydubins{FRET}{DYN-FRET-attach}
 
@@
     \sankeyend{DRIFT-IN}
-    \node[anchor=center, align=center, rotate=0, shift=({0,0})] at (0,-15mm) {Drifting (?)};
+    \node[anchor=center, align=center, rotate=0, shift=({0,0})] at (0,-15mm) {Drifting (\Q{DR})};
 
@@
 
-    \node[anchor=center, align=center, shift=({0mm,1cm})] at (RU-from-MELT.center) {Melt runoff\\(\sankeygetnodeqty{RU-from-MELT})};
+    %% \node[anchor=center, align=center, shift=({0mm,1cm})] at (RU-from-MELT.center) {Melt runoff\\(\sankeygetnodeqty{RU-from-MELT})};
 
@@
 
-    \node[anchor=center, align=center, shift=({-2.5cm,-1.5cm})] at (RFZ-from-MELT.center) {Refreezing\\loop (\sankeygetnodeqty{RFZ-from-MELT})};
+    \node[anchor=center, align=center, shift=({0.5cm,1.5cm})] at (RFZ-waypoint1.center) {Refreezing\\(\sankeygetnodeqty{GROSS-from-RFZ})};
 
@@
 
-    \node[anchor=center, align=center, rotate=-90, shift=({0cm,-1cm})] at (RF-NET.center) {Net rainfall (\sankeygetnodeqty{RF-NET})};
+    \node[anchor=east, align=center, rotate=-90, shift=({-3mm,0mm})] at (RF-NET.center) {Net rainfall (\sankeygetnodeqty{RF-NET})};
 

### 2024-04-22 tex/aq_sankey.tex
update aq (sans font in gl)

@@
 
+\renewcommand{\familydefault}{\sfdefault}
+
 \begin{document}
@@
   \begin{sankeydiagram}
-    %% \sankeyset{debug=true}
+    % \sankeyset{debug=true}
 
@@
 
-    %% BM
+    %% BM Basal Melting
     \sankeydubins[ice-to-water]{GROSS-to-BM}{BM}
@@
     \sankeyfork{RU-GROSS}{\Q{RU}+\Q{RF-LIQ}-\Q{EV}/RU-NET, \Q{EV}/EVAP}
-    \sankeyadvance[water]{RU-NET}{1.5cm}
+    \sankeyadvance[water]{RU-NET}{1cm}
     \sankeyend[water]{RU-NET}
@@
     \sankeynode{name=FRET-DYN, quantity=\Q{FRET}, at={DYN-FRET.center}, anchor=center}
-    \sankeyadvance{DYN-FRET}{5mm}
+    \sankeyadvance{DYN-FRET}{1cm}
     \sankeyfork{DYN-FRET}{\Q{ICE-FRET2}/ICE-from-DYN, \Q{SUB-FRET2}/SUB}
-    \sankeyadvance[ice-to-water]{SUB}{2.5cm}
+    \sankeyadvance[ice-to-water]{SUB}{1.5cm}
     \sankeyend[water]{SUB}
@@
 
+    %% Snowfall SF
     \sankeynode{name=SF, quantity=\Q{SF}, at={[xshift=0, yshift=3cm]GROSS-from-SF.center}, align=center}
-    \sankeynode{name=DP, quantity=\Q{DP}, at={[xshift=-3mm, yshift=0]SF.right}, align=left}
-    \sankeynode{name=CD, quantity=\Q{CD}, at={[xshift=-3mm, yshift=0]DP.right}, align=left}
-    \sankeynode{name=RF-NET, quantity=\Q{RF}+\Q{RF-LIQ}, at={[xshift=-6mm, yshift=0]CD.right}, align=left}
+
+    %% Deposition DP
+    \sankeynode{name=DP, quantity=\Q{DP}, at={[xshift=-2mm, yshift=10mm]SF.right}, align=left}
+
+    %% Condensation CD
+    \sankeynode{name=CD, quantity=\Q{CD}, at={[xshift=-3mm, yshift=4mm]SF.right}, align=left}
+
+    %% Net Rainfall RF-NET
+    \sankeynode{name=RF-NET, quantity=\Q{RF}+\Q{RF-LIQ}, at={[xshift=-6mm, yshift=0mm]SF.right}, align=left}
     \sankeystart[water]{RF-NET}
@@
     \sankeydubins[water-to-ice]{RFZ-from-RF-NET}{GROSS-from-RF}
-    \sankeynode{name=RF-wpt1, quantity=\Q{RF-LIQ}, at={[xshift=-10mm]GROSS-to-MELT.right}, align=right}
-    \sankeynode{name=RF-wpt2, quantity=\Q{RF-LIQ}, at={[xshift=-10mm]RFZ-from-MELT.right}, align=right}
-    \sankeydubins[water]{RF-LIQ-from-RF-NET}{RF-wpt1}
+    \sankeynode{name=RF-wpt0, quantity=\Q{RF-LIQ}, at={[xshift=-16mm,yshift=-10mm]GROSS-from-RFZ.right}, align=right}
+    \sankeynode{name=RF-wpt1, quantity=\Q{RF-LIQ}, at={[xshift=-18mm]GROSS-to-MELT.right}, align=right}
+    \sankeynode{name=RF-wpt2, quantity=\Q{RF-LIQ}, at={[xshift=-18mm,yshift=-10mm]RFZ-from-MELT.right}, align=right}
+    \sankeydubins[water]{RF-LIQ-from-RF-NET}{RF-wpt0}
+    \sankeydubins[water]{RF-wpt0}{RF-wpt1}
     \sankeydubins[water]{RF-wpt1}{RF-wpt2}
@@
 
-    \node[anchor=center, align=center, shift=({0mm,-1cm})] at (RU-NET.center) {Net Runoff\\(\sankeygetnodeqty{RU-NET})};
+    \node[anchor=east, align=center, shift=({0mm,-1cm})] at (RU-NET.center) {Net Runoff\\(\sankeygetnodeqty{RU-NET})};
 
-    \node[anchor=center, align=center, shift=({0mm,-1cm})] at (SUB.center) {Submarine\\melt (\sankeygetnodeqty{SUB})};
+    \node[anchor=center, align=center, shift=({0mm,-1cm})] at (SUB.center) {Submarine\\melt \\(\sankeygetnodeqty{SUB})};
 
@@
 
-    \node[anchor=center, align=left, rotate=-90, shift=({1.5cm, -2mm})] at (BM.center) {Basal\\melting (\sankeygetnodeqty{BM})};
+    \node[anchor=north, align=center, rotate=0, shift=({0cm, -0.5cm})] at (BM.center) {Basal\\melting \\(\sankeygetnodeqty{BM})};
 
@@
 
-    \node[anchor=center, align=center, shift=({0mm,1cm})] at (RU-from-MELT.center) {Melt runoff\\(\sankeygetnodeqty{RU-from-MELT})};
+    \node[anchor=center, align=center, shift=({-0.6cm,2cm})] at (RU-from-MELT.center) {Melt\\ runoff\\(\sankeygetnodeqty{RU-from-MELT})};
 
-    \node[anchor=center, align=center, shift=({0mm,-1cm})] at (GROSS-to-MELT.center) {Melting \\(\sankeygetnodeqty{GROSS-to-MELT})};
+    \node[anchor=center, align=right, shift=({-1cm,-1cm})] at (GROSS-to-MELT.center) {Melting \\(\sankeygetnodeqty{GROSS-to-MELT})};
 
-    \node[anchor=center, align=center, shift=({-2.5cm,-1.5cm})] at (RFZ-from-MELT.center) {Refreezing\\loop (\sankeygetnodeqty{RFZ-from-MELT})};
+    \node[anchor=right, align=right, shift=({-2cm,1cm})] at (RFZ-from-MELT.center) {Refreezing\\loop \\(\sankeygetnodeqty{RFZ-from-MELT})};
 
-    \node[anchor=center, align=center, shift=({0cm,-1cm})] at (SU-out.center) {Sublimation\\(\sankeygetnodeqty{SU-out})};
+    \node[anchor=center, align=center, shift=({-0.2cm,-1cm})] at (SU-out.center) {Sublimation\\(\sankeygetnodeqty{SU-out})};
 
@@
 
-    \node[anchor=east, align=center, rotate=-90, shift=({-3mm,0cm})] at (DP.center) {Deposition (\sankeygetnodeqty{DP})};
+    \node[anchor=east, align=center, rotate=0, shift=({-0mm,0.2cm})] at (DP.center) {Deposition (\sankeygetnodeqty{DP})};
 
-    \node[anchor=east, align=center, rotate=-90, shift=({-3mm,0cm})] at (CD.center) {Condensation (\sankeygetnodeqty{CD})};
+    \node[anchor=east, align=center, rotate=0, shift=({-0mm,0.2cm})] at (CD.center) {Condensation (\sankeygetnodeqty{CD})};
 
-    \node[anchor=center, align=center, rotate=-90, shift=({0cm,-1cm})] at (RF-NET.center) {Net rainfall (\sankeygetnodeqty{RF-NET})};
+    \node[anchor=east, align=center, rotate=0, shift=({-0mm,0.3cm})] at (RF-NET.center) {Net rainfall (\sankeygetnodeqty{RF-NET})};
 

### 2024-05-16 tex/gl_sankey.tex
update aq (sans font in gl)
Update w/ MAR values and Antarctic (all/E/W)
Adjust sublimation label position

@@
 \pagestyle{empty}
+
+\renewcommand{\familydefault}{\sfdefault}
 
@@
 
-    \sankeynode{name=SF, quantity=\Q{SF}, at={[xshift=0, yshift=3cm]GROSS-from-SF.center}, align=center}
+    \sankeynode{name=SF, quantity=\Q{SF}, at={[xshift=0, yshift=5cm]GROSS-from-SF.center}, align=center}
     \sankeynode{name=DP, quantity=\Q{DP}, at={[xshift=-3mm, yshift=0]SF.right}, align=left}
@@
     \sankeystart[red]{ML}
+    %% \sankeynode{ name=ML_start, quantity=\Q{FRET}+\Q{IL}, at={[xshift=8mm]SF.left}, 
+      anchor=left, angle=90}
+    %% \sankeyend[red]{ML_start}
     \sankeyadvance[red-to-ice]{ML}{1cm}
@@
 
-    \node[anchor=center, align=center, shift=({0cm,-1cm})] at (SU-out.center) {Sublimation\\(\sankeygetnodeqty{SU-out})};
+    \node[anchor=center, align=center, shift=({0cm,1cm})] at (SU-out.center) {Sublimation\\(\sankeygetnodeqty{SU-out})};
 

### 2024-05-16 tex/aq_sankey.tex
Adjust sublimation label position
ML (mass loss) -> MG (mass gain)

@@
     \sankeyset{ice}
-    \sankeynode{ name=ML, quantity=\Q{FRET}+\Q{IL}, at={[xshift=8mm]SF.left}, anchor=right}
-    \sankeystart[red]{ML}
-    \sankeyadvance[red-to-ice]{ML}{1cm}
-    \sankeyfork{ML}{\Q{FRET}/FRET, \Q{IL}/IL}
+    \sankeynode{ name=MG, quantity=\Q{FRET}+\Q{IL}, at={[xshift=8mm]SF.left}, anchor=right}
+    \sankeystart[red]{MG}
+    \sankeyadvance[red-to-ice]{MG}{1cm}
+    \sankeyfork{MG}{\Q{FRET}/FRET, \Q{IL}/IL}
     \sankeydubins[minimum radius=2mm]{IL}{GROSS-from-IL}
@@
 
-    \node[anchor=center, align=center, shift=({-0.2cm,-1cm})] at (SU-out.center) {Sublimation\\(\sankeygetnodeqty{SU-out})};
+    \node[anchor=center, align=center, shift=({0.2cm,1cm})] at (SU-out.center) {Sublimation\\(\sankeygetnodeqty{SU-out})};
 
-    \node[anchor=center, align=center, shift=({0cm,2.25cm})] at (ML.center) {Mass loss (\sankeygetnodeqty{ML}) = \\SMB \& Dynamics (\sankeygetnodeqty{IL})\\+ Frontal retreat (\sankeygetnodeqty{FRET})};
+    \node[anchor=center, align=center, shift=({0cm,2.25cm})] at (MG.center) {Mass loss (\sankeygetnodeqty{MG}) = \\SMB \& Dynamics (\sankeygetnodeqty{IL})\\+ Frontal retreat (\sankeygetnodeqty{FRET})};
 

### 2024-05-18 tex/aq_sankey.tex
AQ: grounding line retreat and frontal retreat

@@
   \begin{sankeydiagram}
-    % \sankeyset{debug=true}
+    %% \sankeyset{debug=true}
 
@@
         {MELT/Surface melt/\Q{RU}+\Q{RFZ}},
-        {ICE-FRET2/ICE and some FRET/\Q{ICE}+0.5*\Q{FRET}},
-        {SUB-FRET2/SUB and some FRET/\Q{SUB}+0.5*\Q{FRET}},
+        {ICE-GZRET2/ICE and some GZRET/\Q{ICE}+0.5*\Q{GZRET}},
+        {SUB-GZRET2/SUB and some GZRET/\Q{SUB}+0.5*\Q{GZRET}},
         {GROSS/Gross\\accumulation/\Q{DP}+\Q{RFZ}+\Q{SF}+\Q{RF}+\Q{CD}},
-        {IL/Ice Loss/
-          (\Q{SU}+\Q{RU}+\Q{BM}+\Q{DYN}-(\Q{SF}+\Q{RF}+\Q{CD}+\Q{DP}))},
-        {ML/Net mass Loss/\Q{FRET}+\Q{IL}}
+        {MG/Mass gain/
+          (\Q{SF}+\Q{RF}+\Q{CD}+\Q{DP})-(\Q{SU}+\Q{RU}+\Q{BM}+\Q{DYN})},
+        {MGNet/Net mass gain/\Q{MG}-\Q{FRET}-\Q{GZRET}}
       },
@@
       name=GROSS,
-      quantity=\Q{GROSS}+\Q{IL},
+      quantity=\Q{GROSS}, %+\Q{IL},
       at={0,0},
@@
       forked={
-        \Q{IL}/GROSS-from-IL,  % Ice loss
         \Q{SF}/GROSS-from-SF,   % snowfall
@@
     \sankeyfork{GROSS}{
+      \Q{MGNet}/MGNet-from-GROSS,
       \Q{SU}/GROSS-to-SU,
+      \Q{FRET}/FRET-from-GROSS,  
+      \Q{GZRET}/GZRET-from-GROSS,
       \Q{DYN}/GROSS-to-DYN,
@@
     \sankeynode{name=DYN, quantity=\Q{DYN}, at={[xshift=3mm, yshift=0mm]BM.center}, anchor=right}
-    \sankeynode{name=DYN-FRET, quantity=\Q{DYN}+\Q{FRET}, at={[xshift=3mm, yshift=-3cm]DYN.center}, anchor=center}
+    \sankeynode{name=DYN-GZRET, quantity=\Q{DYN}+\Q{GZRET}, at={[xshift=3mm, yshift=-3cm]DYN.center}, anchor=center}
 
@@
     \sankeydubins{GROSS-to-DYN}{DYN}
-    \sankeyoutin{DYN}{DYN-FRET}
-    \sankeynode{name=FRET-DYN, quantity=\Q{FRET}, at={DYN-FRET.center}, anchor=center}
-    \sankeyadvance{DYN-FRET}{1cm}
-    \sankeyfork{DYN-FRET}{\Q{ICE-FRET2}/ICE-from-DYN, \Q{SUB-FRET2}/SUB}
+    \sankeyoutin{DYN}{DYN-GZRET}
+    \sankeynode{name=GZRET-DYN, quantity=\Q{GZRET}, at={DYN-GZRET.center}, anchor=center}
+    \sankeyadvance{DYN-GZRET}{1cm}
+    \sankeyfork{DYN-GZRET}{\Q{ICE-GZRET2}/ICE-from-DYN, \Q{SUB-GZRET2}/SUB}
     \sankeyadvance[ice-to-water]{SUB}{1.5cm}
     \sankeyend[water]{SUB}
-    \sankeynode{name=ICE, quantity=\Q{ICE-FRET2}, at={[xshift=3mm, yshift=0mm]SUB.left}, anchor=right}
+    \sankeynode{name=ICE, quantity=\Q{ICE-GZRET2}, at={[xshift=3mm, yshift=0mm]SUB.left}, anchor=right}
     \sankeydubins{ICE-from-DYN}{ICE}
-    \sankeyend{ICE}
+    \sankeynode{name=ICE-FRET, quantity=\Q{ICE-GZRET2}+\Q{FRET}, at=ICE.right, anchor=right, forked={\Q{FRET}/FRET, \Q{ICE-GZRET2}/ICE-GZRET3}}
+    \sankeyend{ICE-FRET}
+
+    \sankeydubins[minimum radius=1cm]{FRET-from-GROSS}{FRET}
+
+    \sankeyadvance{GZRET-from-GROSS}{1cm}
+    \sankeydubins[minimum radius=1cm]{GZRET-from-GROSS}{GZRET-DYN}
 
@@
     %% SU
+    \sankeynode{name=SU-out-tmp, quantity=\Q{SU},
+      at={[xshift=5mm, yshift=-35mm]GROSS-to-SU.center}, anchor=center}
     \sankeynode{name=SU-out, quantity=\Q{SU},
-      at={[xshift=10mm, yshift=-10mm]GROSS-to-SU.center}, anchor=center, angle=0}
-    \sankeydubins[ice-to-air]{GROSS-to-SU}{SU-out}
+      at={[xshift=5mm, yshift=-5mm]SU-out-tmp.center}, anchor=center, angle=0}
+    \sankeydubins[ice-to-air]{GROSS-to-SU}{SU-out-tmp}
+    \sankeydubins[air,minimum radius=2mm]{SU-out-tmp}{SU-out}
     \sankeyend[air]{SU-out}
+
+    \sankeynode{name=MGNet-out-tmp, quantity=\Q{MGNet},
+      at={[xshift=5mm, yshift=-15mm]GROSS-to-SU.center}, anchor=center}
+    \sankeynode{name=MGNet-out, quantity=\Q{MGNet},
+      at={[xshift=5mm, yshift=-5mm]MGNet-out-tmp.center}, anchor=center, angle=0}
+    \sankeydubins[ice]{MGNet-from-GROSS}{MGNet-out-tmp}
+    \sankeydubins[ice,minimum radius=2mm]{MGNet-out-tmp}{MGNet-out}
+    \sankeyend[ice]{MGNet-out}
 
@@
     \sankeyset{ice}
-    \sankeynode{ name=MG, quantity=\Q{FRET}+\Q{IL}, at={[xshift=8mm]SF.left}, anchor=right}
-    \sankeystart[red]{MG}
-    \sankeyadvance[red-to-ice]{MG}{1cm}
-    \sankeyfork{MG}{\Q{FRET}/FRET, \Q{IL}/IL}
-    \sankeydubins[minimum radius=2mm]{IL}{GROSS-from-IL}
-    \sankeyadvance{FRET}{2cm}
-    \sankeyturnleft{FRET}{30}
-    \sankeyadvance{FRET}{2cm}
-    \sankeyturnleft{FRET}{-30}
-    \sankeyadvance{FRET}{3cm}
-    \sankeydubins{FRET}{FRET-DYN}
+    %% \sankeynode{ name=MG, quantity=\Q{MG}, at={[xshift=20mm, yshift=-25mm]GROSS-from-IL.left}, anchor=right, angle=90}
+    %% \sankeystart[red]{MG}
+    %% %% \sankeyadvance[red-to-ice]{MG}{1cm}
+    %% \sankeyfork{MG}{\Q{GZRET}/GZRET, \Q{IL}/IL}
+    %% \sankeydubins[minimum radius=0mm]{IL}{GROSS-from-IL}
+    %% %% \sankeyadvance{GZRET}{2cm}
+    %% %% \sankeyturnleft{GZRET}{30}
+    %% %% \sankeyadvance{GZRET}{2cm}
+    %% %% \sankeyturnleft{GZRET}{-30}
+    %% %% \sankeyadvance{GZRET}{3cm}
+    %% \sankeydubins{GZRET}{GZRET-DYN}
 
@@
 
-    \node[anchor=center, align=center, shift=({0cm,2.25cm})] at (MG.center) {Mass loss (\sankeygetnodeqty{MG}) = \\SMB \& Dynamics (\sankeygetnodeqty{IL})\\+ Frontal retreat (\sankeygetnodeqty{FRET})};
+    \node[anchor=center, align=center, shift=({0.2cm,1cm})] at (MGNet-out.center) {Net mass gain\\(\sankeygetnodeqty{MGNet-out})};
+
+    \node[anchor=north, align=center, rotate=0, shift=({0cm, 0.0cm})] at (FRET.center) {Frontal retreat \\(\sankeygetnodeqty{FRET})};
+
+    \node[anchor=north, align=center, rotate=0, shift=({0cm, 0.0cm})] at (GZRET-DYN.center) {Grounding line retreat \\(\sankeygetnodeqty{GZRET-DYN})};
 

### 2024-05-24 ms.tex
ms template

%% Use the "review" option when submitting for review.
%% Removing the "review" option will switch the
%% manuscript to a two-column layout.
%%
%% Use option "jog" for submissions to the Journal of
%% Glaciology; use "aog" for submission to the Annals
%% of Glaciology.
\documentclass[review,jog]{igs}
% \documentclass[review,aog]{igs}

\usepackage[utf8]{inputenc}

\usepackage{mathabx}
\usepackage{graphicx}
\usepackage{siunitx}

% the default is for unnumbered section heads
% if you really must have numbered sections, remove
% the % from the beginning of the following command
% and insert the level of sections you wish to be
% numbered (up to 4):

% \setcounter{secnumdepth}{2}

\jourvolume{V}
\jourissue{N}
\jourpubyear{YYYY}

\begin{document}

\title[Sankey mass flows]{Ice sheet mass flows}

\author[Mankoff and others]
       {Kenneth D. MANKOFF,$^{1,2}$
         William COLGAN,$^3$
         Xavier FETTWEIS,$^4$
         Someone, ELSE,$^n$}

\affiliation{%
  $^1$NASA Goddard Institute for Space Studies, New York, NY, 10025 USA\\
  $^2$Autonomic Integra LLC, New York, NY, 10025 USA\\
  $^3$GEUS affiliation\\
  $^4$Xavier affiliation\\
  $^n$Somewhere Else\\
  Correspondence: Ken Mankoff
  \email{ken.mankoff@nasa.gov}}

\begin{frontmatter}
\maketitle
\begin{abstract}
  Mass flows, contributing to mass loss or gain of the ice sheets, is an of-studied topic in the domain of glaciology. However, few studies consider all processes - this is usually the domain of review papers.  Here we...
\end{abstract}
\end{frontmatter}

\section{Overview of mass flows}

\section{Interpretation}

Red indicates drawdown of the ice sheet mass, which here is an input to the mass flow in Greenland, because the outputs are larger than the external inputs (e.g., snowfall).

Black indicates increase of ice sheet mass, which is an output to the mass flow in Antarctica, because the outputs are smaller than the external inputs.

We mark grounding line in Antarctica and frontal retreat in Greenland in red, even if Antarctica has net mass gain, because not all mass is equal when considering ice sheet health. Antarctica is mass positive because of increased snowfalls. Grounding line retreat is a localized mass imbalance term.

\subsection{Missing terms}

Basal accumulation (not ice), avalanche on/off, Subaerial frontal melt/sublimation (rolled into other terms), drifting on/off, gz retreat in Greenland (few ice shelves).

Graphics may not be complete. For example, an earlier version of this graphic by \citet[Fig. 2]{cogley_2011} did not contain frontal nor grounding line retreat. These are two distinct processes when ice shelves exist, but can be treated as synonyms for one process at tidewater glacier margins. These terms were not only not included in \citet{cogley_2011}, but their respective values were highly uncertain, and still are, although recent work by \citet{kochtitzky_2023,greene_2024} have constrained these values in Greenland

\section{Uncertainty}

Note that some uncertainty (e.g. snowfall) is larger than many other terms combined.

Synthesize what each of the uncertainties is a function of (lack of measurements/scale/timing of measurements/lack of process understanding/variability/etc.). Also, rather than singling the uncertainty of each factor the feedbacks between them could be indicated.

\subsection{Drifting snow}

% I think if you were to go into a discussion of snowdrift it should go further than, for example, the works of Lenaerts et al. Perhaps it is beter to plainly list the uncertainties / poor definitions but in terms of process just refer to the existing papers. However, I am in two minds about this.

\section{Mass flow values}

%% \begin{table} % table1, one column
%%   \caption{One-column table captions will extend beyond the rules in two-column format. Do not try to adjust! Table captions do not have full points at the end}
%%   \label{period}
%%   \begin{minipage}{86mm} % you only need this line if you have a table footnote
%%     \begin{tabular}{@{}lcc}\hline

%%       Period\footnote{Please do not use more than one `\&' between columns, and note that if a table includes table footnotes, it must be inside a \texttt{minipage} environment.}%
%%       & Surface elevation change
%%       & Emergence velocity\\ \hline

%%       1975--85   & $-0.50$ & 0.43\\
%%       1986--2002 & $-1.03$ & 0.32\\
%%       Difference & $-0.53$ & \llap{$-$}0.11
%%     \end{tabular}
%%   \end{minipage} % you only need this line if you have a table footnote
%% \end{table}

%% Tables may be typeset in either one- or two-column format. To typeset two-column format, add asterisks\\ (\verb"\begin{table*}...\end{table*}") as shown in Table~\ref{seasonal}. We may change the format in-house if necessary. Please avoid the use of colour or shading. Note that if you choose to refer to tables using labels, \verb"\caption" must precede \verb"\label", as in standard \LaTeX. Vertical rules are not house-style and will be removed. Note the use of the minipage environment in Table~\ref{period} which enables table footnotes to be output. If the table is two-column, use \texttt{\{178mm\}} instead of \texttt{\{86mm\}} on line~6. The source code for Tables~\ref{period} and~\ref{seasonal} is shown immediately below the tables.

%% \begin{table*}% table2, two column
%% \caption{Two-column table. Seasonal and annual SAT trends ($^\circ$C\,decade$^{-1}$) in the Arctic}
%% \label{seasonal}
%% % the following illustrates how to align columns on decimal points using the S column identifier from siunitx
%% \setlength\tabcolsep{2.5pt}% column separation reduced from the default 6pt so the table fits the measure
%% \begin{tabular}{@{}l@{\hspace{20pt}}SSSSS@{\hspace{20pt}}SSSSS}\hline
%% Area                 & \multicolumn{5}{c}{1951--2005} & \multicolumn{5}{c}{1976--2005}\\[5pt]
%%                      & {Dec--Feb}       & {Mar--May}    & {Jun--Aug}  & {Sep--Nov}         & {Annual}
%%                      & {Dec--Feb}       & {Mar--May}    & {Jun--Aug}  & {Sep--Nov}         & {Annual}\\ \hline
%% Atlantic region      & 0.09           & 0.29 & 0.10 & 0.09 & 0.15 & 0.470 & 0.60 & 0.45 & 0.53 & 0.59\\
%% Siberian region      & 0.12           & 0.29 & 0.04 & 0.17 & 0.16 & 0.08 & 0.69 & 0.29 & 0.59 & 0.48\\
%% Pacific region       & 0.45           & 0.46 & 0.25 & 0.26 & 0.35 & 0.712 & 1.08 & 0.27 & 0.66 & 0.52\\
%% Canadian region      & 0.16           & 0.12 & 0.14 & 0.30 & 0.18 & 0.20 & 0.52 & 0.48 & 0.94 & 0.53\\
%% Baffin Bay region    & -0.02 & 0.10 & 0.00 & 0.15 & 0.02 & 0.33 & 0.62 & 0.51 & 0.80 & 0.57\\
%% Arctic 1             & 0.16           & 0.21 & 0.12 & 0.20 & 0.18 & 0.36 & 200.65 & 0.42 & 0.74 & 0.54\\
%% Arctic 2             & 0.22           & 0.29 & 0.14 & 0.14 & 0.19 & 0.38 & 0.60 & 0.40 & 0.51 & 0.45\\
%% Arctic 3             & 0.28           & 0.31 & 0.14 & 0.13 & 0.21 & 0.42 & 40.53 & 0.41 & 0.42 & 0.43\\
%% NH ($\mathrm{land}
%%   + \mathrm{ocean}$) & 0.13           & 0.13 & 0.10 & 0.10 & 0.12 & 0.27 & 0.24 & 0.25 & 0.25 & 0.25\\
%% \hline
%% \end{tabular}
%% \end{table*}

%% \subsection{Figures}

%% Figures may be typeset in either one- or two-column format. One-column format allows up to 86$\,$mm (e.g. Fig.~\ref{tracks}); two-column format up to 178$\,$mm (e.g. Fig.~\ref{filters}). Please do not provide original graphics files in which the figure is a great deal larger or smaller than what you envisage will be the final printed size. To typeset two-column format, add asterisks (\verb"\begin{figure*}...\end{figure*}") as shown in Fig.~\ref{filters}. We may change the format in-house if necessary. Please note that if you choose to refer to figures using labels, \verb"\caption" must precede \verb"\label", as in standard \LaTeX.

%% Please send one file for each figure (in other words do not use subfigures) and use a name that clearly identifies it (e.g. `72A712Fig03.eps').

%% In addition, figures should be eps, ai (illustrator), ps, tif, psd or pdf. Use strong black lines with a width of at least 0.75pt at final printed size (avoid tinting if possible) and SI units in labels. Lettering should ideally be Optima to match the final typeface; Arial or a similar sans serif font for a second choice. Aim to have the final-size lettering at 9pt, if possible. Figures should not be in boxes. The source code for Figs~\ref{tracks} and~\ref{filters} is shown immediately below the figures.

%% \subsection{Equations}

%% We are including some complex equations as examples. Equations should be checked for width by removing the \verb"[review]" option. Note the use of \verb|cases*| in the following equation:
%% \begin{equation}
%% \label{arrayexample}
%% \alpha_{t_2}= %
%%   \begin{cases}
%%     \alpha_{t_1} - a_1 [\ln (T+1)]
%%       \mathrm{e}^{(a_2\sqrt{n})}
%%       & \mbox{$n_\mathrm{d} > 0\enskip$ and
%%       $\enskip T > 0$}\\
%%     \alpha_{t_1} - a_3 \mathrm{e}^{(a_2\sqrt{n})}
%%       & \mbox{$n_\mathrm{d} > 0\enskip$ and
%%       $\enskip T < 0$}\\
%%     \alpha_{t_1} + a_4 P_\mathrm{s}
%%       & \mbox{$n_\mathrm{d} = 0$}
%%   \end{cases}
%% \end{equation}

%% Equations should be aligned on the equals signs where possible. Equations that extend beyond the one-column measure should be turned over before an operator.

%% \begin{equation}
%% \label{eqnarrayexample}
%% \begin{aligned}
%% l_c & =  l_0 \left(\frac{\widebar{R}_m}{R} \right)^2
%%   \psi^{\frac{P}{P_0\cos Z}}\\
%%     & \phantom{=} \quad  \times [\cos\beta\, \cos Z
%%     + \sin\beta\,\sin Z\,\cos(\psi_\mathrm{sun}
%%     - \psi_\mathrm{slope})]
%% \end{aligned}
%% \end{equation}

%% \citep{mankoff_2020_solid}
%% \citet{mankoff_2021}

%% \begin{figure}%fig1, one column
%% \centering{\includegraphics{72A712Fig01.eps}}
%% \caption{One-column figures should be $\leq$86$\,$mm. Good artwork can
%%   make or break a paper. Capitalize the first word of a label and use
%%   round not square brackets for units.}
%% \label{tracks}
%% \end{figure}

%% \paragraph{Citations using natbib commands}
%% Note that the standard natbib style file has been modified to fall into line with IGS style. The modified style file is called igsnatbib.sty (included in this distribution), and works exactly the same as natbib.sty. The default IGS house style is \citep{Yan13}. The following combinations are also available -- refer to the natbib documentation if you require any further explanation:\\*[0.5\baselineskip]
%% \begin{tabular}{@{}l@{\hskip -94pt}l}
%% \citep{Yan13}
%%     & \verb"\citep{Yan13}"\\
%% \citep[see][p.$\,$34]{Yan13}\\
%%     & \verb"\citep[see][p.$\,$34]{Yan13}"\\
%% \citep[e.g.][]{Yan13}
%%     & \verb"\citep[e.g.][]{Yan13}"\\
%% \citep[Section~2.3]{Yan13}\\
%%     & \verb"\citep[Section~2.3]{Yan13}"\\
%% \citep{Yan13, Edwards14}\\
%%     &  \verb"\citep{Yan13, Edwards14}"\\
%% \cite{Yan13, Edwards14}\\
%%     &  \verb"\cite{Yan13, Edwards14}"\\
%% \citealt{Yan13}
%%     & \verb"\citealt{Yan13}"\\
%% \cite{Yan13}
%%     & \verb"\cite{Yan13}"\\
%% \citealp{Yan13}
%%     & \verb"\citealp{Yan13}"\\
%% \citeauthor{Yan13}
%%     & \verb"\citeauthor{Yan13}"\\
%% \citeyearpar{Yan13}
%%     & \verb"\citeyearpar{Yan13}"\\
%% \citeyear{Yan13}
%%     & \verb"\citeyear{Yan13}"
%% \end{tabular}

%% \begin{figure*}%fig2, two column
%% \centering{\includegraphics{72A712Fig02.eps}}
%% \caption{Two-column figures should be $\leq$178$\,$mm. SSA reconstructed components found by 
%%   projecting the SSA filters found using the whole 2000 traces in Fig.~4, on trace number 1, 
%%   ordered by magnitude of variance accounted for in the radar trace.}
%% \label{filters}
%% \vspace\baselineskip\hrule % to separate figure from verbatim
%% \vspace\baselineskip
%% \begin{verbatim}
%% \begin{figure*}%fig2, two column
%% \centering{\includegraphics{72A712Fig02.eps}}
%% \caption{Two-column figures should be $\leq$178$\,$mm. SSA reconstructed components found by 
%%   projecting the SSA filters found using the whole 2000 traces in Fig.~4, on trace number 1, 
%%   ordered by magnitude of variance accounted for in the radar trace.}
%% \label{filters}
%% \end{figure*}
%% \end{verbatim}
%% \vspace\baselineskip\hrule % to separate verbatim from text
%% \end{figure*}

\section{Acknowledgements}

LaTeX software community, Cogley for a reference graphic.

% authors generating their own bbl file would uncomment the following two lines, and comment out/delete the references below:

\bibliography{library}   % reads igsrefs.bib
\bibliographystyle{igs}  % imposes IGS bibliography style on output

%% \appendix
%% \section{Appendix}

%% Start an appendix by \hbox{typing \verb"\appendix\section{Appendix}".} Appendices appear after the references. Equation numbers automatically start again with (\ref{appeqn}).
%% \begin{equation}
%% \label{appeqn}
%%  2\eta\kappa \frac{\partial \skew1\bar u}{\partial t} + \rho_{\mathrm r} g \skew1\bar u + D\kappa^4 \skew1\bar u = \skew3\bar\sigma_{zz}.
%% \end{equation}

%% \section{Handling more than one appendix}
%% Use the following code to achieve heading APPENDIX~A followed by APPENDIX~B and APPENDIX~C, with appropriate equation numbers:

%% \begin{verbatim}

%% \appendix
%% \section{Appendix A}

%% \setcounter{equation}{0}
%% \renewcommand\theequation{B\arabic{equation}}
%% \section{Appendix B}

%% \setcounter{equation}{0}
%% \renewcommand\theequation{C\arabic{equation}}
%% \section{Appendix C}
%% \end{verbatim}

\end{document}

### 2024-05-24 ms.tex
AQ BMB from pattyn_2010
Addresses #14.

@@
 However, I am in two minds about this.
+
+\section{Methods}
+
+Use published values
+
+Approximate where values are not easily accessed.
+
+We do not have access to the basal mass balance for Antarctica in spatial map form, only a total Antarctic-wide value from \citet{pattyn_2010}.
+To estimate basal melt for East and West Antarctica, we scale by the approximate areas of East and West Antarctica, or 75 \% and 25 \% respectively.
 

### 2024-06-02 ms.tex
CRediT

@@
        {Kenneth D. MANKOFF,$^{1,2}$
-         William COLGAN,$^3$
-         Xavier FETTWEIS,$^4$
-         Someone, ELSE,$^n$}
+         Someone, ELSE,$^n$,
+       add your name to CREDIT.CSV in repository}
 
@@
   $^2$Autonomic Integra LLC, New York, NY, 10025 USA\\
-  $^3$GEUS affiliation\\
-  $^4$Xavier affiliation\\
   $^n$Somewhere Else\\

### 2024-06-02 ms.tex
Refactor

@@
 \begin{abstract}
-  Mass flows, contributing to mass loss or gain of the ice sheets, is an of-studied topic in the domain of glaciology.
-However, few studies consider all processes - this is usually the domain of review papers.
-Here we...
+  Mass flows of ice sheets, and subsequent mass gains and losses, is an of-studied topic in the domain of glaciology.
+However, few studies consider all processes and their relative magnitudes and uncertainties - this is usually the domain of review papers.
+Here we present a Sankey diagram that provides an overview of all mass flow terms and processes for both the Greenlandic and Antarctic ice sheets.
+We use this display to highlight a) relative magnitudes among processes within ice sheets and among ice sheets and sectors, and b) the uncertainty associated with each property.
+We also...
+TBD.
+
+  There is no new science here, but presenting existing data in new or clearer ways may still lead to new insights.
 \end{abstract}
@@
 
+\section{Introduction to Sankey diagrams}
+
+Sankey diagrams are graphical representations of flow or movement of any property (e.g., mass, energy, money, etc.).
+An early and famous use was Charles Minard's Map of Napoleon's Russian Campaign of 1812 (c.f., \citet{kraak_2021}) that combines the magnitude of active soldiers overlaid on a geographical map to show attrition during war.
+The method was later refined, popularized, and eventually named after Captain Matthew Henry Phineas Riall Sankey who used it to show, among other things, the energy efficiency of a steam engine.
+
 \section{Overview of mass flows}
 
-\section{Interpretation}
+See \url{https://github.com/mankoff/sankey/issues/16} for discussion of graphical plan. Comment there or here.
 
-Red indicates drawdown of the ice sheet mass, which here is an input to the mass flow in Greenland, because the outputs are larger than the external inputs (e.g., snowfall).
+\begin{figure}
+  \centering
+  \includegraphics[height=.65\linewidth]{fig/gl_baseline.pdf}
+  \includegraphics[height=.65\linewidth]{fig/gl_2019.pdf}
+  \caption{Sankey flow for Greenland, baseline and 2019.
+Widths are proportional.
+Colors represent both phases (gray solid, blue liquid, white gaseous) and net mass change (red loss, black gain).
+TODO: Improved labeling.}
+    \label{fig:gl}
+\end{figure}
 
-Black indicates increase of ice sheet mass, which is an output to the mass flow in Antarctica, because the outputs are smaller than the external inputs.
+\begin{figure}
+  \centering
+  \includegraphics[width=.30\linewidth]{fig/aq_baseline.pdf}
+  \includegraphics[width=.30\linewidth]{fig/aq_baseline.pdf}
+  \includegraphics[width=.30\linewidth]{fig/aq_baseline.pdf}
+  \caption{Antarctica, east, west. NOTE east and west are TBD, this is just all AQ shown 3x.}
+  \label{fig:aq}
+\end{figure}
 
-We mark grounding line in Antarctica and frontal retreat in Greenland in red, even if Antarctica has net mass gain, because not all mass is equal when considering ice sheet health.
+\subsection{Interpretation of graphics}
+
+Sankey diagrams are generally intuitive, but the following section may still be helpful in interpreting the diagrams.
+The widths of any section within any figure is proportional to all other widths.
+Widths are/are-not (TODO, see https://github.com/mankoff/sankey/issues/16) proportional between images.
+The flows can be somewhat tied to physical space or geography, but are not.
+An alternate display by \citet[Fig.
+2]{cogley_2011} shows a similar diagram overlaid an a glacier schematic, but without flow magnitude nor proportionality.
+
+Color here represents both phase and net mass change.
+Colors gray, blue, and white represent solid, liquid, and gaseous phases respectively, while red and black highlight net mass loss or gain.
+The latter may be counter-intuitive - for example to see mass loss as an input at the top (red in Fig.
+\ref{fig:gl}) even though most mass loss terms (runoff, calving, etc.) are at the bottom.
+This is because Sankey diagrams are balanced, outputs are larger than inputs (hence net mass loss), so the mass loss term is an input - drawdown of the historical 'stable' ice mass.
+Conversely, in East Antarctica mass gain is an output at the bottom (black in Fig.
+\ref{fig:aq}b} that balances the diagram, because without it, there are more flows into the system than out of it.
+
+%% We mark grounding line in Antarctica and frontal retreat in Greenland in red, even if Antarctica has net mass gain, because not all mass is equal when considering ice sheet health.
 Antarctica is mass positive because of increased snowfalls.
@@
 
-\subsection{Missing terms}
+\subsection{Missing terms and limitations}
 
-Basal accumulation (not ice), avalanche on/off, Subaerial frontal melt/sublimation (rolled into other terms), drifting on/off, gz retreat in Greenland (few ice shelves).
+See \citet[Fig. 2]{cogley_2011}
 
-Graphics may not be complete.
+Basal freeze-on (e.g., \citet{bell_2014}) is not shown.
+It is currently unknown by Ken if this is refreezing of the basal melting term (e.g.,\citet{karlsson_2021}) and therefore that terms source remains unchanged, but some diverts back to the dynamics flow and the basal melting output is reduced, or if the basal melting term is final basal runoff, in which case the source should be increased by the refreezing amount.
+Nonetheless, we do not know the refreezing value.
+
+Also:
+\begin{itemize}
+\item Avalanche on/off. Maybe matters more for glaciers than ice sheets?
+\item Subaerial frontal melt/sublimation. That is, the vertical face in Greenland above the water line. This is included in other terms.
+\item Snow drifting on/off. Not sure how much snow drifts onto an ice sheet, but snow drift off could be good to include.
+\item Grounding line retreat in Greenland. Different than frontal retreat. Ignored, as there are few ice shelves.
+\end{itemize}
+
+There may be other missing terms.
 For example, an earlier version of this graphic by \citet[Fig.
@@
 These are two distinct processes when ice shelves exist, but can be treated as synonyms for one process at tidewater glacier margins.
-These terms were not only not included in \citet{cogley_2011}, but their respective values were highly uncertain, and still are, although recent work by \citet{kochtitzky_2023,greene_2024} have constrained these values in Greenland
+These terms were not only not included in \citet{cogley_2011}, but their respective values were highly uncertain, and still are, although recent work by \citet{kochtitzky_2023,greene_2024} have constrained these values in Greenland.
 
@@
 
-Synthesize what each of the uncertainties is a function of (lack of measurements/scale/timing of measurements/lack of process understanding/variability/etc.).
+Advice from Hester: Synthesize what each of the uncertainties is a function of (lack of measurements/scale/timing of measurements/lack of process understanding/variability/etc.).
 Also, rather than singling the uncertainty of each factor the feedbacks between them could be indicated.
@@
 
-Use published values
+Use published values. See CSV tables in \texttt{dat/} folder or README.org
 
@@
 
-We do not have access to the basal mass balance for Antarctica in spatial map form, only a total Antarctic-wide value from \citet{pattyn_2010}.
-To estimate basal melt for East and West Antarctica, we scale by the approximate areas of East and West Antarctica, or 75 \% and 25 \% respectively.
+We do not have access to the basal mass balance for Antarctica in spatial map form, only a total Antarctic-wide value from \citet{pattyn_2010} reported at 65 Gt yr^{-1}.
+To estimate basal melt for East and West Antarctica, we scale by the magnitude of geothermal heat flux scaled by area for East, West, and Peninsula which is 70 \%, 25 \%, and 5 \% respectively, compared to scaling by area alone which has 75 \% of the area in East Antarctica.
 
 \section{Mass flow values}
+
+See CSV tables in \texttt{dat/} folder or README.org
 

### 2024-06-02 tex/aq_sankey.tex
Refactor

@@
     \colorlet{red}[rgb]{red!50!orange!75!white}
+    \colorlet{black}[rgb]{black!100!black!100!black!100}
 
@@
       ratio=1mm/50,
-      minimum radius=1mm,
+      minimum radius=0mm,
       water/.style={fill/.style={draw=water,line width=0pt,fill=water,},},
@@
       air/.style={fill/.style={draw=air,line width=0pt,fill=air,},},
+      black/.style={fill/.style={draw=black,line width=0pt,fill=black,},},
       ice-to-water/.style={fill/.style={fill=none, top color=ice, bottom color=water},},
@@
       water-to-ice/.style={fill/.style={fill=none, top color=water, bottom color=ice},},
+      ice-to-black/.style={fill/.style={fill=none, top color=ice, bottom color=black},},
       debug label/.style={
@@
     \sankeyadvance[water]{BM}{3cm}
-    \sankeyadvance[water]{BM}{3cm}
+    \sankeyadvance[water]{BM}{2.5cm}
     \sankeyend[water]{BM}
@@
     \sankeynode{name=MGNet-out-tmp, quantity=\Q{MGNet},
-      at={[xshift=5mm, yshift=-15mm]GROSS-to-SU.center}, anchor=center}
+      at={[xshift=7mm, yshift=-15mm]GROSS-to-SU.center}, anchor=center}
     \sankeynode{name=MGNet-out, quantity=\Q{MGNet},
       at={[xshift=5mm, yshift=-5mm]MGNet-out-tmp.center}, anchor=center, angle=0}
-    \sankeydubins[ice]{MGNet-from-GROSS}{MGNet-out-tmp}
-    \sankeydubins[ice,minimum radius=2mm]{MGNet-out-tmp}{MGNet-out}
-    \sankeyend[ice]{MGNet-out}
+    \sankeydubins[ice-to-black]{MGNet-from-GROSS}{MGNet-out-tmp}
+    \sankeydubins[black,minimum radius=2mm]{MGNet-out-tmp}{MGNet-out}
+    \sankeyend[black]{MGNet-out}
 
@@
 
-    \node[anchor=center, align=center, shift=({0.2cm,1cm})] at (MGNet-out.center) {Net mass gain\\(\sankeygetnodeqty{MGNet-out})};
+    \node[anchor=center, align=center, shift=({1cm,0.1cm})] at (MGNet-out.center) {Net mass gain\\(\sankeygetnodeqty{MGNet-out})};
 

### 2024-06-02 sankey.tex
Refactor

\documentclass[11pt]{article}
\usepackage{sankey}
\usepackage{tikz}
\usepackage{pgf}
\usepackage{ifthen}
\pagestyle{empty}

\renewcommand{\familydefault}{\sfdefault}

\begin{document}
\begin{tikzpicture}[remember picture, overlay, shift={(5.5,-1.5)}]

  \begin{sankeydiagram}
    \sankeyset{debug=true}

    \colorlet{water}[rgb]{cyan!75!blue!50!white}
    \colorlet{ice}[rgb]{gray!20!gray!20!gray!20}
    %% \colorlet{air}[rgb]{gray!1!gray!1!gray!1}
    \colorlet{air}{yellow!50}
    \colorlet{red}[rgb]{red!50!orange!75!white}
    \colorlet{black}[rgb]{black!100!black!100!black!100}

    \sankeyset{
      every node/.style={angle=-90},
      start style=arrow,end style=arrow,
      %% fill/.style={fill=yellow!10,line width=0pt,draw=yellow!10},
      ratio=1mm/33,
      minimum radius=0mm,
      water/.style={fill/.style={draw=water,line width=0pt,fill=water,},},
      ice/.style={fill/.style={draw=ice,line width=0pt,fill=ice,},},
      air/.style={fill/.style={draw=air,line width=0pt,fill=air,},},
      black/.style={fill/.style={draw=black,line width=0pt,fill=black,},},
      ice-to-water/.style={fill/.style={fill=none, top color=ice, bottom color=water},},
      ice-to-water-horiz/.style={fill/.style={fill=none, right color=ice, left color=water},},
      ice-to-air/.style={fill/.style={fill=none, top color=ice, bottom color=air},},
      ice-to-air-horiz/.style={fill/.style={fill=none, left color=ice, right color=air},},
      water-to-air-horiz/.style={fill/.style={fill=none, right color=water, left color=air},},
      water-to-air/.style={fill/.style={fill=none, top color=water, bottom color=air},},
      air-to-ice/.style={fill/.style={fill=none, top color=air, bottom color=ice},},
      red/.style={fill/.style={fill=none, top color=red, bottom color=red},},
      red-to-ice/.style={fill/.style={fill=none, top color=red, bottom color=ice},},
      water-to-ice/.style={fill/.style={fill=none, top color=water, bottom color=ice},},
      ice-to-black/.style={fill/.style={fill=none, left color=ice, right color=black},},
      debug label/.style={
        overlay,
        draw,
        font=\ttfamily,
        text=debug color,text opacity=1,
        inner sep=.1em,
        fill=white,fill opacity=0.5,
        rounded corners=.1em,
        %% node contents={\name (\qty)},
        node contents={\name},
      },
    }

    % storage of labels
    \newcommand\LabSet[2]{% node name, label
      \expandafter\edef\csname#1@Lab\endcsname{#2}}
    \newcommand\Lab[1]{% node name
      \csname#1@Lab\endcsname}
    % storage of quantities
    \newcommand\QSet[2]{% node name, quantity
      \expandafter\edef\csname#1@Q\endcsname{\fpeval{#2}}}
    \newcommand\Q[1]{% node name
      \csname#1@Q\endcsname}
    % all nodes with their name, label and quantity
    \newcommand\qq[1]{\sankeygetnodeqty{#1}}

    \sankeyset{
      def data/.code args={#1/#2/#3}{% node name/label/values
        \LabSet{#1}{#2}
        \QSet{#1}{#3}
        \typeout{#1: \Q{#1} Gt (\Lab{#1})}
      },
      def data/.list={
        %% INSERT_CSV_HERE
        %% COMPUTED 
        {ML/Mass loss/
          (\Q{SU}+\Q{RU}+\Q{BM}+\Q{DYN}+\Q{FRET}+\Q{GZRET}+\Q{EV})-
          (\Q{SF}+\Q{RF}+\Q{CD}+\Q{DP})},
        {MLfd/Mass Loss fixed domain/\Q{ML}-(\Q{FRET}+\Q{GZRET})},
        {MG/Mass gain/
          (\Q{SF}+\Q{RF}+\Q{CD}+\Q{DP})-
          (\Q{SU}+\Q{RU}+\Q{BM}+\Q{DYN}+\Q{FRET}+\Q{GZRET}+\Q{EV})}
      },
    }

    \sankeyset{ice} % default color until otherwise specified
    \newcommand\xoff{2mm}

    \sankeynode{name=SF, quantity=\Q{SF}}
    \sankeynode{name=DP, quantity=\Q{DP}, at={[xshift=-\xoff]SF.right}, anchor=left}
    \sankeynode{name=CD, quantity=\Q{CD}, at={[xshift=-\xoff]DP.right}, anchor=left}
    \sankeynode{name=RF, quantity=\Q{RF}, at={[xshift=-\xoff]CD.right}, anchor=left}

    \sankeystart{SF}
    \sankeystart[air]{DP}
    \sankeystart{CD}
    \sankeystart[water]{RF}

    \sankeynode{name=f0, quantity=\Q{SF}+\Q{DP}+\Q{CD}+\Q{RF},
      at={[yshift=-15mm]SF.left}, anchor=left,
      forked={\Q{SF}/SF-in, \Q{DP}/DP-in, \Q{CD}/CD-in, \Q{RF}/RF-in}}

    \sankeyoutin{SF}{SF-in}
    \sankeydubins[minimum radius=3mm,air-to-ice]{DP}{DP-in}
    \sankeydubins[minimum radius=3mm]{CD}{CD-in}
    \sankeydubins[minimum radius=3mm,water-to-ice]{RF}{RF-in}

    \ifthenelse{\Q{MG} > \Q{ML}}
    {
      \sankeynode{name=f1, quantity=\Q{RFZ}+\qq{f0},
        at={[yshift=-15mm,xshift=-3mm]f0.right}, anchor=right,
        forked={\qq{f0}/f0-in, \Q{RFZ}/RFZ-in}}
    }
    {
      \sankeynode{name=f1, quantity=\Q{RFZ}+\qq{f0}+\Q{ML},
        at={[yshift=-15mm]f0.center}, anchor=center,
        forked={\Q{ML}/ML-in, \qq{f0}/f0-in, \Q{RFZ}/RFZ-in}}
    }

    \sankeyoutin{f0}{f0-in}
    \sankeyadvance{f1}{15mm}

    \ifthenelse{\Q{MG} > \Q{ML}}
    {
      \sankeyfork{f1}{
        \Q{MG}/fMG,
        \Q{SU}/fSU, \Q{FRET}/fFR, \Q{GZRET}/fGZ,
        \Q{DYN}/f6, \Q{BM}/fBM, \Q{RU}+\Q{EV}/f8, \Q{RFZ}/RFZ-out}
    }
    {
      \sankeyfork{f1}{
        %% \Q{MG}/fMG,
        \Q{SU}/fSU, \Q{FRET}/fFR, \Q{GZRET}/fGZ,
        \Q{DYN}/f6, \Q{BM}/fBM, \Q{RU}+\Q{EV}/f8, \Q{RFZ}/RFZ-out}
    }

    \sankeydubins[minimum radius=5mm,ice-to-water-horiz]{RFZ-out}{RFZ-in}

    \sankeyadvance{f6}{10mm}
    \sankeyfork{f6}{\Q{ICE}/fD, \Q{SUB}/fSUB}
    \sankeyadvance{fD}{10mm}

    \sankeynode{name=fSUBa, quantity=\qq{fSUB},
      at={[xshift=-\xoff]fD.right}, anchor=left}
    \sankeyoutin[ice-to-water]{fSUB}{fSUBa}

    \sankeynode{name=fBMa, quantity=\qq{fBM},
      at={[xshift=-\xoff]fSUBa.right}, anchor=left}
    \sankeyoutin[ice-to-water]{fBM}{fBMa}

    \sankeynode{name=f8a, quantity=\qq{f8},
      at={[xshift=-\xoff]fBMa.right}, anchor=left}
    \sankeyoutin[ice-to-water]{f8}{f8a}

    \sankeynode{name=fGZa, quantity=\qq{fGZ},
      at={[xshift=\xoff]fD.left}, anchor=right}
    \sankeyoutin{fGZ}{fGZa}

    \sankeynode{name=fFRa, quantity=\qq{fFR},
      at={[xshift=\xoff]fGZa.left}, anchor=right}
    \sankeyoutin{fFR}{fFRa}

    \sankeynode{name=fSUa, quantity=\qq{fSU},
      at={[xshift=\xoff]fFRa.left}, anchor=right}
    \sankeyoutin{fSU}{fSUa}

    \ifnum \numexpr\Q{MG} > \Q{ML}
      \sankeynode{name=fMGa, quantity=\qq{fMG},
        at={[xshift=\xoff,yshift=8mm]fSUa.left}, anchor=right, angle=0}
      \sankeyoutin{fMG}{fMGa}
    \fi

    \sankeyfork{f8a}{\Q{RU}/fRU, \Q{EV}/fEV}
    \sankeynode{name=fEVa, quantity=\qq{fEV},
      at={[xshift=-3mm,yshift=-5mm]f8a.right}, anchor=right, angle=180}
    \sankeyoutin[water]{fEV}{fEVa}
    \sankeyadvance[water-to-air-horiz]{fEVa}{5mm}
    \sankeyend[air]{fEVa}

    \ifthenelse{\Q{MG} < \Q{ML}}
      {
        %% mass loss
        \sankeynode{name=fML, quantity=\Q{ML}, at={[xshift=\xoff]SF.left}, anchor=right}
        \sankeystart[red]{fML}
        \sankeyadvance[red-to-ice]{fML}{5mm}
        \sankeyfork{fML}{\Q{FRET}/f18, \Q{GZRET}/f17, \Q{MLfd}/f16}
        \sankeyadvance{f16}{10mm}

        \sankeynode{name=f17a, quantity=\qq{f17},
          at={[xshift=\xoff]f16.left}, anchor=right}
        \sankeyoutin{f17}{f17a}

        \sankeynode{name=f18a, quantity=\qq{f18},
          at={[xshift=\xoff]f17a.left}, anchor=right}
        \sankeyoutin{f18}{f18a}

        \sankeynode{name=f19, quantity=\qq{ML-in},
          at={ML-in.center}, anchor=center,
          forked={\qq{f18}/f18b, \qq{f17}/f17b, \qq{f16}/f16b}}
        \sankeyoutin{f16}{f16b}
        \sankeyoutin{f17a}{f17b}
        \sankeyoutin{f18a}{f18b}
      }
      {
        %% mass gain
        \sankeyadvance[ice-to-black]{fMGa}{3mm}
        \sankeyend[black]{fMGa}
      }

      %% \sankeynode{name=fMGa, quantity=\qq{fMG},
      %%   at={[xshift=\xoff,yshift=8mm]fSUa.left}, anchor=right, angle=0}
      %% \sankeyoutin{fMG}{fMGa}
      %% \sankeyadvance[ice-to-air]{fSUa}{10mm}; \sankeyend[air]{fSUa}

      \sankeynode{name=fSUa2, quantity=\qq{fSU},
        at={[xshift=2mm,yshift=-5mm]fSUa.left}, anchor=left, angle=0}
      \sankeyoutin[ice]{fSUa}{fSUa2}
      \sankeyadvance[ice-to-air-horiz]{fSUa2}{2mm}
      \sankeyend[air]{fSUa2}

      %% \sankeyoutin{fMG}{fMGa}

      \sankeyadvance{fFRa}{10mm}; \sankeyend{fFRa}
      \sankeyadvance[ice-to-water]{fGZa}{10mm}; \sankeyend[water]{fGZa}
      \sankeyadvance{fD}{10mm}; \sankeyend{fD}

      \sankeyset{water}
      \sankeyadvance{fBMa}{10mm}; \sankeyend{fBMa}
      \sankeyadvance{fSUBa}{10mm}; \sankeyend{fSUBa}
      \sankeyadvance{fRU}{10mm}; \sankeyend{fRU}

  %%   %% DRIFT
  %%   \sankeynode{name=DRIFT-OUT, quantity=33, at={(3mm, -8mm)}, angle=150}
  %%   \sankeynode{name=DRIFT-IN, quantity=33, at={(-3mm, -8mm)}, angle=210}
  %%   \sankeydubins[minimum radius=5mm]{DRIFT-OUT}{DRIFT-IN}
  %%   \sankeystart{DRIFT-OUT}
  %%   \sankeyend{DRIFT-IN}
  %%   \node[anchor=center, align=center, rotate=0, shift=({0,0})] at (0,-15mm) {Drifting (?)};

  %%   \node[anchor=south east, align=center, shift=({2mm,-3mm})] at (EV-out.right) {Evaporation\\(\Q{EV})};

  %%   \node[anchor=east, align=center, shift=({0mm,-1cm})] at (RU-NET.center) {Net Runoff\\(\sankeygetnodeqty{RU-NET})};

  %%   \node[anchor=center, align=center, shift=({0mm,-1cm})] at (SUB.center) {Submarine\\melt \\(\sankeygetnodeqty{SUB})};

  %%   \node[anchor=center, align=center, shift=({0mm,-1cm})] at (ICE.center) {Calving\\(\sankeygetnodeqty{ICE})};

  %%   \node[anchor=north, align=center, rotate=0, shift=({0cm, -0.5cm})] at (BM.center) {Basal\\melting \\(\sankeygetnodeqty{BM})};

  %%   \node[anchor=center, align=center, shift=({0mm,0cm})] at (DYN.center) {Dynamics\\(\sankeygetnodeqty{DYN})};

  %%   \node[anchor=center, align=center, shift=({-0.6cm,2cm})] at (RU-from-MELT.center) {Melt\\ runoff\\(\sankeygetnodeqty{RU-from-MELT})};

  %%   \node[anchor=center, align=right, shift=({-1cm,-1cm})] at (GROSS-to-MELT.center) {Melting \\(\sankeygetnodeqty{GROSS-to-MELT})};

  %%   \node[anchor=right, align=right, shift=({-2cm,1cm})] at (RFZ-from-MELT.center) {Refreezing\\loop \\(\sankeygetnodeqty{RFZ-from-MELT})};

  %%   \node[anchor=center, align=center, shift=({0.2cm,1cm})] at (SU-out.center) {Sublimation\\(\sankeygetnodeqty{SU-out})};

  %%   \node[anchor=center, align=center, shift=({1cm,0.1cm})] at (MGNet-out.center) {Net mass gain\\(\sankeygetnodeqty{MGNet-out})};

  %%   \node[anchor=north, align=center, rotate=0, shift=({0cm, 0.0cm})] at (FRET.center) {Frontal retreat \\(\sankeygetnodeqty{FRET})};

  %%   \node[anchor=north, align=center, rotate=0, shift=({0cm, 0.0cm})] at (GZRET-DYN.center) {Grounding line retreat \\(\sankeygetnodeqty{GZRET-DYN})};

  %% \node[anchor=center, align=center, shift=({0cm,0.75cm})] at (SF.center) {Snowfall (\sankeygetnodeqty{SF})};

  %%   \node[anchor=east, align=center, rotate=0, shift=({-0mm,0.2cm})] at (DP.center) {Deposition (\sankeygetnodeqty{DP})};

  %%   \node[anchor=east, align=center, rotate=0, shift=({-0mm,0.2cm})] at (CD.center) {Condensation (\sankeygetnodeqty{CD})};

  %%   \node[anchor=east, align=center, rotate=0, shift=({-0mm,0.3cm})] at (RF-NET.center) {Net rainfall (\sankeygetnodeqty{RF-NET})};

  \end{sankeydiagram}
\end{tikzpicture}
\end{document}

### 2024-06-04 sankey.tex
labels (commented)
Comment out labels
Drawdown is not split by components (FRET, GZRET)

@@
         %% mass loss
-        \sankeynode{name=fML, quantity=\Q{ML}, at={[xshift=\xoff]SF.left}, anchor=right}
+        \sankeynode{name=fML, quantity=\Q{ML}, at={[xshift=\xoff,yshift=0mm]SF.left}, anchor=right}%, angle=-145}
         \sankeystart[red]{fML}
         \sankeyadvance[red-to-ice]{fML}{5mm}
-        \sankeyfork{fML}{\Q{FRET}/f18, \Q{GZRET}/f17, \Q{MLfd}/f16}
-        \sankeyadvance{f16}{10mm}
-
-        \sankeynode{name=f17a, quantity=\qq{f17},
-          at={[xshift=\xoff]f16.left}, anchor=right}
-        \sankeyoutin{f17}{f17a}
-
-        \sankeynode{name=f18a, quantity=\qq{f18},
-          at={[xshift=\xoff]f17a.left}, anchor=right}
-        \sankeyoutin{f18}{f18a}
-
-        \sankeynode{name=f19, quantity=\qq{ML-in},
-          at={ML-in.center}, anchor=center,
-          forked={\qq{f18}/f18b, \qq{f17}/f17b, \qq{f16}/f16b}}
-        \sankeyoutin{f16}{f16b}
-        \sankeyoutin{f17a}{f17b}
-        \sankeyoutin{f18a}{f18b}
+        \sankeyoutin{fML}{ML-in}
       }
@@
 
-  %%   \node[anchor=south east, align=center, shift=({2mm,-3mm})] at (EV-out.right) {Evaporation\\(\Q{EV})};
+      %% \foreach \mag in {SF,f0,RFZ-in,f6,fRU,fSUBa,fD,fGZa,fFRa,fSUa2,fEVa} {
+      %%   \node at (\mag.center) {\qq{\mag}};
+      %% }
 
-  %%   \node[anchor=east, align=center, shift=({0mm,-1cm})] at (RU-NET.center) {Net Runoff\\(\sankeygetnodeqty{RU-NET})};
-
-  %%   \node[anchor=center, align=center, shift=({0mm,-1cm})] at (SUB.center) {Submarine\\melt \\(\sankeygetnodeqty{SUB})};
-
-  %%   \node[anchor=center, align=center, shift=({0mm,-1cm})] at (ICE.center) {Calving\\(\sankeygetnodeqty{ICE})};
-
-  %%   \node[anchor=north, align=center, rotate=0, shift=({0cm, -0.5cm})] at (BM.center) {Basal\\melting \\(\sankeygetnodeqty{BM})};
-
-  %%   \node[anchor=center, align=center, shift=({0mm,0cm})] at (DYN.center) {Dynamics\\(\sankeygetnodeqty{DYN})};
-
-  %%   \node[anchor=center, align=center, shift=({-0.6cm,2cm})] at (RU-from-MELT.center) {Melt\\ runoff\\(\sankeygetnodeqty{RU-from-MELT})};
-
-  %%   \node[anchor=center, align=right, shift=({-1cm,-1cm})] at (GROSS-to-MELT.center) {Melting \\(\sankeygetnodeqty{GROSS-to-MELT})};
-
-  %%   \node[anchor=right, align=right, shift=({-2cm,1cm})] at (RFZ-from-MELT.center) {Refreezing\\loop \\(\sankeygetnodeqty{RFZ-from-MELT})};
-
-  %%   \node[anchor=center, align=center, shift=({0.2cm,1cm})] at (SU-out.center) {Sublimation\\(\sankeygetnodeqty{SU-out})};
-
-  %%   \node[anchor=center, align=center, shift=({1cm,0.1cm})] at (MGNet-out.center) {Net mass gain\\(\sankeygetnodeqty{MGNet-out})};
-
-  %%   \node[anchor=north, align=center, rotate=0, shift=({0cm, 0.0cm})] at (FRET.center) {Frontal retreat \\(\sankeygetnodeqty{FRET})};
-
-  %%   \node[anchor=north, align=center, rotate=0, shift=({0cm, 0.0cm})] at (GZRET-DYN.center) {Grounding line retreat \\(\sankeygetnodeqty{GZRET-DYN})};
-
-  %% \node[anchor=center, align=center, shift=({0cm,0.75cm})] at (SF.center) {Snowfall (\sankeygetnodeqty{SF})};
-
-  %%   \node[anchor=east, align=center, rotate=0, shift=({-0mm,0.2cm})] at (DP.center) {Deposition (\sankeygetnodeqty{DP})};
-
-  %%   \node[anchor=east, align=center, rotate=0, shift=({-0mm,0.2cm})] at (CD.center) {Condensation (\sankeygetnodeqty{CD})};
-
-  %%   \node[anchor=east, align=center, rotate=0, shift=({-0mm,0.3cm})] at (RF-NET.center) {Net rainfall (\sankeygetnodeqty{RF-NET})};
+      %% \ifthenelse{\Q{MG} < \Q{ML}} {
+      %%   \node at (fML.center) {\qq{fML}};
+      %% } {
+      %%   \node at (fMGa.center) {\qq{fMGa}};
+      %% }
 

### 2024-06-05 sankey.tex
Remove unused commented code
Red outline for processes that change plan-view boundary

@@
       ice-to-black/.style={fill/.style={fill=none, left color=ice, right color=black},},
+      out-red/.style={draw/.style={draw=red,thick=3},},
       debug label/.style={
@@
 
-      %% \sankeynode{name=fMGa, quantity=\qq{fMG},
-      %%   at={[xshift=\xoff,yshift=8mm]fSUa.left}, anchor=right, angle=0}
-      %% \sankeyoutin{fMG}{fMGa}
-      %% \sankeyadvance[ice-to-air]{fSUa}{10mm}; \sankeyend[air]{fSUa}
-
       \sankeynode{name=fSUa2, quantity=\qq{fSU},
@@
 
-      %% \sankeyoutin{fMG}{fMGa}
-
-      \sankeyadvance{fFRa}{10mm}; \sankeyend{fFRa}
-      \sankeyadvance[ice-to-water]{fGZa}{10mm}; \sankeyend[water]{fGZa}
-      \sankeyadvance{fD}{10mm}; \sankeyend{fD}
+      \sankeyadvance[out-red]{fFRa}{10mm}
+      \sankeyend[out-red]{fFRa}
+      \sankeyadvance[ice-to-water,out-red]{fGZa}{10mm}
+      \sankeyend[water,out-red]{fGZa}
+      \sankeyadvance{fD}{10mm}
+      \sankeyend{fD}
 
@@
 
-      %% \foreach \mag in {SF,f0,RFZ-in,f6,fRU,fSUBa,fD,fGZa,fFRa,fSUa2,fEVa} {
-      %%   \node at (\mag.center) {\qq{\mag}};
-      %% }
-
-      %% \ifthenelse{\Q{MG} < \Q{ML}} {
-      %%   \node at (fML.center) {\qq{fML}};
-      %% } {
-      %%   \node at (fMGa.center) {\qq{fMGa}};
-      %% }
-
   \end{sankeydiagram}

### 2024-06-05 sankey.tex
Move mass loss to top left. Align right edge.

@@
     \ifthenelse{\Q{MG} > \Q{ML}}
-    {
+    { % mass gain
       \sankeynode{name=f1, quantity=\Q{RFZ}+\qq{f0},
-        at={[yshift=-15mm,xshift=-3mm]f0.right}, anchor=right,
+        at={[yshift=-15mm]f0.left}, anchor=left,
         forked={\qq{f0}/f0-in, \Q{RFZ}/RFZ-in}}
     }
-    {
+    { % mass loss
       \sankeynode{name=f1, quantity=\Q{RFZ}+\qq{f0}+\Q{ML},
-        at={[yshift=-15mm]f0.center}, anchor=center,
-        forked={\Q{ML}/ML-in, \qq{f0}/f0-in, \Q{RFZ}/RFZ-in}}
+        at={[yshift=-15mm]f0.left}, anchor=left,
+        forked={\qq{f0}/f0-in, \Q{ML}/ML-in, \Q{RFZ}/RFZ-in}}
     }
@@
         %% mass loss
-        \sankeynode{name=fML, quantity=\Q{ML}, at={[xshift=\xoff,yshift=0mm]SF.left}, anchor=right}%, angle=-145}
+        \sankeynode{name=fML, quantity=\Q{ML},
+          at={[xshift=-\xoff]RF.right}, anchor=left} %, angle=-45}
         \sankeystart[red]{fML}

### 2024-06-05 ms.tex
Red outline for processes that change plan-view boundary
abstract rewrite

@@
 \begin{abstract}
+
+  Observing, quantifying, and understanding mass loss takes field observations, computer models, and insight through knowledge generation and sharing.
+Tabular data is optimized for computers and useful for humans, but graphical presentations can provide significantly higher information density.
+Here we present several Sankey diagrams depicting mass flow of ice in Greenland and Antarctica, and discuss some properties of the graphics and the cryospheric processes represented by the graphics.
+We focus on insight available through this choice of display, methods for estimating the various cryospheric processes, and uncertainty,
+\end{abstract}
+\end{frontmatter}
+
+\section{Introduction and Background}
+
 Mass flows of ice sheets, and subsequent mass gains and losses, is an of-studied topic in the domain of glaciology.
@@
 
-  There is no new science here, but presenting existing data in new or clearer ways may still lead to new insights.
-\end{abstract}
-\end{frontmatter}
-
-\section{Introduction to Sankey diagrams}
+\subsection{Introduction to Sankey diagrams}
 
@@
 Grounding line retreat is a localized mass imbalance term.
+
+See thick red outlines on frontal retreat and grounding line retreat.
+Can't be in steady state if these processes are active.
+Steady state may not be a natural occurrance.
+Could have these and be quasi-steady (cyclical).
 

### 2024-06-05 sankey.tex
refactor internal variable and node names.

--- 2024-06-05 sankey.tex
@@
         %% COMPUTED 
-        {ML/Mass loss/
-          (\Q{SU}+\Q{RU}+\Q{BM}+\Q{DYN}+\Q{FRET}+\Q{GZRET}+\Q{EV})-
-          (\Q{SF}+\Q{RF}+\Q{CD}+\Q{DP})},
-        {MLfd/Mass Loss fixed domain/\Q{ML}-(\Q{FRET}+\Q{GZRET})},
-        {MG/Mass gain/
-          (\Q{SF}+\Q{RF}+\Q{CD}+\Q{DP})-
-          (\Q{SU}+\Q{RU}+\Q{BM}+\Q{DYN}+\Q{FRET}+\Q{GZRET}+\Q{EV})}
+        {vML/Mass loss/
+          (\Q{vSU}+\Q{vRU}+\Q{vBM}+\Q{vDYN}+\Q{vFRET}+\Q{vGZRET}+\Q{vEV})-
+          (\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP})},
+        {vMLfd/Mass Loss fixed domain/\Q{vML}-(\Q{vFRET}+\Q{vGZRET})},
+        {vMG/Mass gain/
+          (\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP})-
+          (\Q{vSU}+\Q{vRU}+\Q{vBM}+\Q{vDYN}+\Q{vFRET}+\Q{vGZRET}+\Q{vEV})}
       },
@@
 
-    \sankeynode{name=SF, quantity=\Q{SF}}
-    \sankeynode{name=DP, quantity=\Q{DP}, at={[xshift=-\xoff]SF.right}, anchor=left}
-    \sankeynode{name=CD, quantity=\Q{CD}, at={[xshift=-\xoff]DP.right}, anchor=left}
-    \sankeynode{name=RF, quantity=\Q{RF}, at={[xshift=-\xoff]CD.right}, anchor=left}
+    \sankeynode{name=SF, quantity=\Q{vSF}}
+    \sankeynode{name=DP, quantity=\Q{vDP}, at={[xshift=-\xoff]SF.right}, anchor=left}
+    \sankeynode{name=CD, quantity=\Q{vCD}, at={[xshift=-\xoff]DP.right}, anchor=left}
+    \sankeynode{name=RF, quantity=\Q{vRF}, at={[xshift=-\xoff]CD.right}, anchor=left}
 
@@
 
-    \sankeynode{name=f0, quantity=\Q{SF}+\Q{DP}+\Q{CD}+\Q{RF},
+    \sankeynode{name=f0, quantity=\Q{vSF}+\Q{vDP}+\Q{vCD}+\Q{vRF},
       at={[yshift=-15mm]SF.left}, anchor=left,
-      forked={\Q{SF}/SF-in, \Q{DP}/DP-in, \Q{CD}/CD-in, \Q{RF}/RF-in}}
+      forked={\Q{vSF}/SFi, \Q{vDP}/DPi, \Q{vCD}/CDi, \Q{vRF}/RFi}}
 
-    \sankeyoutin{SF}{SF-in}
-    \sankeydubins[minimum radius=3mm,air-to-ice]{DP}{DP-in}
-    \sankeydubins[minimum radius=3mm]{CD}{CD-in}
-    \sankeydubins[minimum radius=3mm,water-to-ice]{RF}{RF-in}
+    \sankeyoutin{SF}{SFi}
+    \sankeydubins[minimum radius=3mm,air-to-ice]{DP}{DPi}
+    \sankeydubins[minimum radius=3mm]{CD}{CDi}
+    \sankeydubins[minimum radius=3mm,water-to-ice]{RF}{RFi}
 
-    \ifthenelse{\Q{MG} > \Q{ML}}
-    {
-      \sankeynode{name=f1, quantity=\Q{RFZ}+\qq{f0},
-        at={[yshift=-15mm,xshift=-3mm]f0.right}, anchor=right,
-        forked={\qq{f0}/f0-in, \Q{RFZ}/RFZ-in}}
+    \ifthenelse{\Q{vMG} > \Q{vML}}
+    { % mass gain
+      \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0},
+        at={[yshift=-15mm]f0.left}, anchor=left,
+        forked={\qq{f0}/f0i, \Q{vRFZ}/RFZi}}
     }
-    {
-      \sankeynode{name=f1, quantity=\Q{RFZ}+\qq{f0}+\Q{ML},
-        at={[yshift=-15mm]f0.center}, anchor=center,
-        forked={\Q{ML}/ML-in, \qq{f0}/f0-in, \Q{RFZ}/RFZ-in}}
+    { % mass loss
+      \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vML},
+        at={[yshift=-15mm]f0.left}, anchor=left,
+        forked={\qq{f0}/f0i, \Q{vML}/MLi, \Q{vRFZ}/RFZi}}
     }
 
-    \sankeyoutin{f0}{f0-in}
+    \sankeyoutin{f0}{f0i}
     \sankeyadvance{f1}{15mm}
 
-    \ifthenelse{\Q{MG} > \Q{ML}}
+    \ifthenelse{\Q{vMG} > \Q{vML}}
     {
       \sankeyfork{f1}{
-        \Q{MG}/fMG,
-        \Q{SU}/fSU, \Q{FRET}/fFR, \Q{GZRET}/fGZ,
-        \Q{DYN}/f6, \Q{BM}/fBM, \Q{RU}+\Q{EV}/f8, \Q{RFZ}/RFZ-out}
+        \Q{vMG}/MG,
+        \Q{vSU}/SU, \Q{vFRET}/FR, \Q{vGZRET}/GZ,
+        \Q{vDYN}/f6, \Q{vBM}/BM, \Q{vRU}+\Q{vEV}/f8, \Q{vRFZ}/RFZo}
     }
@@
       \sankeyfork{f1}{
-        %% \Q{MG}/fMG,
-        \Q{SU}/fSU, \Q{FRET}/fFR, \Q{GZRET}/fGZ,
-        \Q{DYN}/f6, \Q{BM}/fBM, \Q{RU}+\Q{EV}/f8, \Q{RFZ}/RFZ-out}
+        %% \Q{vMG}/MG,
+        \Q{vSU}/SU, \Q{vFRET}/FR, \Q{vGZRET}/GZ,
+        \Q{vDYN}/f6, \Q{vBM}/BM, \Q{vRU}+\Q{vEV}/f8, \Q{vRFZ}/RFZo}
     }
 
-    \sankeydubins[minimum radius=5mm,ice-to-water-horiz]{RFZ-out}{RFZ-in}
+    \sankeydubins[minimum radius=5mm,ice-to-water-horiz]{RFZo}{RFZi}
 
     \sankeyadvance{f6}{10mm}
-    \sankeyfork{f6}{\Q{ICE}/fD, \Q{SUB}/fSUB}
-    \sankeyadvance{fD}{10mm}
+    \sankeyfork{f6}{\Q{vICE}/D, \Q{vSUB}/SUB}
+    \sankeyadvance{D}{10mm}
 
-    \sankeynode{name=fSUBa, quantity=\qq{fSUB},
-      at={[xshift=-\xoff]fD.right}, anchor=left}
-    \sankeyoutin[ice-to-water]{fSUB}{fSUBa}
+    \sankeynode{name=SUBa, quantity=\qq{SUB},
+      at={[xshift=-\xoff]D.right}, anchor=left}
+    \sankeyoutin[ice-to-water]{SUB}{SUBa}
 
-    \sankeynode{name=fBMa, quantity=\qq{fBM},
-      at={[xshift=-\xoff]fSUBa.right}, anchor=left}
-    \sankeyoutin[ice-to-water]{fBM}{fBMa}
+    \sankeynode{name=BMa, quantity=\qq{BM},
+      at={[xshift=-\xoff]SUBa.right}, anchor=left}
+    \sankeyoutin[ice-to-water]{BM}{BMa}
 
     \sankeynode{name=f8a, quantity=\qq{f8},
-      at={[xshift=-\xoff]fBMa.right}, anchor=left}
+      at={[xshift=-\xoff]BMa.right}, anchor=left}
     \sankeyoutin[ice-to-water]{f8}{f8a}
 
-    \sankeynode{name=fGZa, quantity=\qq{fGZ},
-      at={[xshift=\xoff]fD.left}, anchor=right}
-    \sankeyoutin{fGZ}{fGZa}
+    \sankeynode{name=GZa, quantity=\qq{GZ},
+      at={[xshift=\xoff]D.left}, anchor=right}
+    \sankeyoutin{GZ}{GZa}
 
-    \sankeynode{name=fFRa, quantity=\qq{fFR},
-      at={[xshift=\xoff]fGZa.left}, anchor=right}
-    \sankeyoutin{fFR}{fFRa}
+    \sankeynode{name=FRa, quantity=\qq{FR},
+      at={[xshift=\xoff]GZa.left}, anchor=right}
+    \sankeyoutin{FR}{FRa}
 
-    \sankeynode{name=fSUa, quantity=\qq{fSU},
-      at={[xshift=\xoff]fFRa.left}, anchor=right}
-    \sankeyoutin{fSU}{fSUa}
+    \sankeynode{name=SUa, quantity=\qq{SU},
+      at={[xshift=\xoff]FRa.left}, anchor=right}
+    \sankeyoutin{SU}{SUa}
 
-    \ifnum \numexpr\Q{MG} > \Q{ML}
-      \sankeynode{name=fMGa, quantity=\qq{fMG},
-        at={[xshift=\xoff,yshift=8mm]fSUa.left}, anchor=right, angle=0}
-      \sankeyoutin{fMG}{fMGa}
+    \ifnum \numexpr\Q{vMG} > \Q{vML}
+      \sankeynode{name=MGa, quantity=\qq{MG},
+        at={[xshift=\xoff,yshift=8mm]SUa.left}, anchor=right, angle=0}
+      \sankeyoutin{MG}{MGa}
     \fi
 
-    \sankeyfork{f8a}{\Q{RU}/fRU, \Q{EV}/fEV}
-    \sankeynode{name=fEVa, quantity=\qq{fEV},
+    \sankeyfork{f8a}{\Q{vRU}/RU, \Q{vEV}/EV}
+    \sankeynode{name=EVa, quantity=\qq{EV},
       at={[xshift=-3mm,yshift=-5mm]f8a.right}, anchor=right, angle=180}
-    \sankeyoutin[water]{fEV}{fEVa}
-    \sankeyadvance[water-to-air-horiz]{fEVa}{5mm}
-    \sankeyend[air]{fEVa}
+    \sankeyoutin[water]{EV}{EVa}
+    \sankeyadvance[water-to-air-horiz]{EVa}{5mm}
+    \sankeyend[air]{EVa}
 
-    \ifthenelse{\Q{MG} < \Q{ML}}
+    \ifthenelse{\Q{vMG} < \Q{vML}}
       {
         %% mass loss
-        \sankeynode{name=fML, quantity=\Q{ML}, at={[xshift=\xoff,yshift=0mm]SF.left}, anchor=right}%, angle=-145}
-        \sankeystart[red]{fML}
-        \sankeyadvance[red-to-ice]{fML}{5mm}
-        \sankeyoutin{fML}{ML-in}
+        \sankeynode{name=ML, quantity=\Q{vML},
+          at={[xshift=-\xoff]RF.right}, anchor=left} %, angle=-45}
+        \sankeystart[red]{ML}
+        \sankeyadvance[red-to-ice]{ML}{5mm}
+        \sankeyoutin{ML}{MLi}
       }
@@
         %% mass gain
-        \sankeyadvance[ice-to-black]{fMGa}{3mm}
-        \sankeyend[black]{fMGa}
+        \sankeyadvance[ice-to-black]{MGa}{3mm}
+        \sankeyend[black]{MGa}
       }
 
-      \sankeynode{name=fSUa2, quantity=\qq{fSU},
-        at={[xshift=2mm,yshift=-5mm]fSUa.left}, anchor=left, angle=0}
-      \sankeyoutin[ice]{fSUa}{fSUa2}
-      \sankeyadvance[ice-to-air-horiz]{fSUa2}{2mm}
-      \sankeyend[air]{fSUa2}
+      \sankeynode{name=SUaa, quantity=\qq{SU},
+        at={[xshift=2mm,yshift=-5mm]SUa.left}, anchor=left, angle=0}
+      \sankeyoutin[ice]{SUa}{SUaa}
+      \sankeyadvance[ice-to-air-horiz]{SUaa}{2mm}
+      \sankeyend[air]{SUaa}
 
-      \sankeyadvance[out-red]{fFRa}{10mm}
-      \sankeyend[out-red]{fFRa}
-      \sankeyadvance[ice-to-water,out-red]{fGZa}{10mm}
-      \sankeyend[water,out-red]{fGZa}
-      \sankeyadvance{fD}{10mm}
-      \sankeyend{fD}
+      \sankeyadvance[out-red]{FRa}{10mm}
+      \sankeyend[out-red]{FRa}
+      \sankeyadvance[ice-to-water,out-red]{GZa}{10mm}
+      \sankeyend[water,out-red]{GZa}
+      \sankeyadvance{D}{10mm}
+      \sankeyend{D}
 
       \sankeyset{water}
-      \sankeyadvance{fBMa}{10mm}; \sankeyend{fBMa}
-      \sankeyadvance{fSUBa}{10mm}; \sankeyend{fSUBa}
-      \sankeyadvance{fRU}{10mm}; \sankeyend{fRU}
+      \sankeyadvance{BMa}{10mm}; \sankeyend{BMa}
+      \sankeyadvance{SUBa}{10mm}; \sankeyend{SUBa}
+      \sankeyadvance{RU}{10mm}; \sankeyend{RU}
 
   %%   %% DRIFT
-  %%   \sankeynode{name=DRIFT-OUT, quantity=33, at={(3mm, -8mm)}, angle=150}
-  %%   \sankeynode{name=DRIFT-IN, quantity=33, at={(-3mm, -8mm)}, angle=210}
-  %%   \sankeydubins[minimum radius=5mm]{DRIFT-OUT}{DRIFT-IN}
-  %%   \sankeystart{DRIFT-OUT}
-  %%   \sankeyend{DRIFT-IN}
+  %%   \sankeynode{name=DRIFTo, quantity=33, at={(3mm, -8mm)}, angle=150}
+  %%   \sankeynode{name=DRIFTi, quantity=33, at={(-3mm, -8mm)}, angle=210}
+  %%   \sankeydubins[minimum radius=5mm]{DRIFTo}{DRIFTi}
+  %%   \sankeystart{DRIFTo}
+  %%   \sankeyend{DRIFTI}
   %%   \node[anchor=center, align=center, rotate=0, shift=({0,0})] at (0,-15mm) {Drifting (?)};

### 2024-06-05 ms.tex
shelf text

@@
 Also, rather than singling the uncertainty of each factor the feedbacks between them could be indicated.
+
+\subsection{Shelves}
+
+Mass change of shelves is a bulk aggregate property, and should not the default reporting metric because it obscures information.
+For example, in theory ice shelf mass can grow even as they collapse, as long as the grounding line retreats (adds mass to the shelf from the upstream ice sheet) faster than the mass loss at the frontal or submarine boundaries.
+A mass flow diagram dedicated to ice shelves (this one is not) would clearly convey each of these processes.
 

### 2024-06-07 sankey.tex
vMLfd unused

@@
           (\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP})},
-        {vMLfd/Mass Loss fixed domain/\Q{vML}-(\Q{vFRET}+\Q{vGZRET})},
         {vMG/Mass gain/

### 2024-06-22 sankey.tex
Labels

@@
   \begin{sankeydiagram}
-    \sankeyset{debug=true}
+    %% \sankeyset{debug=true}
 
@@
     \sankeyset{
+      %% comment for gray outline
+      draw/.style={
+        draw=white,
+        line width=0pt,line cap=round,line join=round,
+      },
       every node/.style={angle=-90},
@@
 
+    \newcommand\abovelabel[2]{ % valname, label
+      \node[font=\tiny, rotate=45, anchor=north west]
+      at ([yshift=5mm,xshift=-3mm]#1.center) {\qq{#1}: #2};
+    }
+    \newcommand\belowlabel[2]{ % valname, label
+      \node[font=\tiny, rotate=-45, anchor=south west]
+      at ([yshift=-6mm,xshift=-3mm]#1.center) {\qq{#1}: #2};
+    }
+
     \sankeyset{
@@
     \sankeynode{name=RF, quantity=\Q{vRF}, at={[xshift=-\xoff]CD.right}, anchor=left}
+
+    \abovelabel{SF}{Snowfall}
+    \abovelabel{DP}{Deposition}
+    \abovelabel{CD}{Condensation}
+    \abovelabel{RF}{Rainfall}
 
@@
         at={[xshift=\xoff,yshift=8mm]SUa.left}, anchor=right, angle=0}
+
+      \node[font=\tiny, anchor=north]
+      at ([yshift=-5mm,xshift=9mm]MGa.bottom) {\qq{MGa}: Accumulation};
+
       \sankeyoutin{MG}{MGa}
@@
           at={[xshift=-\xoff]RF.right}, anchor=left} %, angle=-45}
+        \abovelabel{ML}{Drawdown}
         \sankeystart[red]{ML}
@@
 
+      \belowlabel{RU}{Runoff}
+      \belowlabel{BMa}{Basal melting}
+      \belowlabel{SUBa}{Submarine melting}
+      \belowlabel{D}{Discharge}
+      \belowlabel{GZa}{Grounding line retreat}
+      \belowlabel{FRa}{Frontal retreat}
+
+      \node[font=\tiny, anchor=north]
+      at ([yshift=-2mm,xshift=3mm]SUaa.bottom) {\qq{SUa}: Sublimation};
+
+      \node[font=\tiny, anchor=north]
+      at ([yshift=-2mm,xshift=0mm]EVa.bottom) {\qq{EVa}: Evaporation};
+
+      \node[font=\tiny, anchor=west, rotate=-90]
+      at ([yshift=0.5mm,xshift=1mm]RFZi.center) {\qq{RFZi}: Refreezing};
+
   %%   %% DRIFT

### 2024-06-22 ms.tex
ack
Writing

--- 2024-06-02 ms.tex
@@
 %% manuscript to a two-column layout.
-%%
-%% Use option "jog" for submissions to the Journal of
-%% Glaciology; use "aog" for submission to the Annals
-%% of Glaciology.
-\documentclass[review,jog]{igs}
-% \documentclass[review,aog]{igs}
+
+\documentclass[jog]{igs}
 
@@
        {Kenneth D. MANKOFF,$^{1,2}$
+         Xavier FETTWEIS,
+         Chad GREENE,
          Someone, ELSE,$^n$,
@@
 \begin{abstract}
-  Mass flows of ice sheets, and subsequent mass gains and losses, is an of-studied topic in the domain of glaciology.
-However, few studies consider all processes and their relative magnitudes and uncertainties - this is usually the domain of review papers.
-Here we present a Sankey diagram that provides an overview of all mass flow terms and processes for both the Greenlandic and Antarctic ice sheets.
-We use this display to highlight a) relative magnitudes among processes within ice sheets and among ice sheets and sectors, and b) the uncertainty associated with each property.
-We also...
-TBD.
 
-  There is no new science here, but presenting existing data in new or clearer ways may still lead to new insights.
+  Abstract 1:
+
+  We combine recent estimates of ice mass change that include ice at or near flotation (i.e., terminus change in Greenland) with grounded ice mass balance to produce the first total mass balance product for the Greenlandic and Antarctic ice sheets.
+Greenland mass loss is 30 \% higher than previous estimates due to terminus retreat.
+This additional mass loss does not contribute significantly to sea level rise as it is already at or near flotation.
+
+  Abstract 2:
+
+  Understanding the processes involved in ice sheet mass loss requires knowledge production and sharing, in addition to field observations and computer modeling.
+Tabular data is optimized for computers and useful for humans, but graphical presentations can provide significantly higher information density.
+Here we present several Sankey diagrams depicting all mass flow of water (in various phases) in Greenland and Antarctica.
+These diagrams show mass loss in Greenland is 330 Gt yr$^{-1}$  which is 30 \% higher previous estimates that neglect terminus retreat of ice at or near flotation.
+
 \end{abstract}
@@
 
-\section{Introduction to Sankey diagrams}
+\section{Introduction and Background}
+
+Mass flows of ice sheets, and subsequent mass gains and losses, is an of-studied topic in the domain of glaciology.
+However, few studies consider all processes and their relative magnitudes and uncertainties - this is typically limited to review papers, which may not focus on quantitative assessment of each process.
+
+Net mass flow (mass balance) is typically analyzed using one or a combination of three methods.
+Of these, the gravimetric method cannot observe mass change of floating ice, the volumetric method cannot distinguish between surface and basal processes, and only the input-output method can estimate both mass flow, all processes participating in mass flow, and the subsequent mass change.
+
+Here we present Sankey diagrams that provide an overview of all mass flow terms and processes for both the Greenlandic and Antarctic ice sheets.
+
+% We use this display to highlight a) relative magnitudes among processes within ice sheets and among ice sheets and sectors, and b) the uncertainty associated with each property. We also... TBD.
+
+\begin{figure*}
+\centering{\includegraphics[width=0.85\textwidth]{fig.png}}
+\caption{Sankey mass flow diagrams for Antarctica and Greenland, and Antarctica split into East, West, and Peninsula.
+All widths are proportional within and between images.
+Because Sankey diagrams balance all inputs and outputs, mass losses require a `drawdown' input (red) to balance the larger outputs, and mass gains requires an `accumulation' output (black) to balance the larger inputs.}
+\label{fig}
+\end{figure*}
+
+\subsection{Introduction to Sankey diagrams}
 
 Sankey diagrams are graphical representations of flow or movement of any property (e.g., mass, energy, money, etc.).
-An early and famous use was Charles Minard's Map of Napoleon's Russian Campaign of 1812 (c.f., \citet{kraak_2021}) that combines the magnitude of active soldiers overlaid on a geographical map to show attrition during war.
+An early and famous use was Charles Minard's Map of Napoleon's Russian Campaign of 1812 (c.f., \citet{kraak_2021}) that combines the magnitude of active soldiers overlaid on a geographical map to show attrition during one battle in a war.
 The method was later refined, popularized, and eventually named after Captain Matthew Henry Phineas Riall Sankey who used it to show, among other things, the energy efficiency of a steam engine.
 
-\section{Overview of mass flows}
+  The flows can be somewhat tied to physical space as in the Minard map mentioned above, but are not done so explicitly here.
+An similar display to the diagrams presented here, of glacier mass flow by \citet[Fig.
+2]{cogley_2011}, shows flow overlaid an a glacier schematic, but with neither flow magnitude nor proportionality.
 
-See \url{https://github.com/mankoff/sankey/issues/16} for discussion of graphical plan. Comment there or here.
+\section{Methods}
 
-\begin{figure}
-  \centering
-  \includegraphics[height=.65\linewidth]{fig/gl_baseline.pdf}
-  \includegraphics[height=.65\linewidth]{fig/gl_2019.pdf}
-  \caption{Sankey flow for Greenland, baseline and 2019.
-Widths are proportional.
-Colors represent both phases (gray solid, blue liquid, white gaseous) and net mass change (red loss, black gain).
-TODO: Improved labeling.}
-    \label{fig:gl}
-\end{figure}
+We use existing published values, some new estimates, and some derived values (See Tables \ref{tab:gl} and \ref{tab:aq} in Appendix A).
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.30\linewidth]{fig/aq_baseline.pdf}
-  \includegraphics[width=.30\linewidth]{fig/aq_baseline.pdf}
-  \includegraphics[width=.30\linewidth]{fig/aq_baseline.pdf}
-  \caption{Antarctica, east, west. NOTE east and west are TBD, this is just all AQ shown 3x.}
-  \label{fig:aq}
-\end{figure}
+The derived value is only the net mass change - here shown as `drawdown' in Greenland and West Antarctica, or `accumulation' in Antarctica, East Antarctica, or the Peninsula (Fig \ref{fig}).
+This term balances all the other terms, so that the Sankey diagram has outputs balancing inputs.
 
-\subsection{Interpretation of graphics}
+Here we describe our new estimates.
 
-Sankey diagrams are generally intuitive, but the following section may still be helpful in interpreting the diagrams.
-The widths of any section within any figure is proportional to all other widths.
-Widths are/are-not (TODO, see https://github.com/mankoff/sankey/issues/16) proportional between images.
-The flows can be somewhat tied to physical space or geography, but are not.
-An alternate display by \citet[Fig.
-2]{cogley_2011} shows a similar diagram overlaid an a glacier schematic, but without flow magnitude nor proportionality.
+In Greenland, there is no known assessment of grounding line retreat separate from ice front retreat.
+These two terms are the same in most places in Greenland, because there are few ice shelves.
+In these locations, we used published values from \citet{kochtitzky_2023} and assign this to frontal retreat.
+There are published values of grounding line retreat at some glaciers, but units are typically m and not Gt yr$^{-1}$.
+Here we use published values of Petermann glacier grounding line retreat (units m) from \citet{millan_2022}, ice velocity from \citet{millan_2022}, ice thickness from \citet{ciraci_2023}, and an estimate of ice density of 917 kg m$^{3}$ to calculate grounding line retreat in units of Gt yr$^{-1}$.
+We estimate \sim 1.5 Gt yr$^{-1}$, and round this up to 5 Gt yr$^{-1}$ to account for all other ice shelf grounding line retreat in Greenland.
 
-Color here represents both phase and net mass change.
-Colors gray, blue, and white represent solid, liquid, and gaseous phases respectively, while red and black highlight net mass loss or gain.
-The latter may be counter-intuitive - for example to see mass loss as an input at the top (red in Fig.
-\ref{fig:gl}) even though most mass loss terms (runoff, calving, etc.) are at the bottom.
-This is because Sankey diagrams are balanced, outputs are larger than inputs (hence net mass loss), so the mass loss term is an input - drawdown of the historical 'stable' ice mass.
-Conversely, in East Antarctica mass gain is an output at the bottom (black in Fig.
-\ref{fig:aq}b} that balances the diagram, because without it, there are more flows into the system than out of it.
+Condensation: TODO
 
-%% We mark grounding line in Antarctica and frontal retreat in Greenland in red, even if Antarctica has net mass gain, because not all mass is equal when considering ice sheet health.
-Antarctica is mass positive because of increased snowfalls.
-Grounding line retreat is a localized mass imbalance term.
+Deposition: TODO
 
-\subsection{Missing terms and limitations}
+In Antarctica, we use the MEaSUREs Antarctic Boundaries for IPY 2007-2009 from Satellite Radar, Version 2 (NSIDC product 0709; \citet{mouginot_2017,rignot_2013}) to separate Antarctica into East, West, and Peninsula.
+We drop all unattached islands, so the sum of the regional terms may not equal the total Antarctic values in Fig.
+\ref{fig}.
 
-See \citet[Fig. 2]{cogley_2011}
+Most existing estimates are provided with enough geospatial metadata that we simply assign flow estimates to the appropriate region.
+However, we do not have access to the basal mass balance for Antarctica in spatial map form, only a total Antarctic-wide value from \citet{pattyn_2010} reported at 65 Gt yr$^{-1}$.
+To estimate basal melt for East and West Antarctica, we scale by the magnitude of geothermal heat flux (from \citet{shapiro_2004}) scaled by area for East, West, and Peninsula which is 70 \%, 25 \%, and 5 \% respectively.
+This is slightly better than scaling by area alone which has 75 \% of the area in East Antarctica.
+This method neglects the contribution of shear heating (ice velocity) when partitioning basal melt.
 
-Basal freeze-on (e.g., \citet{bell_2014}) is not shown.
-It is currently unknown by Ken if this is refreezing of the basal melting term (e.g.,\citet{karlsson_2021}) and therefore that terms source remains unchanged, but some diverts back to the dynamics flow and the basal melting output is reduced, or if the basal melting term is final basal runoff, in which case the source should be increased by the refreezing amount.
-Nonetheless, we do not know the refreezing value.
+\section{Missing terms and limitations}
+\label{sec:limits}
 
-Also:
-\begin{itemize}
-\item Avalanche on/off. Maybe matters more for glaciers than ice sheets?
-\item Subaerial frontal melt/sublimation. That is, the vertical face in Greenland above the water line. This is included in other terms.
-\item Snow drifting on/off. Not sure how much snow drifts onto an ice sheet, but snow drift off could be good to include.
-\item Grounding line retreat in Greenland. Different than frontal retreat. Ignored, as there are few ice shelves.
-\end{itemize}
+These figures neglect some mass flow processes (some of which are included in \citet[Fig. 2]{cogley_2011}), and simplify others.
+
+Neglected processes include basal freeze-on (e.g., \citet{bell_2014}).
+It is currently unknown by Ken (TODO: Nanna,  Brice?) if this is refreezing of the basal melting term (i.e., \citet{karlsson_2021} Greenland and \citet{pattyn_2010} Antarctica) and therefore that source remains unchanged, but some diverts back to the dynamics flow and the basal melting output is reduced, or if the basal melting term is final basal runoff, in which case the source should be increased by the refreezing amount.
+Nonetheless, we do not yet know nor estimate the refreezing magnitude.
+
+We also neglect avalanche on and off - these likely matter more for mountain glaciers than ice sheets.
+
+Snow drift on and off is also excluded.
+There is likely little snow drift onto either ice sheets, but drifting off may be of similar magnitude to other terms.
+Some drift off may be implicitly included in the sublimation term (TODO: Xavier?)
 
 There may be other missing terms.
-For example, an earlier version of this graphic by \citet[Fig.
+For example, the earlier version of this graphic by \citet[Fig.
 2]{cogley_2011} did not contain frontal nor grounding line retreat.
@@
 These terms were not only not included in \citet{cogley_2011}, but their respective values were highly uncertain, and still are, although recent work by \citet{kochtitzky_2023,greene_2024} have constrained these values in Greenland.
+
+We make the following simplifications.
+
+Subaerial frontal melt and sublimation (the vertical face in above the water line, \citet[Fig. 2]{cogley_2011}) is not explicitly treated but is included in other terms.
+
+There are a variety of other simplifications.
+For example, rainfall input does not all turn to ice.
+Some enters as part of the refreezing loop, and some remains liquid and leaves as runoff or evaporation.
+Similarly, the evaporation output could pull from the refreezing loop (in the liquid phase, depicted by the blue color) and also directly from rainfall as stated above.
+Although some path details are simplified, the magnitudes are still correct - at least as well as we are able to estimate them.
 
@@
 
-\subsection{Drifting snow}
+This could be the largest section of text.
 
-% I think if you were to go into a discussion of snowdrift it should go further than, for example, the works of Lenaerts et al.
+\section{Comparison with existing estimates}
+
+As previously stated, few existing studies outside of review papers address all terms, and the review papers usually do not focus of quantitative assessment of magnitude.
+Therefore, we compare parts of this graphic to other existing estimates.
+
+The ice sheet mass balance intercomparison experiment (IMBIE; \citet{otosaka_2023}) reports recent Greenlandic ice sheet mass loss as -257 \pm42 Gt yr$^{-1}$.
+Elsewhere the gravimetric method reports recent Greenlandic mass loss of ~277 Gt yr$^{-1}$ (GRACE site, need CITATION).
+These are both significantly less than our estimate of 325 drawdown in order to balance the inputs with outputs.
+This can be directly attributable to the gravimetric method not observing frontal retreat (50 Gt yr$^{-1}$) nor grounding line retreat (5 Gt yr$^{-1}$).
+When these loss terms are removed from our estimate, it becomes 270 Gt yr$^{-1}$ which is well within the uncertainty.
+
+In Antarctica...
+
+%% \subsection{Shelves} %% Suggest remove:
+
+%% Mass change of shelves is a bulk aggregate property, and should not the default reporting metric because it obscures information.
+For example, in theory ice shelf mass can grow even as they collapse, as long as the grounding line retreats (adds mass to the shelf from the upstream ice sheet) faster than the mass loss at the frontal or submarine boundaries.
+A mass flow diagram dedicated to ice shelves (this one is not) would clearly convey each of these processes.
+
+%% \subsection{Drifting snow}
+
+%% %% From Hester: I think if you were to go into a discussion of snowdrift it should go further than, for example, the works of Lenaerts et al.
 Perhaps it is beter to plainly list the uncertainties / poor definitions but in terms of process just refer to the existing papers.
@@
 
-\section{Methods}
+\subsection{Interpretation of graphics}
 
-Use published values. See CSV tables in \texttt{dat/} folder or README.org
+Sankey diagrams are generally intuitive, but the following section may still be helpful in interpreting the diagram(s) shown here.
+The widths of all lines are proportional to all other widths, both within and among figures.
+Color here represents both phase and net mass change.
+Colors gray, blue, and yellow represent solid, liquid, and gaseous phases respectively, while red and black highlight net mass loss or gain.
+The latter may be counter-intuitive - for example to see mass loss as an input at the top (red in Fig.
+\ref{fig}) even though most mass loss terms (runoff, calving, etc.) are at the bottom.
+This is because Sankey diagrams are balanced, outputs are larger than inputs (hence net mass loss), so the mass loss term is an input - drawdown of the historical 'stable' ice mass.
+Conversely, in East Antarctica mass gain is an output at the bottom (black in Fig.
+\ref{fig}} that balances the diagram, because without it, there are more flows into the system than out of it.
 
-Approximate where values are not easily accessed.
+  These diagrams also do not represent every process perfectly.
+For example, frontal retreat is a combination of discharge and submarine melting (and should therefore divide between ice and liquid, with the same uncertainty \citep{enderlin_2013}, but frontal retreat is shown separately here because it is usually treated separately in the literature.
 
-We do not have access to the basal mass balance for Antarctica in spatial map form, only a total Antarctic-wide value from \citet{pattyn_2010} reported at 65 Gt yr^{-1}.
-To estimate basal melt for East and West Antarctica, we scale by the magnitude of geothermal heat flux scaled by area for East, West, and Peninsula which is 70 \%, 25 \%, and 5 \% respectively, compared to scaling by area alone which has 75 \% of the area in East Antarctica.
+  We highlight frontal retreat and grounding line retreat both with a red outline, and by not including frontal retreat in the larger (in Greenland) discharge and submarine melting flow.
+We do this for two reasons.
 
-\section{Mass flow values}
+  First these two terms imply an imbalance.
+Regardless of if a system is gaining mass (Antarctica), losing mass (Greenland) or appears to be in steady state (Peninsula), if there is grounding line and frontal retreat, it implies a system imbalance in the long term, even if not a numerical imbalance as represented here.
 
-See CSV tables in \texttt{dat/} folder or README.org
-
-%% \begin{table} % table1, one column
-%%   \caption{One-column table captions will extend beyond the rules in two-column format. Do not try to adjust! Table captions do not have full points at the end}
-%%   \label{period}
-%%   \begin{minipage}{86mm} % you only need this line if you have a table footnote
-%%     \begin{tabular}{@{}lcc}\hline
-
-%%       Period\footnote{Please do not use more than one `\&' between columns, and note that if a table includes table footnotes, it must be inside a \texttt{minipage} environment.}%
-%%       & Surface elevation change
-%%       & Emergence velocity\\ \hline
-
-%%       1975--85   & $-0.50$ & 0.43\\
-%%       1986--2002 & $-1.03$ & 0.32\\
-%%       Difference & $-0.53$ & \llap{$-$}0.11
-%%     \end{tabular}
-%%   \end{minipage} % you only need this line if you have a table footnote
-%% \end{table}
-
-%% Tables may be typeset in either one- or two-column format.
-To typeset two-column format, add asterisks\\ (\verb"\begin{table*}...\end{table*}") as shown in Table~\ref{seasonal}.
-We may change the format in-house if necessary.
-Please avoid the use of colour or shading.
-Note that if you choose to refer to tables using labels, \verb"\caption" must precede \verb"\label", as in standard \LaTeX.
-Vertical rules are not house-style and will be removed.
-Note the use of the minipage environment in Table~\ref{period} which enables table footnotes to be output.
-If the table is two-column, use \texttt{\{178mm\}} instead of \texttt{\{86mm\}} on line~6.
-The source code for Tables~\ref{period} and~\ref{seasonal} is shown immediately below the tables.
-
-%% \begin{table*}% table2, two column
-%% \caption{Two-column table. Seasonal and annual SAT trends ($^\circ$C\,decade$^{-1}$) in the Arctic}
-%% \label{seasonal}
-%% % the following illustrates how to align columns on decimal points using the S column identifier from siunitx
-%% \setlength\tabcolsep{2.5pt}% column separation reduced from the default 6pt so the table fits the measure
-%% \begin{tabular}{@{}l@{\hspace{20pt}}SSSSS@{\hspace{20pt}}SSSSS}\hline
-%% Area                 & \multicolumn{5}{c}{1951--2005} & \multicolumn{5}{c}{1976--2005}\\[5pt]
-%%                      & {Dec--Feb}       & {Mar--May}    & {Jun--Aug}  & {Sep--Nov}         & {Annual}
-%%                      & {Dec--Feb}       & {Mar--May}    & {Jun--Aug}  & {Sep--Nov}         & {Annual}\\ \hline
-%% Atlantic region      & 0.09           & 0.29 & 0.10 & 0.09 & 0.15 & 0.470 & 0.60 & 0.45 & 0.53 & 0.59\\
-%% Siberian region      & 0.12           & 0.29 & 0.04 & 0.17 & 0.16 & 0.08 & 0.69 & 0.29 & 0.59 & 0.48\\
-%% Pacific region       & 0.45           & 0.46 & 0.25 & 0.26 & 0.35 & 0.712 & 1.08 & 0.27 & 0.66 & 0.52\\
-%% Canadian region      & 0.16           & 0.12 & 0.14 & 0.30 & 0.18 & 0.20 & 0.52 & 0.48 & 0.94 & 0.53\\
-%% Baffin Bay region    & -0.02 & 0.10 & 0.00 & 0.15 & 0.02 & 0.33 & 0.62 & 0.51 & 0.80 & 0.57\\
-%% Arctic 1             & 0.16           & 0.21 & 0.12 & 0.20 & 0.18 & 0.36 & 200.65 & 0.42 & 0.74 & 0.54\\
-%% Arctic 2             & 0.22           & 0.29 & 0.14 & 0.14 & 0.19 & 0.38 & 0.60 & 0.40 & 0.51 & 0.45\\
-%% Arctic 3             & 0.28           & 0.31 & 0.14 & 0.13 & 0.21 & 0.42 & 40.53 & 0.41 & 0.42 & 0.43\\
-%% NH ($\mathrm{land}
-%%   + \mathrm{ocean}$) & 0.13           & 0.13 & 0.10 & 0.10 & 0.12 & 0.27 & 0.24 & 0.25 & 0.25 & 0.25\\
-%% \hline
-%% \end{tabular}
-%% \end{table*}
-
-%% \subsection{Figures}
-
-%% Figures may be typeset in either one- or two-column format.
-One-column format allows up to 86$\,$mm (e.g.
-Fig.~\ref{tracks}); two-column format up to 178$\,$mm (e.g.
-Fig.~\ref{filters}).
-Please do not provide original graphics files in which the figure is a great deal larger or smaller than what you envisage will be the final printed size.
-To typeset two-column format, add asterisks (\verb"\begin{figure*}...\end{figure*}") as shown in Fig.~\ref{filters}.
-We may change the format in-house if necessary.
-Please note that if you choose to refer to figures using labels, \verb"\caption" must precede \verb"\label", as in standard \LaTeX.
-
-%% Please send one file for each figure (in other words do not use subfigures) and use a name that clearly identifies it (e.g. `72A712Fig03.eps').
-
-%% In addition, figures should be eps, ai (illustrator), ps, tif, psd or pdf.
-Use strong black lines with a width of at least 0.75pt at final printed size (avoid tinting if possible) and SI units in labels.
-Lettering should ideally be Optima to match the final typeface; Arial or a similar sans serif font for a second choice.
-Aim to have the final-size lettering at 9pt, if possible.
-Figures should not be in boxes.
-The source code for Figs~\ref{tracks} and~\ref{filters} is shown immediately below the figures.
-
-%% \subsection{Equations}
-
-%% We are including some complex equations as examples. Equations should be checked for width by removing the \verb"[review]" option. Note the use of \verb|cases*| in the following equation:
-%% \begin{equation}
-%% \label{arrayexample}
-%% \alpha_{t_2}= %
-%%   \begin{cases}
-%%     \alpha_{t_1} - a_1 [\ln (T+1)]
-%%       \mathrm{e}^{(a_2\sqrt{n})}
-%%       & \mbox{$n_\mathrm{d} > 0\enskip$ and
-%%       $\enskip T > 0$}\\
-%%     \alpha_{t_1} - a_3 \mathrm{e}^{(a_2\sqrt{n})}
-%%       & \mbox{$n_\mathrm{d} > 0\enskip$ and
-%%       $\enskip T < 0$}\\
-%%     \alpha_{t_1} + a_4 P_\mathrm{s}
-%%       & \mbox{$n_\mathrm{d} = 0$}
-%%   \end{cases}
-%% \end{equation}
-
-%% Equations should be aligned on the equals signs where possible. Equations that extend beyond the one-column measure should be turned over before an operator.
-
-%% \begin{equation}
-%% \label{eqnarrayexample}
-%% \begin{aligned}
-%% l_c & =  l_0 \left(\frac{\widebar{R}_m}{R} \right)^2
-%%   \psi^{\frac{P}{P_0\cos Z}}\\
-%%     & \phantom{=} \quad  \times [\cos\beta\, \cos Z
-%%     + \sin\beta\,\sin Z\,\cos(\psi_\mathrm{sun}
-%%     - \psi_\mathrm{slope})]
-%% \end{aligned}
-%% \end{equation}
-
-%% \citep{mankoff_2020_solid}
-%% \citet{mankoff_2021}
-
-%% \begin{figure}%fig1, one column
-%% \centering{\includegraphics{72A712Fig01.eps}}
-%% \caption{One-column figures should be $\leq$86$\,$mm. Good artwork can
-%%   make or break a paper. Capitalize the first word of a label and use
-%%   round not square brackets for units.}
-%% \label{tracks}
-%% \end{figure}
-
-%% \paragraph{Citations using natbib commands}
-%% Note that the standard natbib style file has been modified to fall into line with IGS style.
-The modified style file is called igsnatbib.sty (included in this distribution), and works exactly the same as natbib.sty.
-The default IGS house style is \citep{Yan13}.
-The following combinations are also available -- refer to the natbib documentation if you require any further explanation:\\*[0.5\baselineskip]
-%% \begin{tabular}{@{}l@{\hskip -94pt}l}
-%% \citep{Yan13}
-%%     & \verb"\citep{Yan13}"\\
-%% \citep[see][p.$\,$34]{Yan13}\\
-%%     & \verb"\citep[see][p.$\,$34]{Yan13}"\\
-%% \citep[e.g.][]{Yan13}
-%%     & \verb"\citep[e.g.][]{Yan13}"\\
-%% \citep[Section~2.3]{Yan13}\\
-%%     & \verb"\citep[Section~2.3]{Yan13}"\\
-%% \citep{Yan13, Edwards14}\\
-%%     &  \verb"\citep{Yan13, Edwards14}"\\
-%% \cite{Yan13, Edwards14}\\
-%%     &  \verb"\cite{Yan13, Edwards14}"\\
-%% \citealt{Yan13}
-%%     & \verb"\citealt{Yan13}"\\
-%% \cite{Yan13}
-%%     & \verb"\cite{Yan13}"\\
-%% \citealp{Yan13}
-%%     & \verb"\citealp{Yan13}"\\
-%% \citeauthor{Yan13}
-%%     & \verb"\citeauthor{Yan13}"\\
-%% \citeyearpar{Yan13}
-%%     & \verb"\citeyearpar{Yan13}"\\
-%% \citeyear{Yan13}
-%%     & \verb"\citeyear{Yan13}"
-%% \end{tabular}
-
-%% \begin{figure*}%fig2, two column
-%% \centering{\includegraphics{72A712Fig02.eps}}
-%% \caption{Two-column figures should be $\leq$178$\,$mm. SSA reconstructed components found by 
-%%   projecting the SSA filters found using the whole 2000 traces in Fig.~4, on trace number 1, 
-%%   ordered by magnitude of variance accounted for in the radar trace.}
-%% \label{filters}
-%% \vspace\baselineskip\hrule % to separate figure from verbatim
-%% \vspace\baselineskip
-%% \begin{verbatim}
-%% \begin{figure*}%fig2, two column
-%% \centering{\includegraphics{72A712Fig02.eps}}
-%% \caption{Two-column figures should be $\leq$178$\,$mm. SSA reconstructed components found by 
-%%   projecting the SSA filters found using the whole 2000 traces in Fig.~4, on trace number 1, 
-%%   ordered by magnitude of variance accounted for in the radar trace.}
-%% \label{filters}
-%% \end{figure*}
-%% \end{verbatim}
-%% \vspace\baselineskip\hrule % to separate verbatim from text
-%% \end{figure*}
+  Secondly, these two terms are rarely included in mass change estimates.
+The gravimetric method does not see these processes, the volumetric method is usually cropped at the some fixed grounding line upstream of these processes, and the IO method has typically ignored these two terms as downstream of the flux gates.
 
@@
 
-LaTeX software community, Cogley for a reference graphic.
+We thank \citep{sankey} for the \LaTeX\ TikZ Sankey package, and \citet{cogley_2011} for a reference graphic.
+Analysis was aided by the software packages Pandas (\citet{pandas_team}), Xarray (\citet{xarray}), and GRASS GIS (\citet{GRASS}), among other tools.
 
-% authors generating their own bbl file would uncomment the following two lines, and comment out/delete the references below:
+We thank Jakob Steiner, Xavier Fettweis, Benjamin Davison, Anna Hogg, Chad Greene, Katie Leonard, Jan Lenaerts, Damien Ringeisen, Liam Colgan, Robert Fausto, Dominik Fahrner, Nanna Karlsson, Brice Van Liefferinge, and Andreas Ahlstrøm for conversations in the development of this work.
 
-\bibliography{library}   % reads igsrefs.bib
-\bibliographystyle{igs}  % imposes IGS bibliography style on output
+\bibliography{library}
+\bibliographystyle{igs}
 
-%% \appendix
-%% \section{Appendix}
+\appendix
+\section{Appendix A}
+\label{sec:appendix:A}
 
-%% Start an appendix by \hbox{typing \verb"\appendix\section{Appendix}".} Appendices appear after the references. Equation numbers automatically start again with (\ref{appeqn}).
-%% \begin{equation}
-%% \label{appeqn}
-%%  2\eta\kappa \frac{\partial \skew1\bar u}{\partial t} + \rho_{\mathrm r} g \skew1\bar u + D\kappa^4 \skew1\bar u = \skew3\bar\sigma_{zz}.
-%% \end{equation}
+\begin{table*}
+\caption{Mass flow values used for Greenland in Fig. \ref{fig}}
+\label{tab:gl}
+\begin{tabular}{@{}lcc}\hline
+  Term & Source\\
+  Snowfall & \citet{fettweis_2020}\\
+  Rainfall & \citet{fettweis_2020}\\
+  Condensation & TODO\\
+  Deposition & TODO\\
+  Sublimation & \citet{fettweis_2020}\\
+  Evaporation & \citet{fettweis_2020}\\
+  Runoff & \citet{fettweis_2020}\\
+  Refreezing & \citet{fettweis_2020}\\
+  Frontal retreat & \citet{kochtitzky_2023,greene_2024}\\
+  Grounding line retreat & Estimated here\\
+  Discharge + submarine melting & \citet{mankoff_2020_solid}\\
+  Split of discharge \& submarine melting & \citet{enderlin_2013}\\
+  Basal melting & \citet{karlsson_2021}\\
+  Drawdown & Derived from above terms
+\end{tabular}
+\end{table*}
 
-%% \section{Handling more than one appendix}
-%% Use the following code to achieve heading APPENDIX~A followed by APPENDIX~B and APPENDIX~C, with appropriate equation numbers:
-
-%% \begin{verbatim}
-
-%% \appendix
-%% \section{Appendix A}
-
-%% \setcounter{equation}{0}
-%% \renewcommand\theequation{B\arabic{equation}}
-%% \section{Appendix B}
-
-%% \setcounter{equation}{0}
-%% \renewcommand\theequation{C\arabic{equation}}
-%% \section{Appendix C}
-%% \end{verbatim}
+\begin{table*}
+\caption{Mass flow values used for Antarctica and Antarctica regions in Fig. \ref{fig}}
+\label{tab:aq}
+\begin{tabular}{@{}lcc}\hline
+  Term & Source\\
+  Snowfall & \citet{fettweis_2020}\\
+  Rainfall & \citet{fettweis_2020}\\
+  Condensation & TODO\\
+  Deposition & TODO\\
+  Sublimation & \citet{fettweis_2020}\\
+  Evaporation & \citet{fettweis_2020}\\
+  Runoff & \citet{fettweis_2020}\\
+  Refreezing & \citet{fettweis_2020}\\
+  Frontal retreat & TODO\\
+  Grounding line retreat & TODO\\
+  Discharge + submarine melting & \citet{davison_2023}\\
+  Split of discharge \& submarine melting & \citet{davison_2023}\\
+  Basal melting & \citet{pattyn_2010}\\
+  Regional distribution terms & See Sect. \ref{sec:methods:split}\\
+  Drawdown or accumulation & Derived from above
+\end{tabular}
+\end{table*}
 

### 2024-06-24 sankey.tex
Antarctic frontal retreat from Greene (2022)
Closes #11

@@
         {vML/Mass loss/
-          (\Q{vSU}+\Q{vRU}+\Q{vBM}+\Q{vDYN}+\Q{vFRET}+\Q{vGZRET}+\Q{vEV})-
-          (\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP})},
+          (\Q{vSU}+\Q{vRU}+\Q{vBM}+\Q{vDYN}+\Q{vFRLOSS}+\Q{vGZRET}+\Q{vEV})-
+          (\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP}+\Q{vFRGAIN})},
         {vMG/Mass gain/
-          (\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP})-
-          (\Q{vSU}+\Q{vRU}+\Q{vBM}+\Q{vDYN}+\Q{vFRET}+\Q{vGZRET}+\Q{vEV})}
+          (\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP}+\Q{vFRGAIN})-
+          (\Q{vSU}+\Q{vRU}+\Q{vBM}+\Q{vDYN}+\Q{vFRLOSS}+\Q{vGZRET}+\Q{vEV})}
       },
@@
 
-    \ifthenelse{\Q{vMG} > \Q{vML}}
-    { % mass gain
+    \ifthenelse{\Q{vMG} > \Q{vML}} {
+      \ifthenelse{\Q{vFRLOSS} > 0}{
         \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0},
@@
           forked={\qq{f0}/f0i, \Q{vRFZ}/RFZi}}
+      }{
+        \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vFRGAIN},
+          at={[yshift=-15mm]f0.left}, anchor=left,
+          forked={\Q{vFRGAIN}/FG, \qq{f0}/f0i, \Q{vRFZ}/RFZi}}
       }
-    { % mass loss
+    } {
+      \ifthenelse{\Q{vFRLOSS} > 0}{
         \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vML},
@@
           forked={\qq{f0}/f0i, \Q{vML}/MLi, \Q{vRFZ}/RFZi}}
+      }{
+        \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vML}+\Q{vFRGAIN},
+          at={[yshift=-15mm]f0.left}, anchor=left,
+          forked={\Q{vFRGAIN}/FG, \qq{f0}/f0i, \Q{vML}/MLi, \Q{vRFZ}/RFZi}}
       }
+    }
+
+    \ifthenelse{\Q{vFRGAIN} > 0}{
+      \sankeynode{name=FGi, quantity=\Q{vFRGAIN},
+        at={[yshift=15mm,xshift=\xoff]FG.left}, anchor=right}
+      \sankeystart{FGi}
+      \abovelabel{FGi}{Frontal advance}
+      \sankeyoutin{FGi}{FG}
+    }{}
 
@@
         \Q{vMG}/MG,
-        \Q{vSU}/SU, \Q{vFRET}/FR, \Q{vGZRET}/GZ,
+        \Q{vSU}/SU,
+        \Q{vFRLOSS}/FL,
+        \Q{vGZRET}/GZ,
         \Q{vDYN}/f6, \Q{vBM}/BM, \Q{vRU}+\Q{vEV}/f8, \Q{vRFZ}/RFZo}
@@
         %% \Q{vMG}/MG,
-        \Q{vSU}/SU, \Q{vFRET}/FR, \Q{vGZRET}/GZ,
+        \Q{vSU}/SU,
+        \Q{vFRLOSS}/FL,
+        \Q{vGZRET}/GZ,
         \Q{vDYN}/f6, \Q{vBM}/BM, \Q{vRU}+\Q{vEV}/f8, \Q{vRFZ}/RFZo}
@@
 
-    \sankeynode{name=FRa, quantity=\qq{FR},
+    % FLa must exist for placement, but only draw lines if > 0
+    \ifthenelse{\Q{vFRLOSS} > 0} {
+        \sankeynode{name=FLa, quantity=\qq{FL},
           at={[xshift=\xoff]GZa.left}, anchor=right}
-    \sankeyoutin{FR}{FRa}
+        \sankeyoutin{FL}{FLa}
+      }{
+        \sankeynode{name=FLa, quantity=\qq{FL},
+          at={[yshift=0mm]GZa.left}, anchor=right}
+      }
 
     \sankeynode{name=SUa, quantity=\qq{SU},
-      at={[xshift=\xoff]FRa.left}, anchor=right}
+      at={[xshift=\xoff]FLa.left}, anchor=right}
     \sankeyoutin{SU}{SUa}
+
+    \ifthenelse{\Q{vFRLOSS} > 0} {
+      \sankeyadvance[out-red]{FLa}{10mm}
+      \sankeyend[out-red]{FLa}
+      \belowlabel{FLa}{Frontal retreat}
+    }{}
 
@@
       \sankeynode{name=MGa, quantity=\qq{MG},
-        at={[xshift=\xoff,yshift=8mm]SUa.left}, anchor=right, angle=0}
+        at={[xshift=3mm,yshift=-3mm]MG.left}, anchor=left, angle=0}
 
       \node[font=\tiny, anchor=north]
-      at ([yshift=-5mm,xshift=9mm]MGa.bottom) {\qq{MGa}: Accumulation};
+      at ([yshift=0mm,xshift=3mm]MGa.south) {\qq{MGa}: Accumulation};
 
@@
 
-      \sankeyadvance[out-red]{FRa}{10mm}
-      \sankeyend[out-red]{FRa}
       \sankeyadvance[ice-to-water,out-red]{GZa}{10mm}
@@
       \belowlabel{GZa}{Grounding line retreat}
-      \belowlabel{FRa}{Frontal retreat}
 
       \node[font=\tiny, anchor=north]
-      at ([yshift=-2mm,xshift=3mm]SUaa.bottom) {\qq{SUa}: Sublimation};
+      at ([yshift=-2mm,xshift=3mm]SUaa.east) {\qq{SUa}: Sublimation};
 
       \node[font=\tiny, anchor=north]
-      at ([yshift=-2mm,xshift=0mm]EVa.bottom) {\qq{EVa}: Evaporation};
+      at ([yshift=-2mm,xshift=0mm]EVa.west) {\qq{EVa}: Evaporation};
 
@@
 
-  %%   %% DRIFT
-  %%   \sankeynode{name=DRIFTo, quantity=33, at={(3mm, -8mm)}, angle=150}
-  %%   \sankeynode{name=DRIFTi, quantity=33, at={(-3mm, -8mm)}, angle=210}
-  %%   \sankeydubins[minimum radius=5mm]{DRIFTo}{DRIFTi}
-  %%   \sankeystart{DRIFTo}
-  %%   \sankeyend{DRIFTI}
-  %%   \node[anchor=center, align=center, rotate=0, shift=({0,0})] at (0,-15mm) {Drifting (?)};
-
   \end{sankeydiagram}

### 2024-06-24 ms.tex
Basal melt in AQ. Closes #14

@@
 \ref{fig}.
-
-Most existing estimates are provided with enough geospatial metadata that we simply assign flow estimates to the appropriate region.
-However, we do not have access to the basal mass balance for Antarctica in spatial map form, only a total Antarctic-wide value from \citet{pattyn_2010} reported at 65 Gt yr$^{-1}$.
-To estimate basal melt for East and West Antarctica, we scale by the magnitude of geothermal heat flux (from \citet{shapiro_2004}) scaled by area for East, West, and Peninsula which is 70 \%, 25 \%, and 5 \% respectively.
-This is slightly better than scaling by area alone which has 75 \% of the area in East Antarctica.
-This method neglects the contribution of shear heating (ice velocity) when partitioning basal melt.
 
@@
   Split of discharge \& submarine melting & \citet{davison_2023}\\
-  Basal melting & \citet{pattyn_2010}\\
-  Regional distribution terms & See Sect. \ref{sec:methods:split}\\
+  Basal melting & \citet{van-liefferinge_2013}\\
   Drawdown or accumulation & Derived from above

### 2024-06-24 sankey.tex
Improve labels (all rotated the same)

@@
     \newcommand\belowlabel[2]{ % valname, label
-      \node[font=\tiny, rotate=-45, anchor=south west]
-      at ([yshift=-6mm,xshift=-3mm]#1.center) {\qq{#1}: #2};
+      \node[font=\tiny, rotate=45, anchor=north east]
+      at ([yshift=-2mm,xshift=0mm]#1.center) {#2: \qq{#1}};
     }
@@
 
-      \node[font=\tiny, anchor=north]
-      at ([yshift=-2mm,xshift=3mm]SUaa.east) {\qq{SUa}: Sublimation};
+      \node[font=\tiny, rotate=45, anchor=north east]
+      at ([yshift=1mm,xshift=3mm]SUaa.south) {Sublimation: \qq{SUaa}};
 
-      \node[font=\tiny, anchor=north]
-      at ([yshift=-2mm,xshift=0mm]EVa.west) {\qq{EVa}: Evaporation};
+      \node[font=\tiny, rotate=45, anchor=north east]
+      at ([yshift=0mm,xshift=5mm]EVa.north) {Evaporation: \qq{EVa}};
 
-      \node[font=\tiny, anchor=west, rotate=-90]
-      at ([yshift=0.5mm,xshift=1mm]RFZi.center) {\qq{RFZi}: Refreezing};
+      \node[font=\tiny, anchor=west, rotate=90]
+      at ([yshift=-0.5mm,xshift=1mm]RFZo.center) {Refreezing: \qq{RFZi}};
 

### 2024-06-24 ms.tex
Brice is co-author
There is an implicit physical flow

@@
          Chad GREENE,
+         Brice VAN LIEFFERINGE,
          Someone, ELSE,$^n$,
@@
 2]{cogley_2011}, shows flow overlaid an a glacier schematic, but with neither flow magnitude nor proportionality.
+Nonetheless, there is an implicit physical connection here for most flow.
+Inputs all come from the atmosphere (with the exception of drawdown and frontal advance), and outputs pointing down are all mass flow into the ocean.
 

### 2024-06-24 sankey.tex
Adjust frontal advance

@@
       ice-to-black/.style={fill/.style={fill=none, left color=ice, right color=black},},
+      black-to-ice/.style={fill/.style={fill=none, bottom color=ice, top color=black},},
       out-red/.style={draw/.style={draw=red,thick=3},},
@@
           at={[yshift=-15mm]f0.left}, anchor=left,
-          forked={\Q{vFRGAIN}/FG, \qq{f0}/f0i, \Q{vRFZ}/RFZi}}
+          forked={\qq{f0}/f0i, \Q{vFRGAIN}/FG, \Q{vRFZ}/RFZi}}
       }
@@
           at={[yshift=-15mm]f0.left}, anchor=left,
-          forked={\Q{vFRGAIN}/FG, \qq{f0}/f0i, \Q{vML}/MLi, \Q{vRFZ}/RFZi}}
+          forked={\qq{f0}/f0i, \Q{vML}/MLi, \Q{vFRGAIN}/FG, \Q{vRFZ}/RFZi}}
       }
@@
       \sankeynode{name=FGi, quantity=\Q{vFRGAIN},
-        at={[yshift=15mm,xshift=\xoff]FG.left}, anchor=right}
-      \sankeystart{FGi}
-      \abovelabel{FGi}{Frontal advance}
-      \sankeyoutin{FGi}{FG}
+        at={[yshift=15mm,xshift=-\xoff]FG.right}, anchor=left}
+      \sankeystart[black]{FGi}
+      \sankeyoutin[black-to-ice]{FGi}{FG}
+      \node[font=\tiny, rotate=45, anchor=north east]
+      at ([yshift=12mm,xshift=3mm]FGi.right) {Frontal advance: \qq{FGi}}; 
    }{}

### 2024-06-24 sankey.tex
Accumulation -> Mass gain
Discharge -> Calving
Show both frontal advance and retreat in same graphic

--- 2024-06-24 sankey.tex
@@
     \colorlet{black}[rgb]{black!100!black!100!black!100}
+    \colorlet{gray}[rgb]{black!30!black!30!black!30}
 
@@
       ice-to-black/.style={fill/.style={fill=none, left color=ice, right color=black},},
+      black-to-ice/.style={fill/.style={fill=none, bottom color=ice, top color=black},},
+      in-black/.style={draw/.style={draw=gray,thick=3},},
       out-red/.style={draw/.style={draw=red,thick=3},},
@@
     \ifthenelse{\Q{vMG} > \Q{vML}} {
-      \ifthenelse{\Q{vFRLOSS} > 0}{
-        \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0},
-          at={[yshift=-15mm]f0.left}, anchor=left,
-          forked={\qq{f0}/f0i, \Q{vRFZ}/RFZi}}
-      }{
         \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vFRGAIN},
           at={[yshift=-15mm]f0.left}, anchor=left,
-          forked={\Q{vFRGAIN}/FG, \qq{f0}/f0i, \Q{vRFZ}/RFZi}}
-      }
+          forked={\qq{f0}/f0i, \Q{vFRGAIN}/FG, \Q{vRFZ}/RFZi}}
     } {
-      \ifthenelse{\Q{vFRLOSS} > 0}{
-        \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vML},
-          at={[yshift=-15mm]f0.left}, anchor=left,
-          forked={\qq{f0}/f0i, \Q{vML}/MLi, \Q{vRFZ}/RFZi}}
-      }{
         \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vML}+\Q{vFRGAIN},
           at={[yshift=-15mm]f0.left}, anchor=left,
-          forked={\Q{vFRGAIN}/FG, \qq{f0}/f0i, \Q{vML}/MLi, \Q{vRFZ}/RFZi}}
+          forked={\qq{f0}/f0i, \Q{vML}/MLi, \Q{vFRGAIN}/FG, \Q{vRFZ}/RFZi}}
     }
-    }
-
-    \ifthenelse{\Q{vFRGAIN} > 0}{
-      \sankeynode{name=FGi, quantity=\Q{vFRGAIN},
-        at={[yshift=15mm,xshift=\xoff]FG.left}, anchor=right}
-      \sankeystart{FGi}
-      \abovelabel{FGi}{Frontal advance}
-      \sankeyoutin{FGi}{FG}
-    }{}
 
@@
 
-    % FLa must exist for placement, but only draw lines if > 0
-    \ifthenelse{\Q{vFRLOSS} > 0} {
     \sankeynode{name=FLa, quantity=\qq{FL},
@@
     \sankeyoutin{FL}{FLa}
-      }{
-        \sankeynode{name=FLa, quantity=\qq{FL},
-          at={[yshift=0mm]GZa.left}, anchor=right}
-      }
 
@@
 
-    \ifthenelse{\Q{vFRLOSS} > 0} {
     \sankeyadvance[out-red]{FLa}{10mm}
@@
     \belowlabel{FLa}{Frontal retreat}
-    }{}
 
@@
       \node[font=\tiny, anchor=north]
-      at ([yshift=0mm,xshift=3mm]MGa.south) {\qq{MGa}: Accumulation};
+      at ([yshift=0mm,xshift=3mm]MGa.south) {\qq{MGa}: Mass gain};
 
@@
 
-    \ifthenelse{\Q{vMG} < \Q{vML}}
-      {
+    \ifthenelse{\Q{vMG} < \Q{vML}} {
       %% mass loss
@@
       \abovelabel{ML}{Drawdown}
+
+      % frontal advance next to drawdown (mass loss)
+      \sankeynode{name=FGi, quantity=\Q{vFRGAIN},
+        at={[yshift=-5mm,xshift=-\xoff]ML.right}, anchor=left}
+
       \sankeystart[red]{ML}
@@
       \sankeyoutin{ML}{MLi}
-      }
-      {
+    }{
       %% mass gain
@@
       \sankeyend[black]{MGa}
+
+      % frontal with no drawdown
+      \sankeynode{name=FGi, quantity=\Q{vFRGAIN},
+        at={[yshift=15mm,xshift=3mm]RFZi.right}, anchor=left}
     }
+
+    \ifnum \numexpr\qq{FGi} > 0
+    \sankeystart[in-black]{FGi}
+    \sankeyoutin[in-black]{FGi}{FG}
+    \node[font=\tiny, rotate=45, anchor=north east]
+    at ([yshift=12mm,xshift=3mm]FGi.right) {Frontal advance: \qq{FGi}};
+    \fi
 
@@
     \belowlabel{SUBa}{Submarine melting}
-      \belowlabel{D}{Discharge}
+    \belowlabel{D}{Calving}
     \belowlabel{GZa}{Grounding line retreat}

### 2024-06-25 sankey.tex
solving the difference of color problem in arrows

@@
 \begin{document}
+
+% Can be rbg or cmyk, cmyk is better for printing
+\selectcolormodel{cmyk}
+
 \begin{tikzpicture}[remember picture, overlay, shift={(5.5,-1.5)}]
@@
       ice-to-air/.style={fill/.style={fill=none, top color=ice, bottom color=air},},
-      ice-to-air-horiz/.style={fill/.style={fill=none, left color=ice, right color=air},},
+      ice-to-air-horiz/.style={fill/.style={fill=air, left color=ice, right color=air},},
       water-to-air-horiz/.style={fill/.style={fill=none, right color=water, left color=air},},

### 2024-06-29 sankey.tex
tiny correction
Formatting
Labels from table

--- 2024-06-24 sankey.tex
@@
 \begin{document}
+
+\selectcolormodel{cmyk} % Can be rbg or cmyk, cmyk is better for printing
+
 \begin{tikzpicture}[remember picture, overlay, shift={(5.5,-1.5)}]
@@
 
-    \newcommand\abovelabel[2]{ % valname, label
+    \newcommand\abovelabel[1]{ % valname
       \node[font=\tiny, rotate=45, anchor=north west]
-      at ([yshift=5mm,xshift=-3mm]#1.center) {\qq{#1}: #2};
+      at ([yshift=5mm,xshift=-3mm]#1.center) {\qq{#1}: \Lab{v#1}};
     }
-    \newcommand\belowlabel[2]{ % valname, label
+    \newcommand\belowlabel[2]{ % valname
       \node[font=\tiny, rotate=45, anchor=north east]
-      at ([yshift=-2mm,xshift=0mm]#1.center) {#2: \qq{#1}};
+      at ([yshift=-2mm,xshift=0mm]#1.center) {\Lab{#2}: \qq{#1}};
     }
@@
 
-    \abovelabel{SF}{Snowfall}
-    \abovelabel{DP}{Deposition}
-    \abovelabel{CD}{Condensation}
-    \abovelabel{RF}{Rainfall}
+    \abovelabel{SF}
+    \abovelabel{DP}
+    \abovelabel{CD}
+    \abovelabel{RF}
 
@@
 
-    \sankeynode{name=SUBa, quantity=\qq{SUB},
-      at={[xshift=-\xoff]D.right}, anchor=left}
+    \sankeynode{name=SUBa, quantity=\qq{SUB}, at={[xshift=-\xoff]D.right}, anchor=left}
     \sankeyoutin[ice-to-water]{SUB}{SUBa}
 
-    \sankeynode{name=BMa, quantity=\qq{BM},
-      at={[xshift=-\xoff]SUBa.right}, anchor=left}
+    \sankeynode{name=BMa, quantity=\qq{BM}, at={[xshift=-\xoff]SUBa.right}, anchor=left}
     \sankeyoutin[ice-to-water]{BM}{BMa}
 
-    \sankeynode{name=f8a, quantity=\qq{f8},
-      at={[xshift=-\xoff]BMa.right}, anchor=left}
+    \sankeynode{name=f8a, quantity=\qq{f8}, at={[xshift=-\xoff]BMa.right}, anchor=left}
     \sankeyoutin[ice-to-water]{f8}{f8a}
 
-    \sankeynode{name=GZa, quantity=\qq{GZ},
-      at={[xshift=\xoff]D.left}, anchor=right}
+    \sankeynode{name=GZa, quantity=\qq{GZ}, at={[xshift=\xoff]D.left}, anchor=right}
     \sankeyoutin{GZ}{GZa}
 
-    \sankeynode{name=FLa, quantity=\qq{FL},
-      at={[xshift=\xoff]GZa.left}, anchor=right}
+    \sankeynode{name=FLa, quantity=\qq{FL}, at={[xshift=\xoff]GZa.left}, anchor=right}
     \sankeyoutin{FL}{FLa}
 
-    \sankeynode{name=SUa, quantity=\qq{SU},
-      at={[xshift=\xoff]FLa.left}, anchor=right}
+    \sankeynode{name=SUa, quantity=\qq{SU}, at={[xshift=\xoff]FLa.left}, anchor=right}
     \sankeyoutin{SU}{SUa}
@@
     \sankeyend[out-red]{FLa}
-    \belowlabel{FLa}{Frontal retreat}
+    \belowlabel{FLa}{vFRLOSS}
 
@@
         at={[xshift=3mm,yshift=-3mm]MG.left}, anchor=left, angle=0}
-
       \node[font=\tiny, anchor=north]
       at ([yshift=0mm,xshift=3mm]MGa.south) {\qq{MGa}: Mass gain};
-
       \sankeyoutin{MG}{MGa}
@@
 
-    \ifthenelse{\Q{vMG} < \Q{vML}} {
-      %% mass loss
-      \sankeynode{name=ML, quantity=\Q{vML},
-        at={[xshift=-\xoff]RF.right}, anchor=left} %, angle=-45}
-      \abovelabel{ML}{Drawdown}
+    \ifthenelse{\Q{vMG} < \Q{vML}} { % mass loss
+      \sankeynode{name=ML, quantity=\Q{vML}, at={[xshift=-\xoff]RF.right}, anchor=left}
+      \abovelabel{ML}
 
@@
       \sankeyoutin{ML}{MLi}
-    }{
-      %% mass gain
+    }{ % mass gain
       \sankeyadvance[ice-to-black]{MGa}{3mm}
       \sankeyend[black]{MGa}
-
-      % frontal with no drawdown
       \sankeynode{name=FGi, quantity=\Q{vFRGAIN},
@@
 
-    \belowlabel{RU}{Runoff}
-    \belowlabel{BMa}{Basal melting}
-    \belowlabel{SUBa}{Submarine melting}
-    \belowlabel{D}{Calving}
-    \belowlabel{GZa}{Grounding line retreat}
+    \belowlabel{RU}{vRU}
+    \belowlabel{BMa}{vBM}
+    \belowlabel{SUBa}{vSUB}
+    \belowlabel{D}{vICE}
+    \belowlabel{GZa}{vGZRET}
 
     \node[font=\tiny, rotate=45, anchor=north east]
-    at ([yshift=1mm,xshift=3mm]SUaa.south) {Sublimation: \qq{SUaa}};
+    at ([yshift=1mm,xshift=3mm]SUaa.south) {\Lab{vSU}: \qq{SUaa}};
 
     \node[font=\tiny, rotate=45, anchor=north east]
-    at ([yshift=0mm,xshift=5mm]EVa.north) {Evaporation: \qq{EVa}};
+    at ([yshift=0mm,xshift=5mm]EVa.north) {\Lab{vEV}: \qq{EVa}};
 
     \node[font=\tiny, anchor=west, rotate=90]
-    at ([yshift=-0.5mm,xshift=1mm]RFZo.center) {Refreezing: \qq{RFZi}};
+    at ([yshift=-0.5mm,xshift=1mm]RFZo.center) {\Lab{vRFZ}: \qq{RFZi}};
 

### 2024-07-05 sankey.tex
Horizontal flow and re-order outputs top-to-bottom

@@
 
-\begin{tikzpicture}[remember picture, overlay, shift={(5.5,-1.5)}]
+\begin{tikzpicture}[remember picture, overlay, shift={(5,-5)}]
 
@@
       },
-      every node/.style={angle=-90},
       start style=arrow,end style=arrow,
       %% fill/.style={fill=yellow!10,line width=0pt,draw=yellow!10},
-      ratio=1mm/33,
+      ratio=1mm/60,
       minimum radius=0mm,
@@
       black/.style={fill/.style={draw=black,line width=0pt,fill=black,},},
-      ice-to-water/.style={fill/.style={fill=none, top color=ice, bottom color=water},},
-      ice-to-water-horiz/.style={fill/.style={fill=none, right color=ice, left color=water},},
-      ice-to-air/.style={fill/.style={fill=none, top color=ice, bottom color=air},},
-      ice-to-air-horiz/.style={fill/.style={fill=none, left color=ice, right color=air},},
-      water-to-air-horiz/.style={fill/.style={fill=none, right color=water, left color=air},},
-      water-to-air/.style={fill/.style={fill=none, top color=water, bottom color=air},},
-      air-to-ice/.style={fill/.style={fill=none, top color=air, bottom color=ice},},
       red/.style={fill/.style={fill=none, top color=red, bottom color=red},},
-      red-to-ice/.style={fill/.style={fill=none, top color=red, bottom color=ice},},
-      water-to-ice/.style={fill/.style={fill=none, top color=water, bottom color=ice},},
-      ice-to-black/.style={fill/.style={fill=none, left color=ice, right color=black},},
+      ice-to-water/.style={fill/.style={fill=none, left color=ice, right color=water},},
+      ice-to-air/.style={fill/.style={fill=none, left color=ice, right color=air},},
+      ice-to-black/.style={fill/.style={fill=none, top color=ice, bottom color=black},},
+      ice-to-air-vert/.style={fill/.style={fill=none, top color=ice, bottom color=air},},
+      ice-to-water-vert/.style={fill/.style={fill=none, bottom color=ice, top color=water},},
+      water-to-air-vert/.style={fill/.style={fill=none, right color=water, left color=air},},
+      water-to-air/.style={fill/.style={fill=none, left color=water, right color=air},},
+      air-to-ice/.style={fill/.style={fill=none, left color=air, right color=ice},},
+      red-to-ice/.style={fill/.style={fill=none, left color=red, right color=ice},},
+      water-to-ice/.style={fill/.style={fill=none, left color=water, right color=ice},},
       black-to-ice/.style={fill/.style={fill=none, bottom color=ice, top color=black},},
@@
         draw,
-        font=\ttfamily,
+        font=\tiny,
         text=debug color,text opacity=1,
@@
 
-    \newcommand\abovelabel[1]{ % valname
-      \node[font=\tiny, rotate=45, anchor=north west]
-      at ([yshift=5mm,xshift=-3mm]#1.center) {\qq{#1}: \Lab{v#1}};
+    \newcommand\labelleft[1]{ % valname
+      \node[font=\tiny, anchor=east]
+      at ([yshift=0mm,xshift=-3mm]#1.center) {\Lab{v#1}: \qq{#1}};
     }
-    \newcommand\belowlabel[2]{ % valname
-      \node[font=\tiny, rotate=45, anchor=north east]
-      at ([yshift=-2mm,xshift=0mm]#1.center) {\Lab{#2}: \qq{#1}};
+    \newcommand\labelright[2]{ % valname
+      \node[font=\tiny, anchor=west]
+      at ([yshift=0mm,xshift=3mm]#1.center) {\qq{#1}: \Lab{#2}};
     }
@@
     \sankeyset{ice} % default color until otherwise specified
-    \newcommand\xoff{2mm}
+    \newcommand\yoff{2mm}
 
     \sankeynode{name=SF, quantity=\Q{vSF}}
-    \sankeynode{name=DP, quantity=\Q{vDP}, at={[xshift=-\xoff]SF.right}, anchor=left}
-    \sankeynode{name=CD, quantity=\Q{vCD}, at={[xshift=-\xoff]DP.right}, anchor=left}
-    \sankeynode{name=RF, quantity=\Q{vRF}, at={[xshift=-\xoff]CD.right}, anchor=left}
+    \sankeynode{name=DP, quantity=\Q{vDP}, at={[yshift=\yoff]SF.north}, anchor=right}
+    \sankeynode{name=CD, quantity=\Q{vCD}, at={[yshift=\yoff]DP.north}, anchor=right}
+    \sankeynode{name=RF, quantity=\Q{vRF}, at={[yshift=\yoff]CD.north}, anchor=right}
 
-    \abovelabel{SF}
-    \abovelabel{DP}
-    \abovelabel{CD}
-    \abovelabel{RF}
+    \labelleft{SF}
+    \labelleft{DP}
+    \labelleft{CD}
+    \labelleft{RF}
 
@@
     \sankeynode{name=f0, quantity=\Q{vSF}+\Q{vDP}+\Q{vCD}+\Q{vRF},
-      at={[yshift=-15mm]SF.left}, anchor=left,
-      forked={\Q{vSF}/SFi, \Q{vDP}/DPi, \Q{vCD}/CDi, \Q{vRF}/RFi}}
+      at={[xshift=15mm, yshift=0mm]SF.right}, anchor=right,
+      forked={\Q{vRF}/RFi, \Q{vCD}/CDi, \Q{vDP}/DPi, \Q{vSF}/SFi}}
 
@@
         \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vFRGAIN},
-          at={[yshift=-15mm]f0.left}, anchor=left,
-          forked={\qq{f0}/f0i, \Q{vFRGAIN}/FG, \Q{vRFZ}/RFZi}}
+          at={[xshift=15mm]f0.right}, anchor=right,
+          forked={\Q{vRFZ}/RFZi, \Q{vFRGAIN}/FG, \qq{f0}/f0i}}
     } {
         \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vML}+\Q{vFRGAIN},
-          at={[yshift=-15mm]f0.left}, anchor=left,
-          forked={\qq{f0}/f0i, \Q{vML}/MLi, \Q{vFRGAIN}/FG, \Q{vRFZ}/RFZi}}
+          at={[xshift=15mm]f0.right}, anchor=right,
+          forked={\Q{vRFZ}/RFZi, \Q{vFRGAIN}/FG, \Q{vML}/MLi, \qq{f0}/f0i}}
     }
@@
     {
-      \sankeyfork{f1}{
-        \Q{vMG}/MG,
-        \Q{vSU}/SU,
-        \Q{vFRLOSS}/FL,
-        \Q{vGZRET}/GZ,
-        \Q{vDYN}/f6, \Q{vBM}/BM, \Q{vRU}+\Q{vEV}/f8, \Q{vRFZ}/RFZo}
-    }
-    {
+      \sankeyfork{f1}{\Q{vRFZ}/RFZo, \Q{vSU}/SU, \Q{vRU}+\Q{vEV}/f8, \Q{vDYN}/f6,
+        \Q{vGZRET}/GZ, \Q{vFRLOSS}/FL, \Q{vBM}/BM, \Q{vMG}/MG}
+    }{
       \sankeyfork{f1}{
         %% \Q{vMG}/MG,
-        \Q{vSU}/SU,
-        \Q{vFRLOSS}/FL,
-        \Q{vGZRET}/GZ,
-        \Q{vDYN}/f6, \Q{vBM}/BM, \Q{vRU}+\Q{vEV}/f8, \Q{vRFZ}/RFZo}
+        \Q{vRFZ}/RFZo, \Q{vSU}/SU, \Q{vRU}+\Q{vEV}/f8, \Q{vDYN}/f6,
+        \Q{vGZRET}/GZ, \Q{vFRLOSS}/FL, \Q{vBM}/BM}
     }
 
-    \sankeydubins[minimum radius=5mm,ice-to-water-horiz]{RFZo}{RFZi}
+    \sankeyturnleft{RFZo}{1}
+    \sankeydubins[minimum radius=5mm,ice-to-water-vert]{RFZo}{RFZi}
 
     \sankeyadvance{f6}{10mm}
-    \sankeyfork{f6}{\Q{vICE}/D, \Q{vSUB}/SUB}
-    \sankeyadvance{D}{10mm}
+    \sankeyfork{f6}{\Q{vICE}/CALV, \Q{vSUB}/SUB}
+    \sankeyadvance[ice-to-water]{SUB}{10mm}
 
-    \sankeynode{name=SUBa, quantity=\qq{SUB}, at={[xshift=-\xoff]D.right}, anchor=left}
-    \sankeyoutin[ice-to-water]{SUB}{SUBa}
+    \sankeynode{name=CALVa, quantity=\qq{CALV}, at={[yshift=\yoff]SUB.left}, anchor=right}
+    \sankeyoutin{CALV}{CALVa}
 
-    \sankeynode{name=BMa, quantity=\qq{BM}, at={[xshift=-\xoff]SUBa.right}, anchor=left}
-    \sankeyoutin[ice-to-water]{BM}{BMa}
-
-    \sankeynode{name=f8a, quantity=\qq{f8}, at={[xshift=-\xoff]BMa.right}, anchor=left}
+    \sankeynode{name=f8a, quantity=\qq{f8}, at={[yshift=\yoff]CALVa.left}, anchor=right}
     \sankeyoutin[ice-to-water]{f8}{f8a}
 
-    \sankeynode{name=GZa, quantity=\qq{GZ}, at={[xshift=\xoff]D.left}, anchor=right}
+    \sankeynode{name=GZa, quantity=\qq{GZ}, at={[yshift=-\yoff]SUB.right}, anchor=left}
     \sankeyoutin{GZ}{GZa}
 
-    \sankeynode{name=FLa, quantity=\qq{FL}, at={[xshift=\xoff]GZa.left}, anchor=right}
+    \sankeynode{name=FLa, quantity=\qq{FL}, at={[yshift=-\yoff]GZa.right}, anchor=left}
     \sankeyoutin{FL}{FLa}
 
-    \sankeynode{name=SUa, quantity=\qq{SU}, at={[xshift=\xoff]FLa.left}, anchor=right}
-    \sankeyoutin{SU}{SUa}
+    \sankeynode{name=BMa, quantity=\qq{BM}, at={[yshift=-\yoff]FLa.right}, anchor=left}
+    \sankeyoutin[ice-to-water]{BM}{BMa}
 
@@
     \sankeyend[out-red]{FLa}
-    \belowlabel{FLa}{vFRLOSS}
+    \labelright{FLa}{vFRLOSS}
 
     \ifnum \numexpr\Q{vMG} > \Q{vML}
-      \sankeynode{name=MGa, quantity=\qq{MG},
-        at={[xshift=3mm,yshift=-3mm]MG.left}, anchor=left, angle=0}
-      \node[font=\tiny, anchor=north]
-      at ([yshift=0mm,xshift=3mm]MGa.south) {\qq{MGa}: Mass gain};
+      \sankeynode{name=MGa, quantity=\qq{MG}, angle=-90,
+        at={[xshift=5mm,yshift=-3mm]MG.right}, anchor=right}
+      \node[font=\tiny, anchor=east]
+      at ([yshift=0mm,xshift=0mm]MGa.right) {Mass gain: \qq{MGa}};
       \sankeyoutin{MG}{MGa}
@@
 
-    \sankeyfork{f8a}{\Q{vRU}/RU, \Q{vEV}/EV}
+    \sankeyfork{f8a}{\Q{vEV}/EV, \Q{vRU}/RU}
     \sankeynode{name=EVa, quantity=\qq{EV},
-      at={[xshift=-3mm,yshift=-5mm]f8a.right}, anchor=right, angle=180}
-    \sankeyoutin[water]{EV}{EVa}
-    \sankeyadvance[water-to-air-horiz]{EVa}{5mm}
+      at={[yshift=2mm,xshift=5mm]f8a.left}, anchor=left, angle=45}
+    \sankeyoutin[water-to-air]{EV}{EVa}
+    \sankeyadvance[air]{EVa}{1mm}
     \sankeyend[air]{EVa}
 
+    \sankeynode{name=SUa, quantity=\qq{SU},
+      at={[xshift=-6mm]EVa.left}, anchor=right, angle=45}
+    %% \sankeyoutin[ice-to-air{SU}{SUa}
+    \sankeydubins[minimum radius=6mm,ice-to-air]{SU}{SUa}
+    \sankeyend[air]{SUa}
+
     \ifthenelse{\Q{vMG} < \Q{vML}} { % mass loss
-      \sankeynode{name=ML, quantity=\Q{vML}, at={[xshift=-\xoff]RF.right}, anchor=left}
-      \abovelabel{ML}
+      \sankeynode{name=ML, quantity=\Q{vML},
+        at={[yshift=6mm]RF.north}, anchor=right, angle=-45}
+      \node[font=\tiny, anchor=east]
+      at ([xshift=-3mm,yshift=2mm]ML.west) {Mass loss: \qq{ML}};
 
@@
       \sankeynode{name=FGi, quantity=\Q{vFRGAIN},
-        at={[yshift=-5mm,xshift=-\xoff]ML.right}, anchor=left}
+        at={[yshift=0mm,xshift=6mm]ML.left}, anchor=right, angle=-45}
 
       \sankeystart[red]{ML}
-      \sankeyadvance[red-to-ice]{ML}{5mm}
-      \sankeyoutin{ML}{MLi}
+      \sankeyadvance[red]{ML}{5mm}
+      \sankeyoutin[red-to-ice]{ML}{MLi}
     }{ % mass gain
@@
       \sankeynode{name=FGi, quantity=\Q{vFRGAIN},
-        at={[yshift=15mm,xshift=3mm]RFZi.right}, anchor=left}
+        at={[xshift=-15mm,yshift=8mm]RFZi.right}, anchor=left, angle=-45}
     }
@@
     \sankeyoutin[in-black]{FGi}{FG}
-    \node[font=\tiny, rotate=45, anchor=north east]
-    at ([yshift=12mm,xshift=3mm]FGi.right) {Frontal advance: \qq{FGi}};
+    \node[font=\tiny, anchor=east]
+    at ([xshift=-3mm,yshift=3mm]FGi.west) {Frontal advance: \qq{FGi}};
     \fi
-
-    \sankeynode{name=SUaa, quantity=\qq{SU},
-      at={[xshift=2mm,yshift=-5mm]SUa.left}, anchor=left, angle=0}
-    \sankeyoutin[ice]{SUa}{SUaa}
-    \sankeyadvance[ice-to-air-horiz]{SUaa}{2mm}
-    \sankeyend[air]{SUaa}
 
@@
     \sankeyend[water,out-red]{GZa}
-    \sankeyadvance{D}{10mm}
-    \sankeyend{D}
+    \sankeyadvance{CALVa}{10mm}
+    \sankeyend{CALVa}
 
     \sankeyset{water}
+    \sankeyadvance{RU}{10mm}; \sankeyend{RU}
     \sankeyadvance{BMa}{10mm}; \sankeyend{BMa}
-    \sankeyadvance{SUBa}{10mm}; \sankeyend{SUBa}
-    \sankeyadvance{RU}{10mm}; \sankeyend{RU}
+    \sankeyadvance{SUB}{10mm}; \sankeyend{SUB}
 
-    \belowlabel{RU}{vRU}
-    \belowlabel{BMa}{vBM}
-    \belowlabel{SUBa}{vSUB}
-    \belowlabel{D}{vICE}
-    \belowlabel{GZa}{vGZRET}
+    \labelright{RU}{vRU}
+    \labelright{BMa}{vBM}
+    \labelright{SUB}{vSUB}
+    \labelright{CALVa}{vICE}
+    \labelright{GZa}{vGZRET}
 
-    \node[font=\tiny, rotate=45, anchor=north east]
-    at ([yshift=1mm,xshift=3mm]SUaa.south) {\Lab{vSU}: \qq{SUaa}};
+    \node[font=\tiny, anchor=west]
+    at ([yshift=4mm,xshift=\yoff]SUa.west) {\qq{SUa}: \Lab{vSU}};
 
-    \node[font=\tiny, rotate=45, anchor=north east]
-    at ([yshift=0mm,xshift=5mm]EVa.north) {\Lab{vEV}: \qq{EVa}};
+    \node[font=\tiny, anchor=west]
+    at ([yshift=0mm,xshift=\yoff]EVa.west) {\qq{EVa}: \Lab{vEV}};
 
-    \node[font=\tiny, anchor=west, rotate=90]
+    \node[font=\tiny, anchor=east]
     at ([yshift=-0.5mm,xshift=1mm]RFZo.center) {\Lab{vRFZ}: \qq{RFZi}};

### 2024-07-24 sankey.tex
Mass gain not at 90 degrees

@@
       ice-to-air/.style={fill/.style={fill=none, left color=ice, right color=air},},
-      ice-to-black/.style={fill/.style={fill=none, top color=ice, bottom color=black},},
+      ice-to-black/.style={fill/.style={fill=none, left color=ice, right color=black},},
       ice-to-air-vert/.style={fill/.style={fill=none, top color=ice, bottom color=air},},
@@
     \ifnum \numexpr\Q{vMG} > \Q{vML}
-      \sankeynode{name=MGa, quantity=\qq{MG}, angle=-90,
+      \sankeynode{name=MGa, quantity=\qq{MG}, angle=-35,
         at={[xshift=5mm,yshift=-3mm]MG.right}, anchor=right}
@@
       at ([yshift=0mm,xshift=0mm]MGa.right) {Mass gain: \qq{MGa}};
-      \sankeyoutin{MG}{MGa}
+      \sankeyoutin[ice-to-black]{MG}{MGa}
     \fi
@@
     }{ % mass gain
-      \sankeyadvance[ice-to-black]{MGa}{3mm}
+      \sankeyadvance[black]{MGa}{3mm}
       \sankeyend[black]{MGa}

### 2024-07-28 ms.tex
Writing

@@
 
-  Abstract 1:
-
-  We combine recent estimates of ice mass change that include ice at or near flotation (i.e., terminus change in Greenland) with grounded ice mass balance to produce the first total mass balance product for the Greenlandic and Antarctic ice sheets.
-Greenland mass loss is 30 \% higher than previous estimates due to terminus retreat.
-This additional mass loss does not contribute significantly to sea level rise as it is already at or near flotation.
-
-  Abstract 2:
-
   Understanding the processes involved in ice sheet mass loss requires knowledge production and sharing, in addition to field observations and computer modeling.
 Tabular data is optimized for computers and useful for humans, but graphical presentations can provide significantly higher information density.
-Here we present several Sankey diagrams depicting all mass flow of water (in various phases) in Greenland and Antarctica.
-These diagrams show mass loss in Greenland is 330 Gt yr$^{-1}$  which is 30 \% higher previous estimates that neglect terminus retreat of ice at or near flotation.
+Here we present several Sankey diagrams depicting all mass flow of water in various phases in Greenland and Antarctica.
+Sankey diagrams help bring perspective to the interplay and relative importance of snowfall, rain, drawdown, refreezing, submarine melt, iceberg calving, and grounding line retreat in contributing to ice sheet mass balance.
 
+  This work quantifies both grounded and floating ice mass loss.
+From this, mass loss in Greenland is 330 Gt yr$^{-1}$ which is \(\sim\)25 \% higher than the \(\sim\)270 Gt yr$^{-1}$ estimates that neglect terminus retreat of ice at or near flotation.
+Mass loss in Antarctica is 275 Gt yr$^{-1}$ which is \(\sim\)85 \% higher than the \(\sim\)150 Gt yr$^{-1}$ estimates of grounded ice mass loss.
+
+  This work also reports gross not net values, which is useful for example for quantifying total freshwater export from an ice sheet.
+For example, freshwater export due to frontal retreat is not 60 Gt yr$^{-1}$, but rather nearly six times larger at 345 Gt yr$^{-1}$ due to 345 Gt yr$^{-1}$ of frontal retreat offset by 285 Gt yr$^{-1}$ of frontal advance, with the difference as the commonly reported value of 60 Gt yr$^{-1}$.
 \end{abstract}
@@
 
-Mass flows of ice sheets, and subsequent mass gains and losses, is an of-studied topic in the domain of glaciology.
-However, few studies consider all processes and their relative magnitudes and uncertainties - this is typically limited to review papers, which may not focus on quantitative assessment of each process.
+Mass loss from the grounded portion of the Greenlandic and Antarctic ice sheets contributes to sea level rise (SLR), and can be captured by a single value per ice sheet.
+Reporting a single value is beneficial because it simplifies interpretation and comparison.
+However, ice sheets are complex systems and while their contribution to sea level rise is important, it is not a holistic view and does not capture the general health and state of the ice sheet.
+Few studies consider all processes and their relative magnitudes and uncertainties - this is typically limited to review papers, which may not focus on quantitative assessment of each process.
+As one example, an ice sheet that increases output via discharge or submarine melting by X \% but has that offset by an equal increase in snowfall would report no net mass change or net SLR contribution, but has entered a different state when viewing constituent terms or gross values.
 
@@
 \begin{figure*}
-\centering{\includegraphics[width=0.85\textwidth]{fig.png}}
+\centering{\includegraphics[width=0.85\textwidth]{fig_aq_gl.png}}
 \caption{Sankey mass flow diagrams for Antarctica and Greenland, and Antarctica split into East, West, and Peninsula.
@@
 
-  The flows can be somewhat tied to physical space as in the Minard map mentioned above, but are not done so explicitly here.
-An similar display to the diagrams presented here, of glacier mass flow by \citet[Fig.
-2]{cogley_2011}, shows flow overlaid an a glacier schematic, but with neither flow magnitude nor proportionality.
-Nonetheless, there is an implicit physical connection here for most flow.
-Inputs all come from the atmosphere (with the exception of drawdown and frontal advance), and outputs pointing down are all mass flow into the ocean.
+A similar display to the diagrams presented here by \citet[Fig.
+2]{cogley_2011} shows flows overlaid an a glacier schematic and here we build on that work by adding magnitude of processes and making the graphics proportional to magnitudes.
 
@@
 
-We use existing published values, some new estimates, and some derived values (See Tables \ref{tab:gl} and \ref{tab:aq} in Appendix A).
+We use a combination of existing published values, new estimates, and derived values.
+
+Existing published values include
 
@@
 This term balances all the other terms, so that the Sankey diagram has outputs balancing inputs.
-
-Here we describe our new estimates.
 
@@
 \ref{fig}.
+
+\subsection{Sankey diagrams}
+
+The Sankey diagrams shown here are generated from a script that combines a CSV file of values with a \LaTeX\ template that uses the TikZ Sankey package \citep{sankey}.
+This architecture makes it trivial to generate similar diagrams for other time periods, differences between time periods, after generating CSV files with representative values.
 
@@
 
+\section{Appendix B}
+\label{sec:appendix:B}
+
+\begin{figure*}
+\centering{\includegraphics[width=0.85\textwidth]{fig_aq_parts.png}}
+\caption{Sankey mass flow diagrams for Antarctica and Greenland, and Antarctica split into East, West, and Peninsula.
+All widths are proportional within and between images.
+Because Sankey diagrams balance all inputs and outputs, mass losses require a `drawdown' input (red) to balance the larger outputs, and mass gains requires an `accumulation' output (black) to balance the larger inputs.}
+\label{fig}
+\end{figure*}
+
 \end{document}

### 2024-07-30 sankey.tex
Ice shelf submarine freeze/melt separated into gross not net terms
Uses Paolo (2023) http://doi.org/10.5194/tc-17-3409-2023 rather than Davison (2023) http://doi.org/10.1126/sciadv.adi0186

@@
           (\Q{vSU}+\Q{vRU}+\Q{vBM}+\Q{vDYN}+\Q{vFRLOSS}+\Q{vGZRET}+\Q{vEV})-
-          (\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP}+\Q{vFRGAIN})},
+          (\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP}+\Q{vFRGAIN}+\Q{vSUBF})},
         {vMG/Mass gain/
-          (\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP}+\Q{vFRGAIN})-
+          (\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP}+\Q{vFRGAIN}+\Q{vSUBF})-
           (\Q{vSU}+\Q{vRU}+\Q{vBM}+\Q{vDYN}+\Q{vFRLOSS}+\Q{vGZRET}+\Q{vEV})}
@@
 
+    \sankeynode{name=SUBF, quantity=\Q{vSUBF},
+      at={[yshift=-\yoff]SF.south}, anchor=left}
+
     \ifthenelse{\Q{vMG} > \Q{vML}} {
-        \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vFRGAIN},
-          at={[xshift=15mm]f0.right}, anchor=right,
-          forked={\Q{vRFZ}/RFZi, \Q{vFRGAIN}/FG, \qq{f0}/f0i}}
+      \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vFRGAIN}+\Q{vSUBF},
+        at={[xshift=30mm, yshift=\yoff]SUBF.south}, anchor=right,
+        forked={\Q{vRFZ}/RFZi, \Q{vFRGAIN}/FG, \qq{f0}/f0i, \Q{vSUBF}/SUBFi}}
     } {
-        \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vML}+\Q{vFRGAIN},
-          at={[xshift=15mm]f0.right}, anchor=right,
-          forked={\Q{vRFZ}/RFZi, \Q{vFRGAIN}/FG, \Q{vML}/MLi, \qq{f0}/f0i}}
+      \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vML}+\Q{vFRGAIN}+\Q{vSUBF},
+        at={[xshift=30mm, yshift=\yoff]SUBF.south}, anchor=right,
+        forked={\Q{vRFZ}/RFZi, \Q{vFRGAIN}/FG, \Q{vML}/MLi, \qq{f0}/f0i, \Q{vSUBF}/SUBFi}}
     }
@@
     \sankeyadvance{f1}{15mm}
+
+    \ifnum \numexpr\Q{vSUBF} > 0
+    \labelleft{SUBF}
+    \sankeystart[water]{SUBF}
+    \sankeydubins[water-to-ice]{SUBF}{SUBFi}
+    \fi
 

### 2024-08-18 sankey_gl.tex
Major refactor and update

\documentclass[11pt]{article}
\usepackage{sankey}
\usepackage{tikz}
\usepackage{pgf}
\usepackage{ifthen}
\pagestyle{empty}

\renewcommand{\familydefault}{\sfdefault}

\begin{document}

\selectcolormodel{cmyk} % Can be rbg or cmyk, cmyk is better for printing

\begin{tikzpicture}[remember picture, overlay, shift={(5,-5)}]

  \begin{sankeydiagram}
    %% \sankeyset{debug=true}

    \colorlet{water}[rgb]{cyan!75!blue!50!white}
    \colorlet{ice}[rgb]{gray!20!gray!20!gray!20}
    %% \colorlet{air}[rgb]{gray!1!gray!1!gray!1}
    \colorlet{air}{yellow!50}
    \colorlet{red}[rgb]{red!50!orange!75!white}
    \colorlet{black}[rgb]{black!100!black!100!black!100}
    \colorlet{gray}[rgb]{black!30!black!30!black!30}

    \sankeyset{
      %% comment for gray outline
      %% draw/.style={
      %%   draw=white,
      %%   line width=0pt,line cap=round,line join=round,
      %% },
      start style=arrow,end style=arrow,
      %% fill/.style={fill=yellow!10,line width=0pt,draw=yellow!10},
      ratio=1mm/60,
      minimum radius=0mm,
      water/.style={fill/.style={draw=water,line width=0pt,fill=water,},},
      ice/.style={fill/.style={draw=ice,line width=0pt,fill=ice,},},
      air/.style={fill/.style={draw=air,line width=0pt,fill=air,},},
      black/.style={fill/.style={draw=black,line width=0pt,fill=black,},},
      red/.style={fill/.style={fill=none, top color=red, bottom color=red},},
      ice-to-water/.style={fill/.style={fill=none, left color=ice, right color=water},},
      ice-to-air/.style={fill/.style={fill=none, left color=ice, right color=air},},
      ice-to-black/.style={fill/.style={fill=none, left color=ice, right color=black},},
      ice-to-air-vert/.style={fill/.style={fill=none, top color=ice, bottom color=air},},
      ice-to-water-vert/.style={fill/.style={fill=none, bottom color=ice, top color=water},},
      water-to-air-vert/.style={fill/.style={fill=none, right color=water, left color=air},},
      water-to-air/.style={fill/.style={fill=none, left color=water, right color=air},},
      air-to-ice/.style={fill/.style={fill=none, left color=air, right color=ice},},
      red-to-ice/.style={fill/.style={fill=none, left color=red, right color=ice},},
      water-to-ice/.style={fill/.style={fill=none, left color=water, right color=ice},},
      black-to-ice/.style={fill/.style={fill=none, bottom color=ice, top color=black},},
      in-black/.style={draw/.style={draw=gray,thick=3},},
      out-red/.style={draw/.style={draw=red,thick=3},},
      out-blue/.style={draw/.style={draw=water,thick=3},},
      debug label/.style={
        overlay,
        draw,
        font=\tiny,
        text=debug color,text opacity=1,
        inner sep=.1em,
        fill=white,fill opacity=0.5,
        rounded corners=.1em,
        %% node contents={\name: \qty},
        node contents={\qty},
        %% node contents={\name},
      },
    }

    % storage of labels
    \newcommand\LabSet[2]{% node name, label
      \expandafter\edef\csname#1@Lab\endcsname{#2}}
    \newcommand\Lab[1]{% node name
      \csname#1@Lab\endcsname}
    % storage of quantities
    \newcommand\QSet[2]{% node name, quantity
      \expandafter\edef\csname#1@Q\endcsname{\fpeval{#2}}}
    \newcommand\Q[1]{% node name
      \csname#1@Q\endcsname}
    % all nodes with their name, label and quantity
    \newcommand\qq[1]{\sankeygetnodeqty{#1}}

    \newcommand\labelleft[1]{ % valname
      \node[font=\tiny, anchor=east]
      at ([yshift=0mm,xshift=-3mm]#1.center) {\Lab{v#1}: \qq{#1}};
    }
    \newcommand\labelright[2]{ % valname
      \node[font=\tiny, anchor=west]
      at ([yshift=0mm,xshift=3mm]#1.center) {\qq{#1}: \Lab{#2}};
    }
    \newcommand\labelbold[2]{ % valname
      \node[font=\tiny, anchor=west]
      at ([yshift=0mm,xshift=3mm]#1.center) {\textbf{\qq{#1}: \Lab{#2}}};
    }

    \sankeyset{
      def data/.code args={#1/#2/#3}{% node name/label/values
        \LabSet{#1}{#2}
        \QSet{#1}{#3}
        \typeout{#1: \Q{#1} Gt (\Lab{#1})}
      },
      def data/.list={
        %% INSERT_CSV_HERE
        %% COMPUTED
        {vSMBin/SMB Inputs/\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP}},
        {vINPUT/Inputs/\Q{vSMBin}+\Q{vFRONTADV}+\Q{vSHELFFREEZE}},
        {vSMBout/SMB Outputs/\Q{vSU}+\Q{vRU}+\Q{vEV}},
        {vDYN/Dynamics/\Q{vCALV}+\Q{vFRONTMELT}+\Q{vSHELFMELT}+\Q{vFRONTRET}+\Q{vGZRET}+\Q{vBMB}},
        {vOUTPUT/Outputs/\Q{vSMBout}+\Q{vDYN}},
        {vSMB/SMB/\Q{vSMBin}-\Q{vSMBout}},
        {vML/Mass loss/\Q{vOUTPUT}-\Q{vINPUT}},
        {vMG/Mass gain/\Q{vINPUT}-\Q{vOUTPUT}}
      },
    }

    \sankeyset{ice} % default color until otherwise specified
    \newcommand\yoff{3mm}

    \sankeynode{name=SF, quantity=\Q{vSF}}
    \sankeynode{name=DP, quantity=\Q{vDP}, at={[yshift=\yoff]SF.north}, anchor=right}
    \sankeynode{name=CD, quantity=\Q{vCD}, at={[yshift=\yoff]DP.north}, anchor=right}
    \sankeynode{name=RF, quantity=\Q{vRF}, at={[yshift=\yoff]CD.north}, anchor=right}

    \sankeystart{SF}; \labelleft{SF}
    \sankeystart[air]{DP}; \labelleft{DP}
    \sankeystart{CD}; \labelleft{CD}
    \sankeystart[water]{RF}; \labelleft{RF}

    \sankeynode{name=SMBIN, quantity=\Q{vSF}+\Q{vDP}+\Q{vCD}+\Q{vRF},
      at={[xshift=15mm, yshift=0mm]SF.right}, anchor=right,
      forked={\Q{vRF}/RFi, \Q{vCD}/CDi, \Q{vDP}/DPi, \Q{vSF}/SFi}}

    \sankeyoutin{SF}{SFi}
    \sankeydubins[minimum radius=3mm,air-to-ice]{DP}{DPi}
    \sankeydubins[minimum radius=3mm]{CD}{CDi}
    \sankeydubins[minimum radius=3mm,water-to-ice]{RF}{RFi}

    \sankeynode{name=SHELFFREEZE, quantity=\Q{vSHELFFREEZE},
      at={[yshift=-\yoff]SF.south}, anchor=left}

    \sankeynode{name=f1, at={[xshift=30mm, yshift=\yoff]SHELFFREEZE.south}, anchor=right,
      quantity=\qq{SMBIN}+\Q{vFRONTADV}+\Q{vSHELFFREEZE}+\Q{vML}+\Q{vRFZ},
      forked={\Q{vRFZ}/RFZi, \Q{vFRONTADV}/FG, \Q{vML}/MLi,
        \qq{SMBIN}/SMBINi, \Q{vSHELFFREEZE}/SHELFFREEZEi}}

    \sankeyoutin{SMBIN}{SMBINi}
    \sankeyadvance{f1}{15mm}

    \ifnum \numexpr\Q{vSHELFFREEZE} > 0
    \labelleft{SHELFFREEZE}
    \sankeystart[water]{SHELFFREEZE}
    \sankeydubins[water-to-ice]{SHELFFREEZE}{SHELFFREEZEi}
    \fi

    \sankeyfork{f1}{\Q{vRFZ}/RFZo, \Q{vSU}/SU, \Q{vRU}+\Q{vEV}/RUEV,
      \Q{vCALV}+\Q{vFRONTMELT}/DISCH, \Q{vSHELFMELT}/SHELFMELT,
      \Q{vFRONTRET}/FRONTRET, \Q{vGZRET}/GZ, \Q{vBMB}/BMB}

    \sankeyturnleft{RFZo}{1}
    \sankeydubins[minimum radius=5mm,ice-to-water-vert]{RFZo}{RFZi}
    \node[font=\tiny, anchor=east]
    at ([yshift=+1.5mm,xshift=1mm]RFZo.north) {\Lab{vRFZ}: \qq{RFZi}};

    \sankeyadvance{DISCH}{10mm}
    \sankeyfork{DISCH}{\Q{vCALV}/CALV, \Q{vFRONTMELT}/FRONTMELT}

    \sankeynode{name=SHELFMELTa,quantity=\qq{SHELFMELT}, at={[yshift=-\yoff, xshift=10mm]FRONTMELT.right}}
    \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET}, at={[yshift=-\yoff]SHELFMELTa.right}}
    \sankeyadvance[ice-to-water]{FRONTMELT}{10mm}
    \sankeynode{name=CALVa, quantity=\qq{CALV}, at={[yshift=\yoff]FRONTMELT.left}, anchor=right}

    \sankeynode{name=GZa, quantity=\qq{GZ}, at={[yshift=-\yoff]FRONTRETa.right}, anchor=left}
    \sankeynode{name=BMBa, quantity=\qq{BMB}, at={[yshift=-\yoff]GZa.right}, anchor=left}

    \sankeyoutin{CALV}{CALVa};
    \sankeyoutin[ice-to-water]{SHELFMELT}{SHELFMELTa}
    \sankeyoutin[out-blue]{FRONTRET}{FRONTRETa};
    \sankeyoutin[ice-to-water]{BMB}{BMBa}
    \sankeyoutin[ice-to-water]{GZ}{GZa}

    %% Finish dynamic outputs
    \sankeyadvance{CALVa}{10mm}; \sankeyend{CALVa}; \labelright{CALVa}{vCALV}
    \sankeyadvance[water]{FRONTMELT}{10mm}; \sankeyend[water]{FRONTMELT}; \labelright{FRONTMELT}{vFRONTMELT}
    \sankeyadvance[water]{SHELFMELTa}{10mm}; \sankeyend[water]{SHELFMELTa}; \labelright{SHELFMELTa}{vSHELFMELT}
    \sankeyadvance[out-blue]{FRONTRETa}{10mm}; \sankeyend[out-blue]{FRONTRETa}; \labelbold{FRONTRETa}{vFRONTRET}
    \sankeyadvance[water]{GZa}{10mm}; \sankeyend[water]{GZa}; \labelbold{GZa}{vGZRET}
    \sankeyadvance[water]{BMBa}{10mm}; \sankeyend[water]{BMBa}; \labelright{BMBa}{vBMB}

    %% SMB outputs
    \sankeynode{name=RUEVa, quantity=\qq{RUEV}, at={[yshift=\yoff, xshift=-10mm]CALVa.left}, anchor=right}
    \sankeyoutin[ice-to-water]{RUEV}{RUEVa}
    \sankeyfork{RUEVa}{\Q{vEV}/EV, \Q{vRU}/RU}
    \sankeynode{name=EVa, quantity=\qq{EV},
      at={[yshift=\yoff,xshift=5mm]RUEVa.left}, anchor=left, angle=45}
    \sankeyoutin[water-to-air]{EV}{EVa}; \sankeyadvance[air]{EVa}{1mm}
    \sankeyend[air]{EVa};
    \node[font=\tiny, anchor=west] at ([yshift=0mm,xshift=\yoff]EVa.west) {\qq{EVa}: \Lab{vEV}};

    \sankeyadvance[water]{RU}{10mm}; \sankeyend[water]{RU}; \labelright{RU}{vRU}

    \sankeynode{name=SUa, quantity=\qq{SU}, at={[xshift=-6mm]EVa.left}, anchor=right, angle=45}
    \sankeydubins[minimum radius=6mm,ice-to-air]{SU}{SUa}; \sankeyend[air]{SUa}
    \node[font=\tiny, anchor=west] at ([yshift=4mm,xshift=\yoff]SUa.west) {\qq{SUa}: \Lab{vSU}};

    \sankeynode{name=ML, quantity=\Q{vML},
      at={[yshift=6mm]RF.north}, anchor=right, angle=-45}
    \node[font=\tiny, anchor=east]
    at ([xshift=-3mm,yshift=2mm]ML.west) {\textbf{Glacier drawdown: \qq{ML}}};

    % frontal advance next to drawdown (mass loss)
    \sankeynode{name=FRONTADVi, quantity=\Q{vFRONTADV},
      at={[yshift=0mm,xshift=6mm]ML.left}, anchor=right, angle=-45}

    \sankeystart[red]{ML}
    \sankeyadvance[red]{ML}{5mm}
    \sankeyoutin[red-to-ice]{ML}{MLi}

    \ifnum \numexpr\qq{FRONTADVi} > 0
    \sankeystart[in-black]{FRONTADVi}
    \sankeyoutin[in-black]{FRONTADVi}{FG}
    \node[font=\tiny, anchor=east]
    at ([xshift=-3mm,yshift=3mm]FRONTADVi.west) {Frontal advance: \qq{FRONTADVi}};
    \fi

    %% % fancy labels
    %% \sankeynode[debug]{name=Discharge, at={[xshift=-10mm]DISCH.north}, anchor=left,
    %%   quantity=\Q{vDISCH}}
    %% \sankeynode[debug]{name=SMB Inputs, at={[xshift=3mm]SMBIN.north}, anchor=left,
    %%   quantity=\Q{vSMBin}}
    %% \sankeynode[debug]{name=SMB Outputs, at={[xshift=0mm]RFZo.south}, anchor=left,
    %%   quantity=\Q{vSMBout}}

  \end{sankeydiagram}
\end{tikzpicture}
\end{document}

### 2025-03-10 sankey.tex
Adjustments and rebalance
Exclude ISLANDS from Calving (75 from Rignot 2019)
Fix runoff for AQ regions
Fix basal melting for AQ regions
refreezing text
co-author edits and comments incorporated

--- 2024-07-24 sankey.tex
@@
       %% comment for gray outline
+      %% draw/.style={
+      %%   draw=white,
+      %%   line width=0pt,line cap=round,line join=round,
+      %% },
+      %% comment for gray outline
       draw/.style={
-        draw=white,
-        line width=0pt,line cap=round,line join=round,
+        draw=black,
+        line width=0.1mm,line cap=round,line join=round,
       },
@@
       out-red/.style={draw/.style={draw=red,thick=3},},
+      out-blue/.style={draw/.style={draw=water,thick=3},},
       debug label/.style={
@@
         rounded corners=.1em,
-        %% node contents={\name (\qty)},
-        node contents={\name},
+        %% node contents={\name: \qty},
+        node contents={\qty},
+        %% node contents={\name},
       },
@@
     }
+    \newcommand\labelbold[2]{ % valname
+      \node[font=\tiny, anchor=west]
+      at ([yshift=0mm,xshift=3mm]#1.center) {\textbf{\qq{#1}: \Lab{#2}}};
+    }
 
@@
         %% COMPUTED
-        {vML/Mass loss/
-          (\Q{vSU}+\Q{vRU}+\Q{vBM}+\Q{vDYN}+\Q{vFRLOSS}+\Q{vGZRET}+\Q{vEV})-
-          (\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP}+\Q{vFRGAIN})},
-        {vMG/Mass gain/
-          (\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP}+\Q{vFRGAIN})-
-          (\Q{vSU}+\Q{vRU}+\Q{vBM}+\Q{vDYN}+\Q{vFRLOSS}+\Q{vGZRET}+\Q{vEV})}
+        {vSMBin/SMB Inputs/\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP}},
+        {vINPUT/Inputs/\Q{vSMBin}+\Q{vFRONTADV}+\Q{vSHELFFREEZE}},
+        {vSMBout/SMB Outputs/\Q{vSU}+\Q{vRU}+\Q{vEV}},
+        {vDYN/Dynamics/\Q{vCALV}+\Q{vFRONTMELT}+\Q{vSHELFMELT}+\Q{vFRONTRET}+\Q{vGZRET}+\Q{vBMB}},
+        {vOUTPUT/Outputs/\Q{vSMBout}+\Q{vDYN}},
+        {vSMB/SMB/\Q{vSMBin}-\Q{vSMBout}},
+        {vML/Drawdown; Retreat/\Q{vOUTPUT}-\Q{vINPUT}},
+        {vMG/Build-up; Expansion/\Q{vINPUT}-\Q{vOUTPUT}}
       },
@@
     \sankeyset{ice} % default color until otherwise specified
-    \newcommand\yoff{2mm}
+    \newcommand\yoff{3mm}
 
@@
 
-    \labelleft{SF}
-    \labelleft{DP}
-    \labelleft{CD}
-    \labelleft{RF}
+    \sankeystart{SF}; \labelleft{SF}
+    \sankeystart[air]{DP}; \labelleft{DP}
+    \sankeystart{CD}; \labelleft{CD}
+    \sankeystart[water]{RF}; \labelleft{RF}
 
-    \sankeystart{SF}
-    \sankeystart[air]{DP}
-    \sankeystart{CD}
-    \sankeystart[water]{RF}
-
-    \sankeynode{name=f0, quantity=\Q{vSF}+\Q{vDP}+\Q{vCD}+\Q{vRF},
+    \sankeynode{name=SMBIN, quantity=\Q{vSF}+\Q{vDP}+\Q{vCD}+\Q{vRF},
       at={[xshift=15mm, yshift=0mm]SF.right}, anchor=right,
@@
 
-    \ifthenelse{\Q{vMG} > \Q{vML}} {
-        \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vFRGAIN},
-          at={[xshift=15mm]f0.right}, anchor=right,
-          forked={\Q{vRFZ}/RFZi, \Q{vFRGAIN}/FG, \qq{f0}/f0i}}
-    } {
-        \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vML}+\Q{vFRGAIN},
-          at={[xshift=15mm]f0.right}, anchor=right,
-          forked={\Q{vRFZ}/RFZi, \Q{vFRGAIN}/FG, \Q{vML}/MLi, \qq{f0}/f0i}}
+    \sankeynode{name=SHELFFREEZE, quantity=\Q{vSHELFFREEZE},
+      at={[yshift=-\yoff]SF.south}, anchor=left}
+
+    \ifthenelse{\Q{vML} > \Q{vMG}} { % mass loss
+      \sankeynode{name=f1, at={[xshift=30mm, yshift=\yoff]SHELFFREEZE.south}, anchor=right,
+        quantity=\qq{SMBIN}+\Q{vFRONTADV}+\Q{vSHELFFREEZE}+\Q{vML}+\Q{vRFZ},
+        forked={\Q{vRFZ}/RFZi, \Q{vFRONTADV}/FG, \Q{vML}/MLi,
+          \qq{SMBIN}/SMBINi, \Q{vSHELFFREEZE}/SHELFFREEZEi}}
+    } { % mass gain edge case
+      \sankeynode{name=f1, at={[xshift=30mm, yshift=\yoff]SHELFFREEZE.south}, anchor=right,
+        quantity=\qq{SMBIN}+\Q{vFRONTADV}+\Q{vSHELFFREEZE}+\Q{vRFZ},
+        forked={\Q{vRFZ}/RFZi, \Q{vFRONTADV}/FG, \qq{SMBIN}/SMBINi, \Q{vSHELFFREEZE}/SHELFFREEZEi}}
     }
 
-    \sankeyoutin{f0}{f0i}
+    \sankeyoutin{SMBIN}{SMBINi}
     \sankeyadvance{f1}{15mm}
 
-    \ifthenelse{\Q{vMG} > \Q{vML}}
-    {
-      \sankeyfork{f1}{\Q{vRFZ}/RFZo, \Q{vSU}/SU, \Q{vRU}+\Q{vEV}/f8, \Q{vDYN}/f6,
-        \Q{vGZRET}/GZ, \Q{vFRLOSS}/FL, \Q{vBM}/BM, \Q{vMG}/MG}
-    }{
-      \sankeyfork{f1}{
-        %% \Q{vMG}/MG,
-        \Q{vRFZ}/RFZo, \Q{vSU}/SU, \Q{vRU}+\Q{vEV}/f8, \Q{vDYN}/f6,
-        \Q{vGZRET}/GZ, \Q{vFRLOSS}/FL, \Q{vBM}/BM}
+    \ifnum \numexpr\Q{vSHELFFREEZE} > 0
+    \labelleft{SHELFFREEZE}
+    \sankeystart[water]{SHELFFREEZE}
+    \sankeydubins[water-to-ice]{SHELFFREEZE}{SHELFFREEZEi}
+    \fi
+
+    \ifthenelse{\Q{vML} > \Q{vMG}} { % mass loss
+      \sankeyfork{f1}{\Q{vRFZ}/RFZo, \Q{vSU}/SU, \Q{vRU}+\Q{vEV}/RUEV,
+        \Q{vCALV}/CALV, \Q{vFRONTMELT}/FRONTMELT, \Q{vSHELFMELT}/SHELFMELT,
+        \Q{vFRONTRET}/FRONTRET, \Q{vGZRET}/GZ, \Q{vBMB}/BMB}
+    }{ % mass gain edge case (AQ_east)
+      \sankeyfork{f1}{\Q{vRFZ}/RFZo, \Q{vSU}/SU, \Q{vRU}+\Q{vEV}/RUEV,
+        \Q{vCALV}/CALV, \Q{vFRONTMELT}/FRONTMELT, \Q{vSHELFMELT}/SHELFMELT,
+        \Q{vFRONTRET}/FRONTRET, \Q{vGZRET}/GZ, \Q{vBMB}/BMB, \Q{vMG}/MG}
+    }
+    \sankeyturnleft{RFZo}{1}
+    \sankeydubins[minimum radius=5mm,ice-to-water-vert]{RFZo}{RFZi}
+    \node[font=\tiny, anchor=east]
+           at ([yshift=+1.5mm,xshift=1mm]RFZo.north) {\Lab{vRFZ}: \qq{RFZi}};
+
+    \ifthenelse{\Q{vFRONTMELT} > \Q{vSHELFMELT}} { % Greenland
+      %% Discharge splits to FRONTMELT (straight) & CALV (above).
+      %% SHELFMELT below FRONTMELT. FRONTRET below SHELFMELT.
+      \sankeyadvance{FRONTMELT}{20mm}
+
+      \sankeynode{name=SHELFMELTa,quantity=\qq{SHELFMELT},
+        at={[yshift=-\yoff, xshift=10mm]FRONTMELT.right}}
+      \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET},
+        at={[yshift=-\yoff]SHELFMELTa.right}}
+      \sankeyadvance[ice-to-water]{FRONTMELT}{10mm}
+
+      \sankeynode{name=CALVa, quantity=\qq{CALV},
+        at={[yshift=\yoff,xshift=0mm]FRONTMELT.left}, anchor=right}
+
+      \sankeyset{water}
+      \sankeyadvance{FRONTMELT}{10mm}; \sankeyend{FRONTMELT}
+      \labelright{FRONTMELT}{vFRONTMELT}
+    } {
+      % Antarctica
+      %% Discharge splits to SHELFMELT (straight) & CALV (above).
+      %% No FRONTMELT. Then FRONTRET
+      \sankeyadvance{SHELFMELT}{10mm}
+      \sankeynode{name=SHELFMELTa, quantity=\qq{SHELFMELT},
+        at={[yshift=0, xshift=20mm]SHELFMELT.right}, anchor=right}
+
+      \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET},
+        at={[yshift=-\yoff]SHELFMELT.right}}
+      \sankeynode{name=CALVa, quantity=\qq{CALV},
+        at={[yshift=\yoff,xshift=0mm]SHELFMELTa.left}, anchor=right}
+      \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET},
+        at={[yshift=-\yoff]SHELFMELTa.right}, anchor=left}
     }
 
-    \sankeyturnleft{RFZo}{1}
-    \sankeydubins[minimum radius=5mm,ice-to-water-vert]{RFZo}{RFZi}
+    \sankeyset{ice}
 
-    \sankeyadvance{f6}{10mm}
-    \sankeyfork{f6}{\Q{vICE}/CALV, \Q{vSUB}/SUB}
-    \sankeyadvance[ice-to-water]{SUB}{10mm}
+    \sankeyoutin[ice-to-water]{SHELFMELT}{SHELFMELTa}
+    \sankeyadvance[water]{SHELFMELTa}{10mm}; \sankeyend[water]{SHELFMELTa}
+    \labelright{SHELFMELTa}{vSHELFMELT}
 
-    \sankeynode{name=CALVa, quantity=\qq{CALV}, at={[yshift=\yoff]SUB.left}, anchor=right}
-    \sankeyoutin{CALV}{CALVa}
+    %% \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET}, at={[yshift=-\yoff]SHELFMELTa.right}}
+    \sankeyoutin[out-blue]{FRONTRET}{FRONTRETa};
 
-    \sankeynode{name=f8a, quantity=\qq{f8}, at={[yshift=\yoff]CALVa.left}, anchor=right}
-    \sankeyoutin[ice-to-water]{f8}{f8a}
+    \sankeynode{name=GZa, quantity=\qq{GZ}, at={[yshift=-\yoff]FRONTRETa.right}, anchor=left}
 
-    \sankeynode{name=GZa, quantity=\qq{GZ}, at={[yshift=-\yoff]SUB.right}, anchor=left}
-    \sankeyoutin{GZ}{GZa}
+    \sankeynode{name=BMBa, quantity=\qq{BMB}, at={[yshift=-\yoff]GZa.right}, anchor=left}
+    \sankeyoutin[ice-to-water]{BMB}{BMBa}; \sankeyadvance[water]{BMBa}{10mm};
+    \sankeyend[water]{BMBa}; \labelright{BMBa}{vBMB}
 
-    \sankeynode{name=FLa, quantity=\qq{FL}, at={[yshift=-\yoff]GZa.right}, anchor=left}
-    \sankeyoutin{FL}{FLa}
+    \sankeyoutin[ice-to-water]{GZ}{GZa}; \sankeyadvance[water]{GZa}{10mm};
+    \sankeyend[water]{GZa}; \labelbold{GZa}{vGZRET}
+    \sankeyadvance[out-blue]{FRONTRETa}{10mm}; 
+    \sankeyend[out-blue]{FRONTRETa}; \labelbold{FRONTRETa}{vFRONTRET}
 
-    \sankeynode{name=BMa, quantity=\qq{BM}, at={[yshift=-\yoff]FLa.right}, anchor=left}
-    \sankeyoutin[ice-to-water]{BM}{BMa}
+    \sankeyadvance{CALV}{5mm};
+    \sankeyoutin{CALV}{CALVa};
 
-    \sankeyadvance[out-red]{FLa}{10mm}
-    \sankeyend[out-red]{FLa}
-    \labelright{FLa}{vFRLOSS}
+    %% SMB outputs
+    \sankeynode{name=RUEVa, quantity=\qq{RUEV}, at={[yshift=\yoff]CALVa.left}, anchor=right}
+    \sankeyoutin[ice-to-water]{RUEV}{RUEVa}
+    \sankeyfork{RUEVa}{\Q{vEV}/EV, \Q{vRU}/RU}
+    \sankeynode{name=EVa, quantity=\qq{EV},
+      at={[yshift=\yoff,xshift=5mm]RUEVa.left}, anchor=left, angle=45}
+    \sankeyoutin[water-to-air]{EV}{EVa}; \sankeyadvance[air]{EVa}{1mm}
+    \sankeyend[air]{EVa};
+    \node[font=\tiny, anchor=west] at ([yshift=0mm,xshift=\yoff]EVa.west) {\qq{EVa}: \Lab{vEV}};
 
-    \ifnum \numexpr\Q{vMG} > \Q{vML}
-      \sankeynode{name=MGa, quantity=\qq{MG}, angle=-35,
-        at={[xshift=5mm,yshift=-3mm]MG.right}, anchor=right}
-      \node[font=\tiny, anchor=east]
-      at ([yshift=0mm,xshift=0mm]MGa.right) {Mass gain: \qq{MGa}};
-      \sankeyoutin[ice-to-black]{MG}{MGa}
-    \fi
+    \sankeyadvance[water]{RU}{10mm}; \sankeyend[water]{RU}; \labelright{RU}{vRU}
 
-    \sankeyfork{f8a}{\Q{vEV}/EV, \Q{vRU}/RU}
-    \sankeynode{name=EVa, quantity=\qq{EV},
-      at={[yshift=2mm,xshift=5mm]f8a.left}, anchor=left, angle=45}
-    \sankeyoutin[water-to-air]{EV}{EVa}
-    \sankeyadvance[air]{EVa}{1mm}
-    \sankeyend[air]{EVa}
+    \sankeynode{name=SUa, quantity=\qq{SU}, at={[xshift=-6mm]EVa.left}, anchor=right, angle=45}
+    \sankeydubins[minimum radius=6mm,ice-to-air]{SU}{SUa}; \sankeyend[air]{SUa}
+    \node[font=\tiny, anchor=west] at ([yshift=4mm,xshift=\yoff]SUa.west) {\qq{SUa}: \Lab{vSU}};
 
-    \sankeynode{name=SUa, quantity=\qq{SU},
-      at={[xshift=-6mm]EVa.left}, anchor=right, angle=45}
-    %% \sankeyoutin[ice-to-air{SU}{SUa}
-    \sankeydubins[minimum radius=6mm,ice-to-air]{SU}{SUa}
-    \sankeyend[air]{SUa}
+    \sankeyadvance{CALVa}{10mm}; \sankeyend{CALVa}; \labelright{CALVa}{vCALV}
 
@@
       \node[font=\tiny, anchor=east]
-      at ([xshift=-3mm,yshift=2mm]ML.west) {Mass loss: \qq{ML}};
+      at ([xshift=-3mm,yshift=2mm]ML.west) {\textbf{Drawdown or retreat: \qq{ML}}};
 
       % frontal advance next to drawdown (mass loss)
-      \sankeynode{name=FGi, quantity=\Q{vFRGAIN},
+      \sankeynode{name=FRONTADVi, quantity=\Q{vFRONTADV},
         at={[yshift=0mm,xshift=6mm]ML.left}, anchor=right, angle=-45}
@@
     }{ % mass gain
-      \sankeyadvance[black]{MGa}{3mm}
+      \sankeynode{name=MGa, quantity=\qq{MG}, angle=-35,
+        at={[xshift=5mm,yshift=-3mm]MG.right}, anchor=right}
+      \node[font=\tiny, anchor=east]
+        at ([yshift=0mm,xshift=0mm]MGa.right) {Build-up or expansion: \qq{MGa}};
+      \sankeyoutin[ice-to-black]{MG}{MGa}; \sankeyadvance[black]{MGa}{3mm}
       \sankeyend[black]{MGa}
-      \sankeynode{name=FGi, quantity=\Q{vFRGAIN},
+      \sankeynode{name=FRONTADVi, quantity=\Q{vFRONTADV},
         at={[xshift=-15mm,yshift=8mm]RFZi.right}, anchor=left, angle=-45}
@@
 
-    \ifnum \numexpr\qq{FGi} > 0
-    \sankeystart[in-black]{FGi}
-    \sankeyoutin[in-black]{FGi}{FG}
+    \ifnum \numexpr\qq{FRONTADVi} > 0
+      \sankeystart[in-black]{FRONTADVi}
+      \sankeyoutin[in-black]{FRONTADVi}{FG}
       \node[font=\tiny, anchor=east]
-    at ([xshift=-3mm,yshift=3mm]FGi.west) {Frontal advance: \qq{FGi}};
+          at ([xshift=-3mm,yshift=3mm]FRONTADVi.west) {Frontal advance: \qq{FRONTADVi}};
     \fi
 
-    \sankeyadvance[ice-to-water,out-red]{GZa}{10mm}
-    \sankeyend[water,out-red]{GZa}
-    \sankeyadvance{CALVa}{10mm}
-    \sankeyend{CALVa}
-
-    \sankeyset{water}
-    \sankeyadvance{RU}{10mm}; \sankeyend{RU}
-    \sankeyadvance{BMa}{10mm}; \sankeyend{BMa}
-    \sankeyadvance{SUB}{10mm}; \sankeyend{SUB}
-
-    \labelright{RU}{vRU}
-    \labelright{BMa}{vBM}
-    \labelright{SUB}{vSUB}
-    \labelright{CALVa}{vICE}
-    \labelright{GZa}{vGZRET}
-
-    \node[font=\tiny, anchor=west]
-    at ([yshift=4mm,xshift=\yoff]SUa.west) {\qq{SUa}: \Lab{vSU}};
-
-    \node[font=\tiny, anchor=west]
-    at ([yshift=0mm,xshift=\yoff]EVa.west) {\qq{EVa}: \Lab{vEV}};
-
-    \node[font=\tiny, anchor=east]
-    at ([yshift=-0.5mm,xshift=1mm]RFZo.center) {\Lab{vRFZ}: \qq{RFZi}};
+    %% % fancy labels
+    %% \sankeynode[debug]{name=Discharge, at={[xshift=-10mm]DISCH.north}, anchor=left,
+    %%   quantity=\Q{vDISCH}}
+    %% \sankeynode[debug]{name=SMB Inputs, at={[xshift=3mm]SMBIN.north}, anchor=left,
+    %%   quantity=\Q{vSMBin}}
+    %% \sankeynode[debug]{name=SMB Outputs, at={[xshift=0mm]RFZo.south}, anchor=left,
+    %%   quantity=\Q{vSMBout}}
 

### 2025-06-23 sankey.tex
Remove bolded text from figures

@@
       at ([yshift=0mm,xshift=3mm]#1.center) {\qq{#1}: \Lab{#2}};
-    }
-    \newcommand\labelbold[2]{ % valname
-      \node[font=\tiny, anchor=west]
-      at ([yshift=0mm,xshift=3mm]#1.center) {\textbf{\qq{#1}: \Lab{#2}}};
     }
@@
     \sankeyoutin[ice-to-water]{GZ}{GZa}; \sankeyadvance[water]{GZa}{10mm};
-    \sankeyend[water]{GZa}; \labelbold{GZa}{vGZRET}
+    \sankeyend[water]{GZa}; \labelright{GZa}{vGZRET}
     \sankeyadvance[out-blue]{FRONTRETa}{10mm}; 
-    \sankeyend[out-blue]{FRONTRETa}; \labelbold{FRONTRETa}{vFRONTRET}
+    \sankeyend[out-blue]{FRONTRETa}; \labelright{FRONTRETa}{vFRONTRET}
 
@@
       \node[font=\tiny, anchor=east]
-      at ([xshift=-3mm,yshift=2mm]ML.west) {\textbf{Drawdown or retreat: \qq{ML}}};
+      at ([xshift=-3mm,yshift=2mm]ML.west) {Drawdown or retreat: \qq{ML}};
 

### 2025-06-23 sankey.tex
Use 'mass loss (gain)' not 'drawdown/retreat (buildup/expansion)'

@@
       \node[font=\tiny, anchor=east]
-      at ([xshift=-3mm,yshift=2mm]ML.west) {Drawdown or retreat: \qq{ML}};
+      at ([xshift=-3mm,yshift=2mm]ML.west) {Mass loss: \qq{ML}};
 
@@
       \node[font=\tiny, anchor=east]
-        at ([yshift=0mm,xshift=0mm]MGa.right) {Build-up or expansion: \qq{MGa}};
+        at ([yshift=0mm,xshift=0mm]MGa.right) {Mass gain: \qq{MGa}};
       \sankeyoutin[ice-to-black]{MG}{MGa}; \sankeyadvance[black]{MGa}{3mm}

### 2025-08-20 sankey.tex
Remove mass loss/gain arrows. Replace w/ truncated section and label

@@
 
-\begin{tikzpicture}[remember picture, overlay, shift={(5,-5)}]
+\begin{tikzpicture}[remember picture, overlay, shift={(1,-5)}]
 
@@
         rounded corners=.1em,
-        %% node contents={\name: \qty},
-        node contents={\qty},
+        node contents={\name: \qty},
+        %node contents={\qty},
         %% node contents={\name},
@@
     \ifthenelse{\Q{vML} > \Q{vMG}} { % mass loss
-      \sankeynode{name=f1, at={[xshift=30mm, yshift=\yoff]SHELFFREEZE.south}, anchor=right,
+      \sankeynode{name=f1, at={[xshift=30mm, yshift=0mm]RFi.north}, anchor=left,
         quantity=\qq{SMBIN}+\Q{vFRONTADV}+\Q{vSHELFFREEZE}+\Q{vML}+\Q{vRFZ},
-        forked={\Q{vRFZ}/RFZi, \Q{vFRONTADV}/FG, \Q{vML}/MLi,
-          \qq{SMBIN}/SMBINi, \Q{vSHELFFREEZE}/SHELFFREEZEi}}
+        forked={\Q{vRFZ}/RFZi, \qq{SMBIN}/SMBINi, \Q{vSHELFFREEZE}/SHELFFREEZEi,
+          \Q{vML}/MLi, \Q{vFRONTADV}/FG}}
     } { % mass gain edge case
-      \sankeynode{name=f1, at={[xshift=30mm, yshift=\yoff]SHELFFREEZE.south}, anchor=right,
+      \sankeynode{name=f1, at={[xshift=30mm, yshift=0mm]RFi.north}, anchor=left,
         quantity=\qq{SMBIN}+\Q{vFRONTADV}+\Q{vSHELFFREEZE}+\Q{vRFZ},
-        forked={\Q{vRFZ}/RFZi, \Q{vFRONTADV}/FG, \qq{SMBIN}/SMBINi, \Q{vSHELFFREEZE}/SHELFFREEZEi}}
+        forked={\Q{vRFZ}/RFZi, \qq{SMBIN}/SMBINi, \Q{vSHELFFREEZE}/SHELFFREEZEi, \Q{vFRONTADV}/FG}}
     }
@@
     \ifthenelse{\Q{vMG} < \Q{vML}} { % mass loss
-      \sankeynode{name=ML, quantity=\Q{vML},
-        at={[yshift=6mm]RF.north}, anchor=right, angle=-45}
-      \node[font=\tiny, anchor=east]
-      at ([xshift=-3mm,yshift=2mm]ML.west) {Mass loss: \qq{ML}};
-
-      % frontal advance next to drawdown (mass loss)
-      \sankeynode{name=FRONTADVi, quantity=\Q{vFRONTADV},
-        at={[yshift=0mm,xshift=6mm]ML.left}, anchor=right, angle=-45}
-
-      \sankeystart[red]{ML}
-      \sankeyadvance[red]{ML}{5mm}
-      \sankeyoutin[red-to-ice]{ML}{MLi}
+      \sankeynode{name=ML, quantity=\Q{vML}, at={[xshift=-15mm,yshift=-\yoff]MLi.south}, anchor=right}
+      \sankeyadvance{ML}{1mm} \sankeydubins{ML}{MLi}
+      \draw[<-,gray] ([xshift=-2mm]ML.west) -- ++(-10mm,0)
+      node[left,font=\tiny,anchor=east,text=black] {Mass loss: \Q{vML}};
     }{ % mass gain
-      \sankeynode{name=MGa, quantity=\qq{MG}, angle=-35,
-        at={[xshift=5mm,yshift=-3mm]MG.right}, anchor=right}
-      \node[font=\tiny, anchor=east]
-        at ([yshift=0mm,xshift=0mm]MGa.right) {Mass gain: \qq{MGa}};
-      \sankeyoutin[ice-to-black]{MG}{MGa}; \sankeyadvance[black]{MGa}{3mm}
-      \sankeyend[black]{MGa}
-      \sankeynode{name=FRONTADVi, quantity=\Q{vFRONTADV},
-        at={[xshift=-15mm,yshift=8mm]RFZi.right}, anchor=left, angle=-45}
+      \sankeyturnleft{MG}{-15} \sankeyadvance{MG}{10mm} % \sankeyturnleft{MG}{15}
+      \draw[<-,gray] ([xshift=1mm]MG.east)
+      .. controls +(5mm,0) and +(0,0mm) .. ++(5mm,-5mm)
+      node[below,font=\tiny,text=black] {Mass gain: \qq{MG}};
     }
 
-    \ifnum \numexpr\qq{FRONTADVi} > 0
-      \sankeystart[in-black]{FRONTADVi}
-      \sankeyoutin[in-black]{FRONTADVi}{FG}
-      \node[font=\tiny, anchor=east]
-          at ([xshift=-3mm,yshift=3mm]FRONTADVi.west) {Frontal advance: \qq{FRONTADVi}};
+    \ifnum \numexpr\Q{vFRONTADV} > 0
+      \sankeynode{name=FG0, quantity=\qq{FG}, at={[xshift=-15mm,yshift=-6mm]FG.south}, anchor=right}
+      \sankeyadvance{FG0}{1mm} \sankeydubins{FG0}{FG}
+
+      \draw[<-,gray] ([xshift=-2mm]FG0.west) -- ++(-10mm,0)
+      node[left,font=\tiny,anchor=east,text=black] {Frontal advance: \qq{FG}};
     \fi

### 2025-08-21 sankey.tex
Properly handle frontal advance and retreat
Make frontal advance optional in Sankey diagrams

First, unify advance and retreat

If advance > retreat,
  subtract retreat from advance, and set retreat to 0
If advance < retreat
  subtract advance from retreat and set advance to 0

Then, put it in terms of calving, or calving and frontal melt

AQ: Add retreat and subtract advance from calving

GL: Add (retreat - advance)/2 to calving and again frontal melt.

@@
       %% fill/.style={fill=yellow!10,line width=0pt,draw=yellow!10},
-      ratio=1mm/60,
+      ratio=1mm/40,
       minimum radius=0mm,
@@
 
+      \ifnum \numexpr\Q{vFRONTRET} > 0
         \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET},
           at={[yshift=-\yoff]SHELFMELT.right}}
+        \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET},
+          at={[yshift=-\yoff]SHELFMELTa.right}, anchor=left}
+      \fi
       \sankeynode{name=CALVa, quantity=\qq{CALV},
         at={[yshift=\yoff,xshift=0mm]SHELFMELTa.left}, anchor=right}
-      \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET},
-        at={[yshift=-\yoff]SHELFMELTa.right}, anchor=left}
     }
@@
     %% \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET}, at={[yshift=-\yoff]SHELFMELTa.right}}
+    \ifthenelse{\Q{vFRONTRET} > 0} {
       \sankeyoutin[out-blue]{FRONTRET}{FRONTRETa};
-
       \sankeynode{name=GZa, quantity=\qq{GZ}, at={[yshift=-\yoff]FRONTRETa.right}, anchor=left}
+      \sankeyadvance[out-blue]{FRONTRETa}{10mm}; 
+      \sankeyend[out-blue]{FRONTRETa}; \labelright{FRONTRETa}{vFRONTRET}
+    } {
+      \sankeynode{name=GZa, quantity=\qq{GZ}, at={[xshift=-10mm,yshift=-\yoff]SHELFMELTa.right}, anchor=left}
+    }
 
@@
     \sankeyend[water]{GZa}; \labelright{GZa}{vGZRET}
-    \sankeyadvance[out-blue]{FRONTRETa}{10mm}; 
-    \sankeyend[out-blue]{FRONTRETa}; \labelright{FRONTRETa}{vFRONTRET}
 

### 2025-08-21 sankey.tex
Change figure ratio

@@
 
-      \ifnum \numexpr\Q{vFRONTRET} > 0
       \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET},
         at={[yshift=-\yoff]SHELFMELT.right}}
+      \sankeynode{name=CALVa, quantity=\qq{CALV},
+        at={[yshift=\yoff,xshift=0mm]SHELFMELTa.left}, anchor=right}
       \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET},
         at={[yshift=-\yoff]SHELFMELTa.right}, anchor=left}
-      \fi
-      \sankeynode{name=CALVa, quantity=\qq{CALV},
-        at={[yshift=\yoff,xshift=0mm]SHELFMELTa.left}, anchor=right}
     }
@@
     %% \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET}, at={[yshift=-\yoff]SHELFMELTa.right}}
-    \ifthenelse{\Q{vFRONTRET} > 0} {
     \sankeyoutin[out-blue]{FRONTRET}{FRONTRETa};
+
     \sankeynode{name=GZa, quantity=\qq{GZ}, at={[yshift=-\yoff]FRONTRETa.right}, anchor=left}
-      \sankeyadvance[out-blue]{FRONTRETa}{10mm}; 
-      \sankeyend[out-blue]{FRONTRETa}; \labelright{FRONTRETa}{vFRONTRET}
-    } {
-      \sankeynode{name=GZa, quantity=\qq{GZ}, at={[xshift=-10mm,yshift=-\yoff]SHELFMELTa.right}, anchor=left}
-    }
 
@@
     \sankeyend[water]{GZa}; \labelright{GZa}{vGZRET}
+    \sankeyadvance[out-blue]{FRONTRETa}{10mm}; 
+    \sankeyend[out-blue]{FRONTRETa}; \labelright{FRONTRETa}{vFRONTRET}
 

### 2025-08-21 sankey.tex
No FRONTRET or FRONTADV in Sankey (not needed when 0)

--- 2025-08-21 sankey.tex
@@
         {vSMBin/SMB Inputs/\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP}},
-        {vINPUT/Inputs/\Q{vSMBin}+\Q{vFRONTADV}+\Q{vSHELFFREEZE}},
+        {vINPUT/Inputs/\Q{vSMBin}+\Q{vSHELFFREEZE}},
         {vSMBout/SMB Outputs/\Q{vSU}+\Q{vRU}+\Q{vEV}},
-        {vDYN/Dynamics/\Q{vCALV}+\Q{vFRONTMELT}+\Q{vSHELFMELT}+\Q{vFRONTRET}+\Q{vGZRET}+\Q{vBMB}},
+        {vDYN/Dynamics/\Q{vCALV}+\Q{vFRONTMELT}+\Q{vSHELFMELT}+\Q{vGZRET}+\Q{vBMB}},
         {vOUTPUT/Outputs/\Q{vSMBout}+\Q{vDYN}},
@@
       \sankeynode{name=f1, at={[xshift=30mm, yshift=0mm]RFi.north}, anchor=left,
-        quantity=\qq{SMBIN}+\Q{vFRONTADV}+\Q{vSHELFFREEZE}+\Q{vML}+\Q{vRFZ},
+        quantity=\qq{SMBIN}+\Q{vSHELFFREEZE}+\Q{vML}+\Q{vRFZ},
         forked={\Q{vRFZ}/RFZi, \qq{SMBIN}/SMBINi, \Q{vSHELFFREEZE}/SHELFFREEZEi,
-          \Q{vML}/MLi, \Q{vFRONTADV}/FG}}
+          \Q{vML}/MLi}}
     } { % mass gain edge case
       \sankeynode{name=f1, at={[xshift=30mm, yshift=0mm]RFi.north}, anchor=left,
-        quantity=\qq{SMBIN}+\Q{vFRONTADV}+\Q{vSHELFFREEZE}+\Q{vRFZ},
-        forked={\Q{vRFZ}/RFZi, \qq{SMBIN}/SMBINi, \Q{vSHELFFREEZE}/SHELFFREEZEi, \Q{vFRONTADV}/FG}}
+        quantity=\qq{SMBIN}+\Q{vSHELFFREEZE}+\Q{vRFZ},
+        forked={\Q{vRFZ}/RFZi, \qq{SMBIN}/SMBINi, \Q{vSHELFFREEZE}/SHELFFREEZEi}}
     }
@@
         \Q{vCALV}/CALV, \Q{vFRONTMELT}/FRONTMELT, \Q{vSHELFMELT}/SHELFMELT,
-        \Q{vFRONTRET}/FRONTRET, \Q{vGZRET}/GZ, \Q{vBMB}/BMB}
+        \Q{vGZRET}/GZ, \Q{vBMB}/BMB}
     }{ % mass gain edge case (AQ_east)
@@
         \Q{vCALV}/CALV, \Q{vFRONTMELT}/FRONTMELT, \Q{vSHELFMELT}/SHELFMELT,
-        \Q{vFRONTRET}/FRONTRET, \Q{vGZRET}/GZ, \Q{vBMB}/BMB, \Q{vMG}/MG}
+        \Q{vGZRET}/GZ, \Q{vBMB}/BMB, \Q{vMG}/MG}
     }
@@
       %% Discharge splits to FRONTMELT (straight) & CALV (above).
-      %% SHELFMELT below FRONTMELT. FRONTRET below SHELFMELT.
+      %% SHELFMELT below FRONTMELT.
       \sankeyadvance{FRONTMELT}{20mm}
@@
         at={[yshift=-\yoff, xshift=10mm]FRONTMELT.right}}
-      \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET},
-        at={[yshift=-\yoff]SHELFMELTa.right}}
       \sankeyadvance[ice-to-water]{FRONTMELT}{10mm}
@@
       %% Discharge splits to SHELFMELT (straight) & CALV (above).
-      %% No FRONTMELT. Then FRONTRET
+      %% No FRONTMELT.
       \sankeyadvance{SHELFMELT}{10mm}
@@
         at={[yshift=0, xshift=20mm]SHELFMELT.right}, anchor=right}
-
-      \ifnum \numexpr\Q{vFRONTRET} > 0
-        \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET},
-          at={[yshift=-\yoff]SHELFMELT.right}}
-        \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET},
-          at={[yshift=-\yoff]SHELFMELTa.right}, anchor=left}
-      \fi
       \sankeynode{name=CALVa, quantity=\qq{CALV},
@@
 
-    %% \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET}, at={[yshift=-\yoff]SHELFMELTa.right}}
-    \ifthenelse{\Q{vFRONTRET} > 0} {
-      \sankeyoutin[out-blue]{FRONTRET}{FRONTRETa};
-      \sankeynode{name=GZa, quantity=\qq{GZ}, at={[yshift=-\yoff]FRONTRETa.right}, anchor=left}
-      \sankeyadvance[out-blue]{FRONTRETa}{10mm}; 
-      \sankeyend[out-blue]{FRONTRETa}; \labelright{FRONTRETa}{vFRONTRET}
-    } {
     \sankeynode{name=GZa, quantity=\qq{GZ}, at={[xshift=-10mm,yshift=-\yoff]SHELFMELTa.right}, anchor=left}
-    }
 
@@
 
-    \ifnum \numexpr\Q{vFRONTADV} > 0
-      \sankeynode{name=FG0, quantity=\qq{FG}, at={[xshift=-15mm,yshift=-6mm]FG.south}, anchor=right}
-      \sankeyadvance{FG0}{1mm} \sankeydubins{FG0}{FG}
-
-      \draw[<-,gray] ([xshift=-2mm]FG0.west) -- ++(-10mm,0)
-      node[left,font=\tiny,anchor=east,text=black] {Frontal advance: \qq{FG}};
-    \fi
-
     %% % fancy labels

### 2025-08-22 sankey.tex
Increase Sankey label font size

@@
     \newcommand\labelleft[1]{ % valname
-      \node[font=\tiny, anchor=east]
+      \node[font=\small, anchor=east]
       at ([yshift=0mm,xshift=-3mm]#1.center) {\Lab{v#1}: \qq{#1}};
@@
     \newcommand\labelright[2]{ % valname
-      \node[font=\tiny, anchor=west]
+      \node[font=\small, anchor=west]
       at ([yshift=0mm,xshift=3mm]#1.center) {\qq{#1}: \Lab{#2}};
@@
     \sankeydubins[minimum radius=5mm,ice-to-water-vert]{RFZo}{RFZi}
-    \node[font=\tiny, anchor=east]
-           at ([yshift=+1.5mm,xshift=1mm]RFZo.north) {\Lab{vRFZ}: \qq{RFZi}};
+    \node[font=\small, anchor=east]
+           at ([yshift=+4.5mm,xshift=5mm]RFZo.north) {\Lab{vRFZ}: \qq{RFZi}};
 
@@
     \sankeyend[air]{EVa};
-    \node[font=\tiny, anchor=west] at ([yshift=0mm,xshift=\yoff]EVa.west) {\qq{EVa}: \Lab{vEV}};
+    \node[font=\small, anchor=west] at ([yshift=0mm,xshift=\yoff]EVa.west) {\qq{EVa}: \Lab{vEV}};
 
@@
     \sankeydubins[minimum radius=6mm,ice-to-air]{SU}{SUa}; \sankeyend[air]{SUa}
-    \node[font=\tiny, anchor=west] at ([yshift=4mm,xshift=\yoff]SUa.west) {\qq{SUa}: \Lab{vSU}};
+    \node[font=\small, anchor=west] at ([yshift=4mm,xshift=\yoff]SUa.west) {\qq{SUa}: \Lab{vSU}};
 
@@
       \draw[<-,gray] ([xshift=-2mm]ML.west) -- ++(-10mm,0)
-      node[left,font=\tiny,anchor=east,text=black] {Mass loss: \Q{vML}};
+      node[left,font=\small,anchor=east,text=black] {Mass loss: \Q{vML}};
     }{ % mass gain
@@
       .. controls +(5mm,0) and +(0,0mm) .. ++(5mm,-5mm)
-      node[below,font=\tiny,text=black] {Mass gain: \qq{MG}};
+      node[below,font=\small,text=black] {Mass gain: \qq{MG}};
     }
